### 2020-10-27 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Alison added her sociodem files

```{r Setup, include = FALSE, results = 'hide', warning = FALSE}
library(knitr)
library(tidycensus)
library(tidyverse)
library(ggplot2)
```

```{r}
acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)

view(acs5_2018_vars)
```

```{r}
sex_age_vector <- acs5_2018_vars %>%
  select(name) %>%

```

### 2020-10-27 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Updated tidycensus doc

@@
 library(ggplot2)
+
+Sys.getenv("CENSUS_API_KEY")
 ```
@@
   select(name) %>%
+  slice(5:51)
+```
 
+```{r}
+MI_sex_age_acs5_2018 <- get_acs(geography = "place",
+              variables = c(sex_age_vector),
+              state = "MI",
+              year = 2018)
+
+view(MI_sex_age_acs5_2018)
 ```

### 2020-10-28 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Update Social_Demo_Chars_tidycensus.Rmd

@@
 acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
-
-view(acs5_2018_vars)
 ```
@@
 ```{r}
-sex_age_vector <- acs5_2018_vars %>%
+sex_age_vars <- acs5_2018_vars %>%
+  slice(5:51)
+
+sex_age_vec <- sex_age_vars %>%
   select(name) %>%
-  slice(5:51)
+  unlist()
+
+sex_age_label <- sex_age_vars %>%
+  select(label) %>%
+  unlist()
 ```
 
-```{r}
+```{r, message = FALSE}
 MI_sex_age_acs5_2018 <- get_acs(geography = "place",
-              variables = c(sex_age_vector),
+              variables = sex_age_vec,
+              place = "Manistee",
               state = "MI",
@@
 
-view(MI_sex_age_acs5_2018)
+Manistee_sex_age_acs5_2018 <- MI_sex_age_acs5_2018 %>%
+  slice(18425:18471) %>%
+  mutate(sex_age_label = sex_age_label) %>%
+  select(sex_age_label, estimate)
 ```

### 2020-10-29 KPMG Team/Visualizations/Louisa - Map/Metro Areas/metro_areas.Rmd
Metro Areas & Moving Files

---
title: "Metropolitan Areas (Deaths & Case)"
output: html_notebook
---

```{r}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus) # US census bureau data
library(zoo) # rolling mean

# plotting 
library(ggplot2)
library(hrbrthemes) 
library(geofacet)

# ----------- LOADING DATASETS ----------- #

# metropolitan state areas delineation from file taken off Census Bureau
data <- read_csv("msa.csv")

#nytimes county-level cases & deaths data from online link 
nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")

# census data API, variables can be changed here 
census <- get_acs(geography = "county",
                   variables = c(population = "B01003_001E"), # total population), # black veterans 
                   output = "wide",
                   county = "",
                   year = 2018)

```

```{r}
# ------------------------------------------------------- #
# -----------  DATA MANIPULATION FOR CBSA --------------- # 
# ------------------------------------------------------- #

# --- *** --- core-based statistical areas  --- *** --- 
# comprised of counties 
# is a U.S. geographic area defined by the Office of Management and Budget (OMB) that consists of one or more counties (or equivalents) anchored by an urban center of at least 10,000 people plus adjacent counties that are socioeconomically tied to the urban center by commuting.
# https://en.wikipedia.org/wiki/List_of_metropolitan_statistical_areas

# select only cbsa-related information 
data_cbsa <- data %>% 
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
  mutate(fips_state = `FIPS State Code`) %>%
  mutate(state = `State Name`) %>%
  mutate(fips_county = `FIPS County Code`) %>%
  rename(cbsa = `CBSA Code`) %>%
  rename(cbsa_title = `CBSA Title`) %>% 
  select(cbsa, cbsa_title, fips, state, fips_state, fips_county)

# join cbsa data and census data by GEOID/FIPS, get population (and other demographic) info.
census_cbsa_merge <- census %>% 
  rename(fips = GEOID) %>% 
  left_join(data_cbsa, by = "fips") %>% 
  group_by(cbsa) %>%
  mutate(cbsa_population = sum(population))

# -------- FILTER OR NOT? ----------- # 
# only top 12 csa and census data (run this)

# list of top cbsas by population, if we want
# top_cbsa_popn <- census_cbsa_merge %>%
#   select(cbsa, cbsa_title, cbsa_population) %>%
#   arrange(desc(cbsa_population)) %>%
#   distinct() %>%
#   na.omit() %>%
#   data.frame() %>%
#   slice(1:12)

# census_cbsa <- census_cbsa_merge %>%
#   semi_join(top_cbsa_popn, by="cbsa")

# --------------------------------------# 
# without filtering (run this)
census_cbsa <- census_cbsa_merge
# ------------------------------------ #

# only need county to merge both databsets, will keep just metro areas (csa)
cbsa_covid_merge <- census_cbsa %>% 
  left_join(nyt, by = "fips") %>% 
  select(cbsa, cbsa_title, cbsa_population, date, cases, deaths) %>% 
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date("2020-10-18")) # so that the 7 day mean wouldn't get messed up

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
cbsa_covid_cases <- data.frame(cbsa_covid_merge) %>% 
  group_by(cbsa, date) %>% 
  dplyr::summarise(cases_sum = sum(cases, na.rm = TRUE)) %>% 
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

# same as above, but for deaths
cbsa_covid_deaths <- data.frame(cbsa_covid_merge) %>% 
  group_by(cbsa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
cbsa_covid <- cbsa_covid_merge %>% 
  left_join(cbsa_covid_cases, by = c("cbsa", "date")) %>% 
  left_join(cbsa_covid_deaths, by = c("cbsa", "date")) %>% 
  arrange(desc(cbsa_population))

# final count of cbsa cases, in order to get ranking
cbsa_covid_sum <- cbsa_covid %>% 
  group_by(cbsa, cbsa_title, cbsa_population) %>%
  mutate(cases = sum(cases)) %>% 
  mutate(deaths = sum(deaths)) %>%
  arrange(desc(cases)) %>% 
  select(cbsa, cbsa_title, cbsa_population, cases, deaths) %>% 
  distinct() %>% 
  na.omit()

# ----------------- MOST USEFUL DATASETS ------------------------- #

# time series data of all new cases per day (7-day rolling average) during time period specified
# names: names(cbsa_covid)
head(cbsa_covid, 10)

# Rank of CBSAs according to total cases
head(cbsa_covid_sum, 10)

# ----------------------------------------------------------------- #
```

```{r}

# ------------------------------------------------------------------- #
# -----------  DATA MANIPULATION TO GET DATA BY STATE --------------- # 
# ------------------------------------------------------------------- #

# similarly, calculate rolling daily 7-day average separately for cases
nyt_state_cases <- nyt %>% 
  group_by(state, date) %>%
  dplyr::summarise(cases_sum = sum(cases, na.rm = TRUE)) %>% 
  mutate(new_cases_state = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_state_mean = rollmean(new_cases_state, k = 7, fill = NA))

# rolling daily 7-day average for deaths
nyt_state_deaths <- nyt %>% 
  group_by(state, date) %>%
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths_state = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_state_mean = rollmean(new_deaths_state, k = 7, fill = NA))

# combined total figures
state_covid <- nyt_state_cases %>% 
  left_join(nyt_state_deaths, by = c("state", "date")) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
         date <= lubridate::as_date("2020-10-18")

# preview
head(state_covid, 10)
```

```{r}

# -------------------- FUNCTION TO MAKE A PLOT FOR ONE STATE  ------------------- #

stateData <- function(state_name) {
  state_covid %>% 
    filter(state == state_name)
}

metroData <- function(cbsa_id) {
  cbsa_covid %>% 
    filter(cbsa == cbsa_id)
}

covidCasesPlot <- function(state_data, state_name, metro_data) {

  ggplot() + 
    geom_line(aes(x = state_data$date, y = state_data$new_cases_state_mean, color = "State"), size = 1.2) + 
    geom_line(aes(x = metro_data$date, y = metro_data$new_cases_mean, color = "Metropolitan Area"), size = 1.2) + 
    labs(x = "\nTime Period", 
       y = "7-day average of Cases and Deaths\n",
       title = paste("COVID-19 Cases in", state_name, "and Metropolitan Area\n")) + 
    scale_color_manual(name = "Colors", 
                       values = c("State" = "coral2", "Metropolitan Area" = "cadetblue")) +
    theme_minimal()

}

covidDeathsPlot <- function(state_data, state_name, metro_data) {

  ggplot() + 
    geom_line(aes(x = state_data$date, y = state_data$new_deaths_state_mean, color = "State"), size = 1.2) + 
    geom_line(aes(x = metro_data$date, y = metro_data$new_deaths_mean, color = "Metropolitan Area"), size = 1.2) + 
    labs(x = "\nTime Period", 
       y = "7-day average of Cases and Deaths\n",
       title = paste("COVID-19 Deaths in", state_name, "and Metropolitan Area\n")) + 
    scale_color_manual(name = "Colors", 
                       values = c("State" = "coral4", "Metropolitan Area" = "darkcyan")) +
    theme_minimal()

}

```

`
```{r}

# --------------------- TO MAKE A PLOT FOR ONE STATE -------------------------- #

# Miami & Florida 
fl_covid <- stateData("Florida")
miami_covid <- metroData(33100) # from cbsa table

a <- covidCasesPlot(fl_covid, "Florida State", miami_covid) # cases
b <- covidDeathsPlot(fl_covid, "Florida State", miami_covid) # deaths

# NYC & New York cases
ny_covid <- stateData("New York")
nyc_covid <- metroData(35620)

c <- covidCasesPlot(ny_covid, "New York State", nyc_covid) # cases
d <- covidDeathsPlot(ny_covid, "New York State", nyc_covid) # deaths

# Dallas-Fort Worth-Arlington (19100) & Texas
tx_covid <- stateData("Texas")
dal_covid <- metroData(19100)

e <- covidCasesPlot(tx_covid, "Texas State", dal_covid) # cases
f <- covidDeathsPlot(tx_covid, "Texas State", dal_covid) # deaths

```

### 2020-10-29 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Alison updated files

--- 2020-10-27 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
@@
 library(knitr)
+library(tidyverse)
 library(tidycensus)
-library(tidyverse)
 library(ggplot2)
+library(ggthemes)
+library(jcolors)
+
+Sys.getenv("CENSUS_API_KEY")
 ```
@@
 
-view(acs5_2018_vars)
+write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
 ```
@@
 ```{r}
-sex_age_vector <- acs5_2018_vars %>%
+sex_age_vars <- acs5_2018_vars %>%
+  slice(5:51)
+
+sex_age_vec <- sex_age_vars %>%
   select(name) %>%
+  unlist()
 
+sex_age_label <- sex_age_vars %>%
+  select(label) %>%
+  unlist()
 ```
+
+```{r}
+MI_place_sex_age_acs5_2018 <- get_acs(geography = "place",
+              variables = sex_age_vec,
+              place = "Manistee",
+              state = "MI",
+              year = 2018)
+
+Man_city_sex_age_acs5_2018 <- MI_sex_age_acs5_2018 %>%
+  slice(18425:18471) %>%
+  mutate(sex_age_label = sex_age_label) %>%
+  select(sex_age_label, estimate)
+```
+
+```{r}
+# https://github.com/walkerke/tidycensus/issues/131
+
+years <- lst(2013, 2018)
+
+MI_place_mob_acs5_2013_2018 <- map_dfr(
+  years,
+  ~ get_acs(
+    geography = "place",
+    variables = c("B07001_016", "B07001_033", "B07001_049", "B07001_065", "B07001_081"),
+    state = "MI",
+    place = "Manistee",
+    year = .x
+    ),
+  .id = "year"
+  )
+```
+
+```{r}
+Man_city_mob_acs5_2013_2018 <- MI_place_mob_acs5_2013_2018 %>%
+  filter(NAME == "Manistee city, Michigan") %>%
+  add_column(label = "") %>%
+  mutate(label = case_when(variable == "B07001_016" ~ "Same house",
+                           variable == "B07001_033" ~ "Same county",
+                           variable == "B07001_049" ~ "Different county",
+                           variable == "B07001_065" ~ "Different state",
+                           variable == "B07001_081" ~ "Moved from abroad")) %>%
+  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
+```
+
+```{r}
+p1 <- ggplot(data = Man_city_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
+  geom_bar(stat = "identity", position = position_dodge())
+```
+
+```{r}
+p1 + theme_fivethirtyeight() +
+  scale_fill_jcolors(palette = "pal5") +
+  theme(axis.text.x = element_text(angle = 35, hjust = 1),
+        legend.position = "top",
+        legend.title = element_blank()) +
+  ggtitle("Where Manistee residents lived compared to \nthe year before\n") +
+  labs(caption = "\nSource: 2013 and 2018 ACS 5-year surveys")
+```
+
+```{r}
+MI_place_age_mob_acs5_2013_2018 <- map_dfr(
+  years,
+  ~ get_acs(
+    geography = "place",
+    variables = c("B07002_002", "B07002_003", "B07002_004", "B07002_005", "B07002_006"),
+    state = "MI",
+    place = "Manistee",
+    year = .x
+    ),
+  .id = "year"
+  )
+```
+
+```{r}
+Man_city_age_mob_acs5_2013_2018 <- MI_place_age_mob_acs5_2013_2018 %>%
+  filter(NAME == "Manistee city, Michigan") %>%
+  add_column(label = "") %>%
+  mutate(label = case_when(variable == "B07002_002" ~ "Same house",
+                           variable == "B07002_003" ~ "Same county",
+                           variable == "B07002_004" ~ "Different county",
+                           variable == "B07002_005" ~ "Different state",
+                           variable == "B07002_006" ~ "Moved from abroad")) %>%
+  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
+```
+
+```{r}
+p2 <- ggplot(data = Man_city_age_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
+  geom_bar(stat = "identity", position = position_dodge())
+```
+
+```{r}
+p2 + theme_fivethirtyeight() +
+  scale_fill_jcolors(palette = "pal5") +
+  theme(axis.text.x = element_text(angle = 35, hjust = 1),
+        legend.position = "top",
+        legend.title = element_blank()) +
+  ggtitle("Median age of Manistee residents \nby where they lived the year before\n") +
+  labs(caption = "\nSource: 2013 and 2018 ACS 5-year surveys")
+```

### 2020-10-29 EPA:Census Team/Gretchen Walkability/Walkability Map.Rmd

---
title: "Walkability"
author: "Gretchen Streett"
date: "10/27/2020"
output: html_document
---

```{r}
library(leaflet)
library(rgdal)
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(stringr)
library(ggplot2)

#get data
walk <- readOGR(dsn = "Natl_WI_SHP/Natl_WI.shp")
mwalk <- walk[walk$SFIPS == "26", ]

statefips <- read.csv("~/Documents/Columbia/Practicum - Aracelis Torres/Data/Walkability/State FIPS.csv")
```

```{r}
# Walkability map (Michigan Specific) 
## Target End state: blue water, data for whole country, better names, search by zip code/location, round Walkability Index Score, add layers for parks/schools/nature???
walkmap <- leaflet(mwalk) %>%
 setView(lat = 44.2444, lng = -86.3243, zoom = 6) %>%
 addProviderTiles(providers$Stamen.TonerLite) %>%
  #add bloc borders
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  #chloropleth shading
  fillColor = ~colorQuantile("Greens", NatWalkInd)(NatWalkInd), 
  fillOpacity = 1, 
  #add pinpoint label
  label = ~stringr::str_c(CBSA_Name, ' ',
          formatC(NatWalkInd, big.mark = ',', format='f')),
  labelOptions = labelOptions(direction = 'auto'),
  #add hover highlight
  highlightOptions = highlightOptions(
          color='#000000', weight = 3,
          bringToFront = TRUE, sendToBack = TRUE))

walkmap2 <- walkmap %>% 
  addMiniMap() %>%
  addLegend(
  "bottomleft", 
  pal = colorQuantile("Greens", mwalk$NatWalkInd, n=5),
  values = ~NatWalkInd,
  title = htmltools::HTML("Walkability Index"),
opacity = 1, 
  labFormat = function(type, cuts, p) {
n = length(cuts)
paste0(cuts[-n], " – ", cuts[-1])}
)

walkmap3 <- walkmap2 %>%
    addEasyButton(easyButton(
    icon="fa-globe", 
    title="Zoom to Entire US Map",
    onClick=JS("function(btn, map){ map.setZoom(2.5); }"))) 
walkmap3
```

```{r}
# Comparison of walkability: Manistee to State to comparable rural areas (CRA)
# EQ: Should Manistee county be removed from other categories?, What constitutes "Comparable Rural Area", # Are there areas that we don't expect to be walkable and should thus be excluded? 
# Target End state: Interactive, compare based on alt. features/demographics

#Subset Manistee county block groups (FIPS code: 026101)
manwalk <- walk[(walk$GEOID10 >= "261010000000" & walk$GEOID10 <= "261030000000"), ]
manwalk <- manwalk@data
manwalk <- manwalk %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
manwalk <- manwalk %>% mutate("Location" = "Manistee County" )

# Subset block groups with similar Population Density 
mwalkdata <- mwalk@data
mwalkdata <- mwalkdata %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
mwalkdata <- mwalkdata[(mwalkdata$`Population Density` >= .8 & mwalkdata$`Population Density` <= 1.6), ]
mwalkdata <- mwalkdata %>% mutate("Location" = "Comparable Rural Area" )

# Compare to Michigan as a Whole
mwalkdatawhole <- mwalk@data
mwalkdatawhole <- mwalkdatawhole %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
mwalkdatawhole <- mwalkdatawhole %>% mutate("Location" = "Entire State" )

# Compare Walkability of two groups 
combined <- rbind(manwalk, mwalkdata, mwalkdatawhole)
combined <- combined %>% group_by(`Location`) %>%
            summarize("Walkability" = mean(`NatWalkInd`))

myplot <- ggplot(data = combined, aes(x=`Location`, y=`Walkability`, fill=`Location`)) + 
  geom_bar(stat="identity", width=0.5) +
  scale_fill_brewer(palette="Dark2") +
  scale_x_discrete(limits=c("Manistee County", "Comparable Rural Area", "Entire State")) +
  geom_text(aes(label= round(`Walkability`, 2)), vjust=1.6, color="white", size=3.5)+
  labs(title="Walkability by Location")+
  theme_minimal() + 
  theme(legend.position="none")
myplot

```

```{r}
# Country Comparison

# Join State Names to Data 
allwalk <- walk@data
allwalk <- separate(allwalk, GEOID10, into = "State Code", sep= 2, remove = FALSE)
allwalk$`State Code` <- as.numeric(allwalk$`State Code`)
walkwithstates <- full_join(allwalk, statefips, by = c("State Code" = "FIPS"))
walkwithstates <- walkwithstates %>% filter(`State Code` <= 56)
walkwithstates <- walkwithstates %>% drop_na(`State Code`)
walkwithstates <- walkwithstates %>% drop_na(`State`)

# Summarize for plot
walkwithstates <- walkwithstates %>% mutate("Population Density" = `TOTPOP10`/`AC_TOT` )
walkwithstates <- walkwithstates %>% 
  mutate(
      CRA = case_when(
      (`Population Density` >= .8 & `Population Density` <= 1.6) ~ "Comparable Rural Area",
      TRUE ~ "Not Comparable"))
walkfullstates<- walkwithstates %>% group_by(`State`) %>% 
              summarize("Walkability" = mean(`NatWalkInd`))
walkstates<- walkwithstates %>% group_by(`State`, `CRA`) %>% 
              summarize("Walkability" = mean(`NatWalkInd`))

# Plot

#Reorder Walkability in table to make chart pretty
walkfullstates<- transform(walkfullstates, `State` = reorder(`State`, -Walkability))
#plot it 
 plota <- ggplot(walkfullstates, aes(x=State, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Walkability by State",
       caption="source: EPA") 
plota

#Reorder Walkability for CRA 
walkstates <- walkstates %>% filter(CRA == "Comparable Rural Area")
walkstates<- transform(walkstates, `State` = reorder(`State`, -Walkability))
#plot it 
plota.2 <- ggplot(walkstates, aes(x=State, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Comparable Rural Area Walkability by State",
       caption="source: EPA") 
plota.2

# lollipop Test: is it better? 
# plotb <- ggplot(walkfullstates, aes(x=State, y=Walkability)) + 
#   geom_point(size=3) + 
#   coord_flip() +
#   geom_segment(aes(x=State, 
#                    xend=State, 
#                    y=0, 
#                    yend=Walkability)) +
#   labs(title="Ordered Bar Chart", 
#        subtitle="Walkability by State", 
#        caption="source: EPA") +
#   theme_bw()
# plotb
```

### 2020-10-29 Asahi_housing/housing.R
Add files via upload

@@ -1,0 +1,283 @@
+library(data.table)
+library(dplyr)
+library(extrafont)
+library(ggplot2)
+library(ggpubr)
+library(sf)
+library(tidycensus)
+
+loadfonts()
+
+api_key <- Sys.getenv("CENSUS_API_KEY")
+census_api_key(api_key, install = TRUE, overwrite = TRUE)
+readRenviron("~/.Renviron")
+
+vars_acs5_2018 <- load_variables(year = 2018, dataset = "acs5")
+rent_vars <- vars_acs5_2018[vars_acs5_2018$name %like% "B25063", ][3:26,]
+rent_thresholds <- regmatches(rent_vars$label, gregexpr("[[:digit:]]+", rent_vars$label))
+
+rent_threshold_labels <- c()
+for (i in c(1:length(rent_thresholds))) {
+  length <- length(rent_thresholds[[i]])
+  label <- ""
+  if ((length) == 1) {
+    rent_threshold_labels[i] <- paste("less than $",rent_thresholds[[i]], sep = "")
+  }
+  else if ((length) == 2) {
+    if (as.numeric(rent_thresholds[[i]][1]) < 100) {
+      rent_threshold_labels[i] <- paste("greater than $",rent_thresholds[[i]][1],",",rent_thresholds[[i]][2], sep = "")
+    }
+    else {
+      rent_threshold_labels[i] <- paste("$",rent_thresholds[[i]][1]," to $",rent_thresholds[[i]][2], sep = "")
+    }
+  }
+  else {
+    rent_threshold_labels[i] <- paste("$",rent_thresholds[[i]][1],",",rent_thresholds[[i]][2]," to ","$",rent_thresholds[[i]][3],",",rent_thresholds[[i]][4], sep = "")
+  }
+}
+
+manistee_rent_tract <- get_acs(geography = "tract",
+                               state = "MI",
+                               county = "Manistee",
+                               variables = rent_vars$name,
+                               year = 2018,
+                               output = "wide",
+                               geometry = TRUE)
+manistee_rent_county <- get_acs(geography = "county",
+                                state = "MI",
+                                county = "Manistee",
+                                variables = rent_vars$name,
+                                year = 2018,
+                                output = "wide",
+                                geometry = TRUE)
+
+county_rents <- manistee_rent_county[ ,grepl("B.*E|E.*B" ,names(manistee_rent_county))]
+st_geometry(county_rents) <- NULL
+
+county_rents_DF <- data.frame(thresholds = rent_threshold_labels,
[…]

### 2020-11-03 EPA:Census Team/Alisha_Education/Manistee_Educ.R
Added Education and GDP (Manistee)

@@ -1,0 +1,70 @@
+library(tidycensus)
+library(tidyverse)
+library(ggplot2)
+library(plotly)
+library(dplyr)
+
+#Data from ACS5 2018 table B15003 - Education level totals for population over 25
+
+api_key <- Sys.getenv("CENSUS_API_KEY")
+census_api_key(api_key, overwrite = TRUE, install = TRUE)
+
+readRenviron("~/.Renviron")
+
+
+acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
+
+
+#pulling data for Manistee Education
+manistee_educ2 <- get_acs(geography = "county",
+                         state = "MI",
+                         county = "Manistee",
+                         table =  'B15003',
+                         year = 2018,
+                         output = "wide",
+                         geometry = TRUE
+)
+
+
+#EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER">
+#adding columns for different education level totals
+manistee_educ2$no_degree <- manistee_educ2$B15003_002E + manistee_educ2$B15003_003E + manistee_educ2$B15003_004E + manistee_educ2$B15003_005E + manistee_educ2$B15003_006E +  
+                            manistee_educ2$B15003_007E + manistee_educ2$B15003_008E + manistee_educ2$B15003_009E + manistee_educ2$B15003_010E +
+                            manistee_educ2$B15003_011E + manistee_educ2$B15003_012E + manistee_educ2$B15003_013E + manistee_educ2$B15003_014E +
+                            manistee_educ2$B15003_015E + manistee_educ2$B15003_016E
+manistee_educ2$high_school <- manistee_educ2$B15003_017E + manistee_educ2$B15003_018E
+manistee_educ2$some_college <- manistee_educ2$B15003_019E + manistee_educ2$B15003_020E
+manistee_educ2$associates <- manistee_educ2$B15003_021E
+manistee_educ2$bachelors <- manistee_educ2$B15003_022E
+manistee_educ2$post_grad <- manistee_educ2$B15003_023E + manistee_educ2$B15003_024E + manistee_educ2$B15003_025E
+manistee_educ2$totalpop <- manistee_educ2$B15003_001E
+
+manistee_educ <- subset(manistee_educ2, select =  c(GEOID, NAME,no_degree, high_school, some_college, associates, bachelors, post_grad,totalpop))
+manistee_educ_pct<- manistee_educ %>% mutate(no_degree_pct = (no_degree/ totalpop),
[…]

### 2020-11-03 EPA:Census Team/Alisha_GDP/Manistee_GDP.R

@@ -1,0 +1,39 @@
+
+#Data from Bureau of Economic Analysis - https://www.bea.gov/data/gdp/gdp-county-metro-and-other-areas - table CAGDP1
+
+library(plm)
+library(plyr)
+library(ggplot2)
+library(tidyverse)
+library(plotly)
+library(bea.R)
+
+
+beakey = '386344DA-444A-4E59-A4FC-4DA15935C1A3'
+
+beaspecs <- list(
+  'UserID' = beakey,
+  'Method' = 'GetData',
+  'DatasetName' = 'Regional',
+  'TableName' = 'CAGDP1',
+  'LineCode' = '1',
+  'GeoFips' = '26101',
+  'Year' = 'ALL',
+  'ResultFormat' = 'json'
+)
+
+beagdp <- beaGet(beaspecs, asWide = FALSE)
+
+
+Manistee_GDP_plot <- ggplot(data=beagdp, aes(x = TimePeriod, y = DataValue, group= 1)) +
+  ggtitle("GDP of Manistee County, Michigan") +
+  geom_line() +
+  geom_point() +
+  labs(
+    x = "Year",
+    y = "Thousands of Chained 2012 Dollars"
+  )
+
+Manistee_GDP_plot
+
+ggplotly(Manistee_GDP_plot)

### 2020-11-04 EPA:Census Team/Gretchen Walkability/Walkability Data Manip.Rmd
Updated Walkability Plots

--- 2020-10-29 EPA:Census Team/Gretchen Walkability/Walkability Map.Rmd
@@
 library(ggplot2)
+library(plotly)
 
@@
 mwalk <- walk[walk$SFIPS == "26", ]
+#writeOGR(mwalk, ".", "mwalk", driver="ESRI Shapefile")
+mwalk2 <- readOGR(dsn = "mwalk.shp")
 
 statefips <- read.csv("~/Documents/Columbia/Practicum - Aracelis Torres/Data/Walkability/State FIPS.csv")
+countyfips <- read.csv("FIPS and County Names.csv")
 ```
@@
 # Walkability map (Michigan Specific) 
-## Target End state: blue water, data for whole country, better names, search by zip code/location, round Walkability Index Score, add layers for parks/schools/nature???
+## Target End state: blue water, better names, search by zip code/location, round Walkability Index Score, add points/layers for parks/schools/nature
 walkmap <- leaflet(mwalk) %>%
@@
 
+# plot Walkability by County
+# just Michigan 
+walkmi <- walkwithstates %>% filter(`State Code` == 26)
+walkmi$countyfips <- paste(walkmi$`State Code`, walkmi$CFIPS, sep = "")
+countyfips$X1001 <- as.character(countyfips$X1001)
+walkmi <- left_join(walkmi, countyfips, by = c("countyfips" = "X1001"))
+walkmi <- walkmi %>% group_by(Autauga) %>% 
+              summarize("Walkability" = mean(`NatWalkInd`))
+colnames(walkmi) <- c("County", "Walkability")
+#write_csv(walkmi, "walkmi.csv")
+
+#Plot it 
+#Reorder Walkability in table to make chart pretty
+walkmi <- transform(walkmi, `County` = reorder(`County`, -Walkability))
+#plot it 
+ plotc <- ggplot(walkmi, aes(x=County, y=Walkability)) + 
+  geom_bar(stat="identity", width=.8, fill="tomato3") + 
+  coord_flip() +
+  labs(title="Walkability by County",
+       caption="source: EPA") + 
+  theme(axis.text.y=element_text(size= 5))
+ggplotly(plotc)
+
 # lollipop Test: is it better? 

### 2020-11-04 EPA:Census Team/Gretchen Walkability/Walkability Plots.Rmd
Updated Walkability Plots

---
title: "Walkability Vizs"
author: "Gretchen Streett"
date: "11/3/2020"
output: html_document
---

```{r setup, }
library(leaflet)
library(rgdal)
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(stringr)
library(ggplot2)
library(plotly)

# All data is in this folder
mwalk <- readOGR(dsn = "mwalk.shp")
statefips <- read.csv("State FIPS.csv")
combined <- read.csv("combined.csv")
walkmi <- read.csv("walkmi.csv")
```

```{r Walkability Map: Michigan}
# Walkability map (Michigan Specific) 
## Target End state: blue water, better names, search by zip code/location, round Walkability Index Score, add points/layers for parks/schools/nature
walkmap <- leaflet(mwalk) %>%
 setView(lat = 44.2444, lng = -86.3243, zoom = 6) %>%
 addProviderTiles(providers$Stamen.TonerLite) %>%
  #add bloc borders
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333', opacity=1, 
  #chloropleth shading
  fillColor = ~colorQuantile("Greens", NatWalkInd)(NatWalkInd), 
  fillOpacity = 1, 
  #add pinpoint label
  label = ~stringr::str_c(CBSA_Name, ' ',
          formatC(NatWalkInd, big.mark = ',', format='f')),
  labelOptions = labelOptions(direction = 'auto'),
  #add hover highlight
  highlightOptions = highlightOptions(
          color='#000000', weight = 3,
          bringToFront = TRUE, sendToBack = TRUE))

walkmap2 <- walkmap %>% 
  addMiniMap() %>%
  addLegend(
  "bottomleft", 
  pal = colorQuantile("Greens", mwalk$NatWalkInd, n=5),
  values = ~NatWalkInd,
  title = htmltools::HTML("Walkability Index"),
opacity = 1, 
  labFormat = function(type, cuts, p) {
n = length(cuts)
paste0(cuts[-n], " – ", cuts[-1])}
)

walkmap3 <- walkmap2 %>%
    addEasyButton(easyButton(
    icon="fa-globe", 
    title="Zoom to Entire US Map",
    onClick=JS("function(btn, map){ map.setZoom(2.5); }"))) 
walkmap3
```

```{r Walkability Comparison: Manistee to State}
# Next steps: re-define "comparable rural areas"
myplot <- ggplot(data = combined, aes(x=`Location`, y=`Walkability`, fill=`Location`)) + 
  geom_bar(stat="identity", width=0.5) +
  scale_fill_brewer(palette="Dark2") +
  scale_x_discrete(limits=c("Manistee County", "Comparable Rural Area", "Entire State")) +
  geom_text(aes(label= round(`Walkability`, 2)), vjust=1.5, color="white", size=4)+
  labs(title="Walkability by Location")+
  theme_minimal() + 
  theme(legend.position="none")
myplot
```

```{r Walkability Comparison: All Michigan Counties}
# Next steps: Make font more readable, fix gap between bars and label on y axis

#Reorder Walkability in table to make chart pretty
walkmi <- transform(walkmi, `County` = reorder(`County`, -Walkability))
#plot it 
 plotc <- ggplot(walkmi, aes(x=County, y=Walkability)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  coord_flip() +
  labs(title="Walkability by County",
       caption="source: EPA") + 
  theme(axis.text.y=element_text(size= 5))
ggplotly(plotc)
```

### 2020-11-04 EPA:Census Team/Manistee Bband Access Map/Data Manipulation/Broadband Data Manip.Rmd
Add Broadband Access Map for Manistee County

---
title: "Broadband Rough Edit"
author: "Gretchen Streett"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgdal)
library(tidyverse)
library(stringr)
library(leaflet)
library(RColorBrewer)

# Get Data: MI Shapefiles, MI Broadband
mwalk2 <- readOGR(dsn = "/Users/gretchenilsestreett/Documents/Columbia/Practicum - Aracelis Torres/Data/Walkability/mwalk.shp")
broadband <- read_csv("FCC MI Broadband.csv")
broadband$BlockCode <- as.character(broadband$BlockCode)
maniblocks <- readOGR(dsn = "/Users/gretchenilsestreett/Documents/Columbia/Practicum - Aracelis Torres/Data/MI Broadband/Manistee Census Blocks/Manistee County Blocks/tl_2010_26101_tabblock10.shp")

# get # of Providers per block 
mani_bband <- left_join(maniblocks@data, broadband, by = c("GEOID10" = "BlockCode"))
mani_bband <- mani_bband[, c("GEOID10", "ProviderName")]
mani_bband <- mani_bband %>% group_by(GEOID10) %>% summarize("ProvideNum" = n())
maniblocks@data <- full_join(mani_bband, maniblocks@data, by = c("GEOID10" = "GEOID10"))
#writeOGR(maniblocks, ".", "manistee bband blocks", driver="ESRI Shapefile")

#plot
#define bins 
bins <- c(1:9)
pal <- colorNumeric("YlGnBu", domain = maniblocks@data$ProvideNum)

bbandmap <- leaflet(maniblocks) %>%
 addPolygons(stroke = TRUE, smoothFactor = 0.5,
  weight=1, color='#333333',  
  fillColor = ~pal(ProvideNum), 
  fillOpacity = 1,
  #add pinpoint label
  label = ~stringr::str_c(NAME10, ':',
          formatC(ProvideNum)),
  labelOptions = labelOptions(direction = 'auto'),
  #add hover highlight
  highlightOptions = highlightOptions(
          color='#000000', weight = 3,
          bringToFront = TRUE, sendToBack = TRUE))

bbandmap2 <- bbandmap %>% 
  addLegend("bottomleft", 
  pal = pal, 
  values = ~ProvideNum, 
  title = htmltools::HTML("Number of Providers"),
  setView(bbandmap2, lng = -86.3242, lat = 44.2444, zoom = 14))
bbandmap2
```

### 2020-11-05 EPA:Census Team/Kyung_Business_Establishments_Employeement/business_demo_prep.Rmd
Kyung Business Demo Prep

---
title: "Demo Prep"
author: "Kyung Suk Lee"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
---

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(plotly)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour="viridis",
  ggplot2.continuous.fill="viridis")

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

knitr::opts_chunk$set(comment = NA,
                      message = FALSE,
                      warning = FALSE,
                      echo = FALSE)

setwd("/Users/kxl50/Desktop/Practicum/Kyung_Business_Establishments_Employeement")

Sys.getenv("CENSUS_API_KEY")
readRenviron("~/.Renviron")

county_2015_2019 = 
  read_csv("./data/county_2015_2019.csv")
```

```{r}
est_ind = 
  county_2015_2019 %>% 
  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
  subset(!str_detect(area, "Unknown Or Undefined")) %>%
  filter(st_name == "Michigan") %>%
  separate(area, into = c("county", "state"), sep = -17) %>% 
  select(-state) %>%
  mutate(year = as.numeric(year)) %>% 
  filter(county == "Manistee") %>% 
  group_by(year, industry) %>% 
  ggplot(aes(x = year,
             y = annual_average_establishment_count,
             fill = industry,
             text = paste("Industry: ", industry,
                          "\nEstablishment: ", annual_average_establishment_count)
             )) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Establishments by Industry in Manistee (2015-2019)",
       x = "Year",
       y = "Number of Establishments")

ggplotly(est_ind, tooltip = "text")
```

```{r}
quotient = 
  county_2015_2019 %>% 
  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
  subset(!str_detect(area, "Unknown Or Undefined")) %>%
  filter(st_name == "Michigan") %>%
  separate(area, into = c("county", "state"), sep = -17) %>% 
  select(-state) %>%
  mutate(year = as.numeric(year)) %>% 
  filter(county == "Manistee") %>% 
  group_by(year, industry) %>% 
  ggplot(aes(x = year,
             y = employment_location_quotient_relative_to_u_s,
             fill = industry,
             text = paste("Industry: ", industry,
                          "\nLocation Quotient: ", employment_location_quotient_relative_to_u_s)
             )) +
  geom_bar(stat = "identity", color = "black", position = "dodge") +
  labs(title = "Location Quotient by Industry in Manistee (2015-2019)",
       x = "Year",
       y = "Number of Establishments")

ggplotly(quotient, tooltip = "text")
```

```{r shape_file}
mi_shp =
  get_acs(state = "MI", 
          geography = "county",
          variables = "B19013_001",
          year = 2018,
          geometry = TRUE) %>%
  janitor::clean_names() %>%
  select(geoid, name, geometry) %>%
  separate(name, into = c("county", "state"), sep = -17) %>%
  select(-state)
```

```{r}
payroll_state = 
county_2015_2019 %>% 
  filter(st_name == "Michigan" & year == "2019") %>% 
  separate(area, into = c("county", "state"), sep = -17) %>% 
  select(-state) %>% 
  filter(industry != "Total, all industries") %>% 
  group_by(county) %>% 
  summarize(wage = mean(annual_total_wages/annual_average_employment, na.rm = T))

payroll_plotly = 
  left_join(mi_shp, payroll_state, by = "county") %>%
  ggplot(aes(fill = wage, 
             text = paste("County: ", county, "\nWage: $", round(wage))
             )) +
  geom_sf(color = "black") +
  labs(title = "Employee's Annual Wage in Michigan (2019)",
       x = "Longitude",
       y = "Latitude")

ggplotly(payroll_plotly, tooltip = "text")
```

```{r}
diff_plotly =
  county_2015_2019 %>% 
  filter(st_name == "Michigan") %>% 
  separate(area, into = c("county", "state"), sep = -17) %>% 
  select(-state) %>% 
  filter(industry != "Total, all industries" &
           industry != "Unclassified" &
           county == "Manistee") %>% 
  mutate(lag_annual_average_employment = Hmisc::Lag(annual_average_employment, +12),
         employment_diff = lag_annual_average_employment - annual_average_employment) %>% 
  filter(year != "2019") %>% 
  ggplot(aes(x = employment_diff,
             y = industry, 
             fill = industry,
             text = paste("Industry: ", industry, 
                          "\nEmployment Diff: ", employment_diff)
             )) +
  geom_bar(stat = "identity") +
  facet_grid(.~year) + 
  coord_flip() +
  theme(axis.text.x = element_blank()) +
  labs(title = "Net Change in Employeement by Industry in Manistee (2015-2019)",
       x = "Net Change",
       y = "")

ggplotly(diff_plotly, tooltip = "text")
```

```{r}
employee_state =
  county_2015_2019 %>% 
  filter(st_name == "Michigan" & year == "2019") %>% 
  separate(area, into = c("county", "state"), sep = -17) %>% 
  select(-state) %>% 
  filter(industry != "Total, all industries") %>% 
  group_by(county) %>% 
  summarize(employment = sum(annual_average_employment, na.rm = T))

employee_plotly = 
  left_join(mi_shp, employee_state, by = "county") %>%
  ggplot(aes(fill = employment, 
             text = paste("County: ", county, "\nEmployee: ", employment)
             )) +
  geom_sf(color = "black") +
  labs(title = "Total Employment in Michigan (2019)",
       x = "Longitude",
       y = "Latitude")

ggplotly(employee_plotly, tooltip = "text")
```

### 2020-11-06 EPA:Census Team/Alison_Family_Household_Childcare.Rmd
Moved files

```{r Setup, include = FALSE, results = 'hide', warning = FALSE}
library(knitr)
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(plotly)

Sys.getenv("CENSUS_API_KEY")
```

Table B11001: Household Type: https://censusreporter.org/tables/B11001/
Table B09005: Household Type for Children in Households: https://censusreporter.org/tables/B09005/
Table B10001: Grandchildren Living With a Grandparent Householder by Age of Grandchild: https://censusreporter.org/tables/B10001/
Table B10002: Grandchildren Living With a Grandparent Householder by Grandparent Responsibility and Presence of Parent: https://censusreporter.org/tables/B10002/

```{r}
B11001_vars = c(total = "B11001_001E", 
                t_family = "B11001_002E", 
                tf_married_couple = "B11001_003E",
                tf_other = "B11001_004E",
                tfo_male = "B11001_005E",
                tfo_female = "B11001_006E",
                t_nonfamily = "B11001_007E",
                tn_alone = "B11001_008E",
                tn_not_alone = "B11001_009E")
```

```{r}
B11001_label = B11001_vars %>%
  enframe() %>%
  select(name) %>%
  unlist()
```

```{r}
B11001 <- get_acs(geography = "county",
                  variables = B11001_vars,
                  state = "26",
                  county = "101",
                  survey = "acs5",
                  year = 2018)
```

```{r}
household_type <- B11001 %>%
  mutate(label = B11001_label) %>%
  mutate(estimate_scaled = (estimate/9591)) %>%
  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
  slice(3, 4, 8, 9) %>%
  mutate(label = case_when(label == "tf_married_couple" ~ "Lives with spouse",
                           label == "tf_other" ~ "Lives with other family",
                           label == "tn_alone" ~ "Lives alone",
                           label == "tn_not_alone" ~ "Lives with non-family")) %>%
  mutate(label = factor(label, levels = c("Lives with spouse", "Lives with other family", "Lives with non-family", "Lives alone")))
```

```{r}
p1 <- ggplot(data = household_type, aes(x = label, y = estimate_scaled)) + 
  geom_bar(stat = "identity")
```

```{r}
ggplotly(p1)
```

```{r}
B09005_vars = c(total = "B09005_001E", 
                t_family = "B09005_002E", 
                tf_married_couple = "B09005_003E",
                tf_male = "B09005_004E",
                tf_female = "B09005_005E",
                t_nonfamily = "B09005_006E")
```

```{r}
B09005_label = B09005_vars %>%
  enframe() %>%
  select(name) %>%
  unlist()
```

```{r}
B09005 <- get_acs(geography = "county",
                  variables = B09005_vars,
                  state = "26",
                  county = "101",
                  survey = "acs5",
                  year = 2018)
```

```{r}
household_type_children <- B09005 %>%
  mutate(label = B09005_label) %>%
  mutate(estimate_scaled = (estimate/4298)) %>%
  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
  slice(3:6) %>%
  mutate(label = case_when(label == "tf_married_couple" ~ "Lives with married couple",
                           label == "tf_male" ~ "Lives with male householder",
                           label == "tf_female" ~ "Lives with female householder",
                           label == "t_nonfamily" ~ "Lives with non-family")) %>%
  mutate(label = factor(label, levels = c("Lives with married couple", "Lives with male householder", "Lives with female householder", 
                                          "Lives with non-family")))
```

```{r}
p2 <- ggplot(data = household_type_children, aes(x = label, y = estimate_scaled)) + 
  geom_bar(stat = "identity")
```

```{r}
ggplotly(p2)
```

```{r}
B10001_vars = c(total = "B10001_001E", 
                t_under_6 = "B10001_002E", 
                t_6_to_11 = "B10001_003E",
                t_12_to_17 = "B10001_004E")
```

```{r}
B10001_label = B10001_vars %>%
  enframe() %>%
  select(name) %>%
  unlist()
```

```{r}
B10001 <- get_acs(geography = "county",
                  variables = B10001_vars,
                  state = "26",
                  county = "101",
                  survey = "acs5",
                  year = 2018)
```

```{r}
grandchildren <- B10001 %>%
  mutate(label = B10001_label) %>%
  mutate(estimate_scaled = (estimate/327)) %>%
  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
  slice(2:4) %>%
  mutate(label = case_when(label == "t_under_6" ~ "Under 6 years old",
                           label == "t_6_to_11" ~ "6 to 11 years old",
                           label == "t_12_to_17" ~ "12 to 17 years old")) %>%
  mutate(label = factor(label, levels = c("Under 6 years old", "6 to 11 years old", "12 to 17 years old")))
```

```{r}
p3 <- ggplot(data = grandchildren, aes(x = label, y = estimate_scaled)) + 
  geom_bar(stat = "identity")
```

```{r}
ggplotly(p3)
```

```{r}
B10002_vars = c(total = "B10002_001E", 
                t_grandparent_responsible = "B10002_002E", 
                tgr_parent = "B10002_003E",
                tgr_no_parent = "B10002_004E",
                t_grandparent_not_responsible = "B10002_005E")
```

```{r}
B10002_label = B10002_vars %>%
  enframe() %>%
  select(name) %>%
  unlist()
```

```{r}
B10002 <- get_acs(geography = "county",
                  variables = B10002_vars,
                  state = "26",
                  county = "101",
                  survey = "acs5",
                  year = 2018)
```

```{r}
grandparents <- B10002 %>%
  mutate(label = B10002_label) %>%
  mutate(estimate_scaled = (estimate/327)) %>%
  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
  slice(3:5) %>%
  mutate(label = case_when(label == "tgr_parent" ~ "Parent present",
                           label == "tgr_no_parent" ~ "Parent not present",
                           label == "t_grandparent_not_responsible" ~ "Grandparent not responsible")) %>%
  mutate(label = factor(label, levels = c("Parent present", "Parent not present", "Grandparent not responsible")))
```

```{r}
p4 <- ggplot(data = grandparents, aes(x = label, y = estimate_scaled)) + 
  geom_bar(stat = "identity")
```

```{r}
ggplotly(p4)
```

### 2020-11-06 EPA:Census Team/app.R
added dashboard app.R
added dashboard app.R

@@ -1,0 +1,59 @@
+library(shiny)
+library(shinydashboard)
+library(tidyverse)
+
+source("housing.r")
+source("cost_income.R")
+
+# Define UI for application that draws a histogram
+ui <- dashboardPage(
+  dashboardHeader(title = "Columbia TOP Tool"),
+  dashboardSidebar(
+    sidebarMenu(
+      menuItem("Planner",
+               tabName = "planner_tab",
+               icon = icon("dashboard")),
+      menuItem("Developer",
+               tabName = "developer_tab",
+               icon = icon("dashboard")),
+      menuItem("Prospective Resident",
+               tabName = "prospective_resident_tab",
+               icon = icon("dashboard"))
+    )
+  ),
+  dashboardBody(
+    fluidRow(
+      tabItems(
+        tabItem(tabName = "planner_tab",
+                box(textInput(inputId = "location", label = "Location", value = NULL, placeholder = "Enter location"))),
+        tabItem(tabName = "developer_tab"),
+        tabItem(tabName = "prospective_resident_tab")
+      )
+    ),
+    fluidRow( 
+      box(
+        title = "Rent",
+        solidHeader = TRUE,
+        collapsible = TRUE,
+        plotOutput("rent", height = "300px")
+      ),
+      box(
+        title = "Income/Rent",
+        solidHeader = TRUE,
+        collapsible = TRUE,
+        plotlyOutput("income_rent", height = "300px")
+      )
+    )
+    
+  )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+  output$rent<- renderPlot({income_perc_plot})
+  output$income_rent<- renderPlotly({plotly_income_rent})
+}
+
+# Run the application 
+shinyApp(ui = ui, server = server)
+

### 2020-11-06 EPA:Census Team/Michelle Commute Housing/visualizations.r
Create visualizations.r

@@ -1,0 +1,161 @@
+## Household median income ##
+# Michigan household data on county level from ACS 5-year
+mi_hhincome <- get_acs(geography = "county", 
+                       variables = c(hhincome = "B19013_001"), 
+                       survey = "acs5", 
+                       state = "MI")
+
+# Rename to remove "County, Michigan"
+mi_hhincome2 <- mi_hhincome %>%  
+  mutate(NAME = str_replace(NAME, " County, Michigan", ""))
+
+# Dot plot
+theme_set(theme_bw())
+
+ggplot(mi_hhincome2, 
+       aes(x=estimate, 
+           y= reorder(NAME, estimate))) + 
+  geom_point(size = 2) + 
+  scale_x_continuous(labels = scales::dollar) + 
+  labs(
+    x = "Median Household Income", 
+    y = "", 
+    title = "Michigan Median Household Income by County") +
+  gghighlight(GEOID == 26101, label_key = NAME) +  # highlight County point 
+  theme(axis.text.y = element_text(size = 5, vjust=0.6))
+
+
+
+## Mode of transport for commute ##
+#Variables for Work Commute Transportation
+transportwork_vars <- c("Drove alone" = "B08006_003",
+                        "Carpooled" = "B08006_004",
+                        "Public transportation" = "B08006_008",
+                        "Bicycle" = "B08006_014",
+                        "Walked" = "B08006_015",
+                        "Other means" = "B08006_016",
+                        "Worked at home" = "B08006_017")
+
+#Pull data for Work Commute Transportation for Michigan
+mi_transportwork <- get_acs(geography = "county", 
+                            state = "MI",
+                            variables = transportwork_vars,
+                            summary_var = "B08006_001")
+                            
+#Calculate as percentages
+mi_transportwork_pct <- mi_transportwork %>%
+  mutate(pct = (estimate/ summary_est)) %>%
+  select(GEOID, NAME, variable, pct)
+
+#Subset for Manistee
+manistee_transportwork_pct <- subset(mi_transportwork_pct, GEOID==26101)
+manistee_transportwork_pct
+
+#Bar graph Manistee data
+ggplot(manistee_transportwork_pct, aes(x = reorder(variable, pct), y = pct, label = scales::percent(pct))) +
+  geom_col() + 
+  geom_text(position = position_dodge(width = .9), #move to center of bars
+            vjust = -0.5, #nudge above top of bar
[…]

### 2020-11-06 EPA:Census Team/Asahi_housing/cost_income.R
include all viz in app
include all viz in app

@@ -16,7 +16,7 @@ readRenviron("~/.Renviron")
 
 vars_acs5_2018 <- load_variables(year = 2018, dataset = "acs5")
 
-fair_rent <- read_xlsx("FY2018_50_County_rev.xlsx")
+fair_rent <- read_xlsx("Asahi_housing/data_files/FY2018_50_County_rev.xlsx")
 MI_fair_rent <- fair_rent[which(fair_rent$state_alpha == "MI"),]
 MI_fair_rent$GEOID <- substr(MI_fair_rent$fips2010, 0, 5)
 MI_fair_rent_trim <- MI_fair_rent[ , -which(names(MI_fair_rent) %in% c("state","cbsasub18","areaname18","county","cousub","cntyname","name","pop2010","hu2010","state_alpha"

### 2020-11-06 EPA:Census Team/app.R

@@ -1,47 +1,196 @@
+# 1 load libraries ----
 library(shiny)
 library(shinydashboard)
 library(tidyverse)
 
-source("housing.r")
-source("cost_income.R")
+# 2 run scripts ----
+# * 2.1 Housing (Asahi) --------------------------------------------------------
+source("Asahi_housing/housing.r")
+source("Asahi_housing/cost_income.R")
+# * 2.2 Sociodemographic (Alison) --------------------------------------------------------
+source("viz_scripts/alison_sociodem.R")
+# * 2.3 Demographics (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_demo.R")
+# * 2.4 Walkability (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_walk.R")
+# * 2.5 Business/Industry (Kyung) --------------------------------------------------------
+source("viz_scripts/kyung_business.R")
+# * 2.6 Broadband (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_bband.R")
+# * 2.7 Households (Alison) --------------------------------------------------------
+source("viz_scripts/alison_childcare.R")
+# * 2.8 Education + Earnings (Alisha) --------------------------------------------------------
+source("viz_scripts/alisha_earn.R")
 
 # Define UI for application that draws a histogram
 ui <- dashboardPage(
   dashboardHeader(title = "Columbia TOP Tool"),
   dashboardSidebar(
     sidebarMenu(
-      menuItem("Planner",
-               tabName = "planner_tab",
+      menuItem("Demographics",
+               tabName = "demo_tab",
                icon = icon("dashboard")),
-      menuItem("Developer",
-               tabName = "developer_tab",
+      menuItem("Childcare",
+               tabName = "child_tab",
                icon = icon("dashboard")),
-      menuItem("Prospective Resident",
-               tabName = "prospective_resident_tab",
+      menuItem("Quality of Life",
+               tabName = "qol_tab",
+               icon = icon("dashboard")),
+      menuItem("Housing",
+               tabName = "housing_tab",
+               icon = icon("dashboard")),
+      menuItem("Industry",
+               tabName = "industry_tab",
+               icon = icon("dashboard")),
+      menuItem("Broadband",
+               tabName = "bband_tab",
                icon = icon("dashboard"))
     )
   ),
   dashboardBody(
     fluidRow(
-      tabItems(
-        tabItem(tabName = "planner_tab",
[…]

### 2020-11-06 EPA:Census Team/kyung_business.R

@@ -1,0 +1,152 @@
+
+library(tidyverse)
+library(tidycensus)
+library(readxl)
+library(plotly)
+
+theme_set(theme_minimal() + theme(legend.position = "bottom"))
+
+options(
+  ggplot2.continuous.colour="viridis",
+  ggplot2.continuous.fill="viridis")
+
+scale_colour_discrete = scale_color_viridis_d
+scale_fill_discrete = scale_fill_viridis_d
+
+knitr::opts_chunk$set(comment = NA,
+                      message = FALSE,
+                      warning = FALSE,
+                      echo = FALSE)
+
+
+Sys.getenv("CENSUS_API_KEY")
+readRenviron("~/.Renviron")
+
+county_2015_2019 = 
+  read_csv("data_files/county_2015_2019.csv")
+
+est_ind = 
+  county_2015_2019 %>% 
+  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
+  subset(!str_detect(area, "Unknown Or Undefined")) %>%
+  filter(st_name == "Michigan") %>%
+  separate(area, into = c("county", "state"), sep = -17) %>% 
+  select(-state) %>%
+  mutate(year = as.numeric(year)) %>% 
+  filter(county == "Manistee") %>% 
+  group_by(year, industry) %>% 
+  ggplot(aes(x = year,
+             y = annual_average_establishment_count,
+             fill = industry,
+             text = paste("Industry: ", industry,
+                          "\nEstablishment: ", annual_average_establishment_count)
+  )) +
+  geom_bar(stat = "identity", color = "black") +
+  labs(title = "Establishments by Industry in Manistee (2015-2019)",
+       x = "Year",
+       y = "Number of Establishments")
+
+est_ind_plotly <- ggplotly(est_ind, tooltip = "text")
+
+quotient = 
+  county_2015_2019 %>% 
+  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
+  subset(!str_detect(area, "Unknown Or Undefined")) %>%
+  filter(st_name == "Michigan") %>%
+  separate(area, into = c("county", "state"), sep = -17) %>% 
+  select(-state) %>%
+  mutate(year = as.numeric(year)) %>% 
+  filter(county == "Manistee") %>% 
+  group_by(year, industry) %>% 
+  ggplot(aes(x = year,
+             y = employment_location_quotient_relative_to_u_s,
+             fill = industry,
+             text = paste("Industry: ", industry,
[…]

### 2020-11-06 EPA:Census Team/viz_scripts/alisha_educ.R

@@ -1,0 +1,70 @@
+library(tidycensus)
+library(tidyverse)
+library(ggplot2)
+library(plotly)
+library(dplyr)
+
+#Data from ACS5 2018 table B15003 - Education level totals for population over 25
+
+api_key <- Sys.getenv("CENSUS_API_KEY")
+census_api_key(api_key, overwrite = TRUE, install = TRUE)
+
+readRenviron("~/.Renviron")
+
+
+acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
+
+
+#pulling data for Manistee Education
+manistee_educ2 <- get_acs(geography = "county",
+                          state = "MI",
+                          county = "Manistee",
+                          table =  'B15003',
+                          year = 2018,
+                          output = "wide",
+                          geometry = TRUE
+)
+
+
+#EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER">
+#adding columns for different education level totals
+manistee_educ2$no_degree <- manistee_educ2$B15003_002E + manistee_educ2$B15003_003E + manistee_educ2$B15003_004E + manistee_educ2$B15003_005E + manistee_educ2$B15003_006E +  
+  manistee_educ2$B15003_007E + manistee_educ2$B15003_008E + manistee_educ2$B15003_009E + manistee_educ2$B15003_010E +
+  manistee_educ2$B15003_011E + manistee_educ2$B15003_012E + manistee_educ2$B15003_013E + manistee_educ2$B15003_014E +
+  manistee_educ2$B15003_015E + manistee_educ2$B15003_016E
+manistee_educ2$high_school <- manistee_educ2$B15003_017E + manistee_educ2$B15003_018E
+manistee_educ2$some_college <- manistee_educ2$B15003_019E + manistee_educ2$B15003_020E
+manistee_educ2$associates <- manistee_educ2$B15003_021E
+manistee_educ2$bachelors <- manistee_educ2$B15003_022E
+manistee_educ2$post_grad <- manistee_educ2$B15003_023E + manistee_educ2$B15003_024E + manistee_educ2$B15003_025E
+manistee_educ2$totalpop <- manistee_educ2$B15003_001E
+
+manistee_educ <- subset(manistee_educ2, select =  c(GEOID, NAME,no_degree, high_school, some_college, associates, bachelors, post_grad,totalpop))
+manistee_educ_pct<- manistee_educ %>% mutate(no_degree_pct = (no_degree/ totalpop),
[…]

### 2020-11-06 EPA:Census Team/viz_scripts/alison_childcare.R

@@ -1,0 +1,151 @@
+
+library(knitr)
+library(tidyverse)
+library(tidycensus)
+library(ggplot2)
+library(plotly)
+
+Sys.getenv("CENSUS_API_KEY")
+
+B11001_vars = c(total = "B11001_001E", 
+                t_family = "B11001_002E", 
+                tf_married_couple = "B11001_003E",
+                tf_other = "B11001_004E",
+                tfo_male = "B11001_005E",
+                tfo_female = "B11001_006E",
+                t_nonfamily = "B11001_007E",
+                tn_alone = "B11001_008E",
+                tn_not_alone = "B11001_009E")
+
+B11001_label = B11001_vars %>%
+  enframe() %>%
+  select(name) %>%
+  unlist()
+
+B11001 <- get_acs(geography = "county",
+                  variables = B11001_vars,
+                  state = "26",
+                  county = "101",
+                  survey = "acs5",
+                  year = 2018)
+
+household_type <- B11001 %>%
+  mutate(label = B11001_label) %>%
+  mutate(estimate_scaled = (estimate/9591)) %>%
+  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
+  slice(3, 4, 8, 9) %>%
+  mutate(label = case_when(label == "tf_married_couple" ~ "Lives with spouse",
+                           label == "tf_other" ~ "Lives with other family",
+                           label == "tn_alone" ~ "Lives alone",
+                           label == "tn_not_alone" ~ "Lives with non-family")) %>%
+  mutate(label = factor(label, levels = c("Lives with spouse", "Lives with other family", "Lives with non-family", "Lives alone")))
+
+household_plot <- ggplot(data = household_type, aes(x = label, y = estimate_scaled)) + 
+  geom_bar(stat = "identity")
+
+household_plotly <- ggplotly(household_plot)
+
+
+B09005_vars = c(total = "B09005_001E", 
+                t_family = "B09005_002E", 
+                tf_married_couple = "B09005_003E",
+                tf_male = "B09005_004E",
+                tf_female = "B09005_005E",
+                t_nonfamily = "B09005_006E")
+
+B09005_label = B09005_vars %>%
+  enframe() %>%
+  select(name) %>%
+  unlist()
+
+B09005 <- get_acs(geography = "county",
+                  variables = B09005_vars,
+                  state = "26",
+                  county = "101",
+                  survey = "acs5",
[…]

### 2020-11-06 EPA:Census Team/viz_scripts/alison_childcare.R
Alison updated files

@@ -1,11 +1,18 @@
 
-library(knitr)
 library(tidyverse)
 library(tidycensus)
 library(ggplot2)
 library(plotly)
 
 Sys.getenv("CENSUS_API_KEY")
+
+state = "26"
+county = "101"
+survey = "acs5"
+year = 2018
+
+
+# Table B11001: Household Type (Including Living Alone): https://censusreporter.org/tables/B11001/
 
 B11001_vars = c(total = "B11001_001E", 
                 t_family = "B11001_002E", 
@@ -24,10 +31,10 @@ B11001_label = B11001_vars %>%
 
 B11001 <- get_acs(geography = "county",
                   variables = B11001_vars,
-                  state = "26",
-                  county = "101",
-                  survey = "acs5",
-                  year = 2018)
+                  state = state,
+                  county = county,
+                  survey = survey,
+                  year = year)
 
 household_type <- B11001 %>%
   mutate(label = B11001_label) %>%
@@ -41,10 +48,14 @@ household_type <- B11001 %>%
   mutate(label = factor(label, levels = c("Lives with spouse", "Lives with other family", "Lives with non-family", "Lives alone")))
 
 household_plot <- ggplot(data = household_type, aes(x = label, y = estimate_scaled)) + 
-  geom_bar(stat = "identity")
+  geom_bar(stat = "identity") +
+  ggtitle("Household type") +
+  labs(caption = "Source: 2018 ACS 5-year survey")
 
 household_plotly <- ggplotly(household_plot)
 
+
+# Table B09005: Household Type for Children in Households: https://censusreporter.org/tables/B09005/
 
 B09005_vars = c(total = "B09005_001E", 
                 t_family = "B09005_002E", 
@@ -60,10 +71,10 @@ B09005_label = B09005_vars %>%
 
 B09005 <- get_acs(geography = "county",
                   variables = B09005_vars,
-                  state = "26",
-                  county = "101",
-                  survey = "acs5",
-                  year = 2018)
+                  state = state,
+                  county = county,
+                  survey = survey,
+                  year = year)
 
 household_type_children <- B09005 %>%
   mutate(label = B09005_label) %>%
@@ -78,11 +89,14 @@ household_type_children <- B09005 %>%
                                           "Lives with non-family")))
 
 children_plot <- ggplot(data = household_type_children, aes(x = label, y = estimate_scaled)) + 
-  geom_bar(stat = "identity")
[…]

### 2020-11-08 EPA:Census Team/Alisha/Manistee_Educ.R
.DS_Store banished!

@@ -1,70 +1,0 @@
-library(tidycensus)
-library(tidyverse)
-library(ggplot2)
-library(plotly)
-library(dplyr)
-
-#Data from ACS5 2018 table B15003 - Education level totals for population over 25
-
-api_key <- Sys.getenv("CENSUS_API_KEY")
-census_api_key(api_key, overwrite = TRUE, install = TRUE)
-
-readRenviron("~/.Renviron")
-
-
-acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
-
-
-#pulling data for Manistee Education
-manistee_educ2 <- get_acs(geography = "county",
-                         state = "MI",
-                         county = "Manistee",
-                         table =  'B15003',
-                         year = 2018,
-                         output = "wide",
-                         geometry = TRUE
-)
-
-
-#EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER">
-#adding columns for different education level totals
-manistee_educ2$no_degree <- manistee_educ2$B15003_002E + manistee_educ2$B15003_003E + manistee_educ2$B15003_004E + manistee_educ2$B15003_005E + manistee_educ2$B15003_006E +  
-                            manistee_educ2$B15003_007E + manistee_educ2$B15003_008E + manistee_educ2$B15003_009E + manistee_educ2$B15003_010E +
-                            manistee_educ2$B15003_011E + manistee_educ2$B15003_012E + manistee_educ2$B15003_013E + manistee_educ2$B15003_014E +
-                            manistee_educ2$B15003_015E + manistee_educ2$B15003_016E
-manistee_educ2$high_school <- manistee_educ2$B15003_017E + manistee_educ2$B15003_018E
-manistee_educ2$some_college <- manistee_educ2$B15003_019E + manistee_educ2$B15003_020E
-manistee_educ2$associates <- manistee_educ2$B15003_021E
-manistee_educ2$bachelors <- manistee_educ2$B15003_022E
-manistee_educ2$post_grad <- manistee_educ2$B15003_023E + manistee_educ2$B15003_024E + manistee_educ2$B15003_025E
-manistee_educ2$totalpop <- manistee_educ2$B15003_001E
-
-manistee_educ <- subset(manistee_educ2, select =  c(GEOID, NAME,no_degree, high_school, some_college, associates, bachelors, post_grad,totalpop))
-manistee_educ_pct<- manistee_educ %>% mutate(no_degree_pct = (no_degree/ totalpop),
[…]

### 2020-11-08 EPA:Census Team/Alisha/earn_industry_manistee_1.R

@@ -1,54 +1,0 @@
-library(plm)
-library(plyr)
-library(ggplot2)
-library(tidyverse)
-library(bea.R)
-#install.packages(httr)
-library(httr)
-
-beakey = '386344DA-444A-4E59-A4FC-4DA15935C1A3'
-
-paramvals <- beaParamVals(beakey, 'Regional', 'LineCode' )
-paramvals <- as.data.frame(paramvals)
-
-
-line1 = c( '400', '500', '600', '700', '1000', '1100', '1200', '1700','1800')
-
-
-beaearnindtotal<- data.frame()
-
-
-for (i in line1) {
-  beaspecs2 <- list(
-    'UserID' = beakey,
-    'Method' = 'GetData', 
-    'DatasetName' = 'Regional',
-    'TableName' = 'CAINC5N',
-    'LineCode' = i,
-    'GeoFips' = '26101',
-    'Year' = '2013,2014,2015,2016,2017,2018',
-    'ResultFormat' = 'json'
-  )
-  beaearningsind <- beaGet(beaspecs2, asWide = FALSE)
-  beaearnindtotal <- rbind(beaearnindtotal, beaearningsind, fill=TRUE)
-}
-
-by_year<- beaearnindtotal %>% group_by(TimePeriod)
-
-earn_plot<- ggplot(beaearnindtotal, aes(x = TimePeriod , y= DataValue, fill=factor(Code))) + ggtitle("Earnings Per Industry for Manistee County, Michigan") +
-  geom_bar(stat= "identity", position="dodge") +
-  scale_fill_discrete(name="Industry",
-                      labels=c("CAINC5N-400" = "Construction", 
-                               "CAINC5N-500" = "Manufacturing", "CAINC5N-600" = "Wholesale trade", "CAINC5N-700" = "Retail trade",
-                               "CAINC5N-1000" = "Finance and insurance",
-                               "CAINC5N-1100" = "Real estate and rental and leasing", "CAINC5N-1200" = "Professional, scientific, and technical services",
-                              "CAINC5N-1700" = "Arts, entertainment, and recreation", "CAINC5N-1800" = "Accommodation and food services")) +   
-  labs(
-    x = "Year",
-    y = "Earnings per Industry"
-  )
-
-earn_plot
-
-#ggplotly(earn_plot) <- commented this out because I'm not sure why but when I run this all of my legend labels dissapear
-

### 2020-11-08 EPA:Census Team/Asahi_housing/cost_income.R

@@ -1,59 +1,0 @@
-library(data.table)
-library(dplyr)
-library(extrafont)
-library(ggplot2)
-library(ggpubr)
-library(plotly)
-library(readxl)
-library(sf)
-library(tidycensus)
-
-loadfonts()
-
-api_key <- Sys.getenv("CENSUS_API_KEY")
-census_api_key(api_key, install = TRUE, overwrite = TRUE)
-readRenviron("~/.Renviron")
-
-vars_acs5_2018 <- load_variables(year = 2018, dataset = "acs5")
-
-fair_rent <- read_xlsx("Asahi_housing/data_files/FY2018_50_County_rev.xlsx")
-MI_fair_rent <- fair_rent[which(fair_rent$state_alpha == "MI"),]
-MI_fair_rent$GEOID <- substr(MI_fair_rent$fips2010, 0, 5)
-MI_fair_rent_trim <- MI_fair_rent[ , -which(names(MI_fair_rent) %in% c("state","cbsasub18","areaname18","county","cousub","cntyname","name","pop2010","hu2010","state_alpha"
-))]
-
-MI_income_county <- get_acs(geography = "county",
-                            state = "MI",
-                            variables = "B19001_001",
-                            year = 2018,
-                            output = "wide",
-                            geometry = FALSE)
-
-MI_county_geo <- get_acs(geography = "county",
-                         state = "MI",
-                         variables = "B19001_001",
-                         year = 2017,
-                         output = "wide",
-                         geometry = TRUE)[,c(1,5)]
-
-MI_rent_income <- merge(MI_fair_rent_trim,MI_income_county,by="GEOID")
-MI_rent_income$income_rent_ratio <- MI_rent_income$B19001_001E/(12*MI_rent_income$rent50_2)
-
-MI_rent_income_geo <- st_as_sf(merge(MI_rent_income,MI_county_geo,by="GEOID"))
-
-income_rent_plot <- ggplot(data = MI_rent_income_geo) + 
-  ggtitle("Income/Rent Ratio in Michigan (2018 ACS5)") +
-  labs(fill = "Income/Rent Ratio") +
-  geom_sf(color = "black", aes(fill= income_rent_ratio, text = NAME)) +
-  theme(
-    rect = element_rect(fill = "transparent"),
-    panel.background = element_rect(fill = NA),
-    plot.background = element_rect(fill = NA, color = NA),
-    panel.grid = element_blank(),
[…]

### 2020-11-08 EPA:Census Team/Asahi_housing/housing.R

@@ -1,283 +1,0 @@
-library(data.table)
-library(dplyr)
-library(extrafont)
-library(ggplot2)
-library(ggpubr)
-library(sf)
-library(tidycensus)
-
-loadfonts()
-
-api_key <- Sys.getenv("CENSUS_API_KEY")
-census_api_key(api_key, install = TRUE, overwrite = TRUE)
-readRenviron("~/.Renviron")
-
-vars_acs5_2018 <- load_variables(year = 2018, dataset = "acs5")
-rent_vars <- vars_acs5_2018[vars_acs5_2018$name %like% "B25063", ][3:26,]
-rent_thresholds <- regmatches(rent_vars$label, gregexpr("[[:digit:]]+", rent_vars$label))
-
-rent_threshold_labels <- c()
-for (i in c(1:length(rent_thresholds))) {
-  length <- length(rent_thresholds[[i]])
-  label <- ""
-  if ((length) == 1) {
-    rent_threshold_labels[i] <- paste("less than $",rent_thresholds[[i]], sep = "")
-  }
-  else if ((length) == 2) {
-    if (as.numeric(rent_thresholds[[i]][1]) < 100) {
-      rent_threshold_labels[i] <- paste("greater than $",rent_thresholds[[i]][1],",",rent_thresholds[[i]][2], sep = "")
-    }
-    else {
-      rent_threshold_labels[i] <- paste("$",rent_thresholds[[i]][1]," to $",rent_thresholds[[i]][2], sep = "")
-    }
-  }
-  else {
-    rent_threshold_labels[i] <- paste("$",rent_thresholds[[i]][1],",",rent_thresholds[[i]][2]," to ","$",rent_thresholds[[i]][3],",",rent_thresholds[[i]][4], sep = "")
-  }
-}
-
-manistee_rent_tract <- get_acs(geography = "tract",
-                               state = "MI",
-                               county = "Manistee",
-                               variables = rent_vars$name,
-                               year = 2018,
-                               output = "wide",
-                               geometry = TRUE)
-manistee_rent_county <- get_acs(geography = "county",
-                                state = "MI",
-                                county = "Manistee",
-                                variables = rent_vars$name,
-                                year = 2018,
-                                output = "wide",
-                                geometry = TRUE)
-
-county_rents <- manistee_rent_county[ ,grepl("B.*E|E.*B" ,names(manistee_rent_county))]
-st_geometry(county_rents) <- NULL
-
-county_rents_DF <- data.frame(thresholds = rent_threshold_labels,
[…]

### 2020-11-08 EPA:Census Team/Michelle Commute Housing/visualizations.r

@@ -1,161 +1,0 @@
-## Household median income ##
-# Michigan household data on county level from ACS 5-year
-mi_hhincome <- get_acs(geography = "county", 
-                       variables = c(hhincome = "B19013_001"), 
-                       survey = "acs5", 
-                       state = "MI")
-
-# Rename to remove "County, Michigan"
-mi_hhincome2 <- mi_hhincome %>%  
-  mutate(NAME = str_replace(NAME, " County, Michigan", ""))
-
-# Dot plot
-theme_set(theme_bw())
-
-ggplot(mi_hhincome2, 
-       aes(x=estimate, 
-           y= reorder(NAME, estimate))) + 
-  geom_point(size = 2) + 
-  scale_x_continuous(labels = scales::dollar) + 
-  labs(
-    x = "Median Household Income", 
-    y = "", 
-    title = "Michigan Median Household Income by County") +
-  gghighlight(GEOID == 26101, label_key = NAME) +  # highlight County point 
-  theme(axis.text.y = element_text(size = 5, vjust=0.6))
-
-
-
-## Mode of transport for commute ##
-#Variables for Work Commute Transportation
-transportwork_vars <- c("Drove alone" = "B08006_003",
-                        "Carpooled" = "B08006_004",
-                        "Public transportation" = "B08006_008",
-                        "Bicycle" = "B08006_014",
-                        "Walked" = "B08006_015",
-                        "Other means" = "B08006_016",
-                        "Worked at home" = "B08006_017")
-
-#Pull data for Work Commute Transportation for Michigan
-mi_transportwork <- get_acs(geography = "county", 
-                            state = "MI",
-                            variables = transportwork_vars,
-                            summary_var = "B08006_001")
-                            
-#Calculate as percentages
-mi_transportwork_pct <- mi_transportwork %>%
-  mutate(pct = (estimate/ summary_est)) %>%
-  select(GEOID, NAME, variable, pct)
-
-#Subset for Manistee
-manistee_transportwork_pct <- subset(mi_transportwork_pct, GEOID==26101)
-manistee_transportwork_pct
-
-#Bar graph Manistee data
-ggplot(manistee_transportwork_pct, aes(x = reorder(variable, pct), y = pct, label = scales::percent(pct))) +
-  geom_col() + 
-  geom_text(position = position_dodge(width = .9), #move to center of bars
-            vjust = -0.5, #nudge above top of bar
[…]

### 2020-11-08 EPA:Census Team/app.R

@@ -1,237 +1,0 @@
-# 1 load libraries ----
-library(shiny)
-library(shinydashboard)
-library(tidyverse)
-
-# 2 run scripts ----
-# * 2.1 Housing (Asahi) --------------------------------------------------------
-source("Asahi_housing/housing.r")
-source("Asahi_housing/cost_income.R")
-# * 2.2 Sociodemographic (Alison) --------------------------------------------------------
-source("viz_scripts/alison_sociodem.R")
-# * 2.3 Demographics (Gretchen) --------------------------------------------------------
-source("viz_scripts/gretchen_demo.R")
-# * 2.4 Walkability (Gretchen) --------------------------------------------------------
-source("viz_scripts/gretchen_walk.R")
-# * 2.5 Business/Industry (Kyung) --------------------------------------------------------
-source("viz_scripts/kyung_business.R")
-# * 2.6 Broadband (Gretchen) --------------------------------------------------------
-source("viz_scripts/gretchen_bband.R")
-# * 2.7 Households (Alison) --------------------------------------------------------
-source("viz_scripts/alison_childcare.R")
-# * 2.8 Education + Earnings (Alisha) --------------------------------------------------------
-source("viz_scripts/alisha_earn.R")
-
-# Define UI for application that draws a histogram
-ui <- dashboardPage(
-  dashboardHeader(title = "Columbia TOP Tool"),
-  dashboardSidebar(
-    sidebarMenu(
-      menuItem("Demographics",
-               tabName = "demo_tab",
-               icon = icon("dashboard")),
-      menuItem("Childcare",
-               tabName = "child_tab",
-               icon = icon("dashboard")),
-      menuItem("Quality of Life",
-               tabName = "qol_tab",
-               icon = icon("dashboard")),
-      menuItem("Housing",
-               tabName = "housing_tab",
-               icon = icon("dashboard")),
-      menuItem("Industry",
-               tabName = "industry_tab",
-               icon = icon("dashboard")),
-      menuItem("Broadband",
-               tabName = "bband_tab",
-               icon = icon("dashboard"))
-    )
-  ),
-  dashboardBody(
-    fluidRow(
-        box(textInput(inputId = "location", label = "Location", value = NULL, placeholder = "Enter location")
-      )
-    ),
-    fluidRow(
-      tabItems(
-        tabItem(tabName = "demo_tab",
-                box(
-                  solidHeader = TRUE,
-                  collapsible = TRUE,
[…]

### 2020-11-08 EPA:Census Team/kyung_business.R

@@ -1,152 +1,0 @@
-
-library(tidyverse)
-library(tidycensus)
-library(readxl)
-library(plotly)
-
-theme_set(theme_minimal() + theme(legend.position = "bottom"))
-
-options(
-  ggplot2.continuous.colour="viridis",
-  ggplot2.continuous.fill="viridis")
-
-scale_colour_discrete = scale_color_viridis_d
-scale_fill_discrete = scale_fill_viridis_d
-
-knitr::opts_chunk$set(comment = NA,
-                      message = FALSE,
-                      warning = FALSE,
-                      echo = FALSE)
-
-
-Sys.getenv("CENSUS_API_KEY")
-readRenviron("~/.Renviron")
-
-county_2015_2019 = 
-  read_csv("data_files/county_2015_2019.csv")
-
-est_ind = 
-  county_2015_2019 %>% 
-  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
-  subset(!str_detect(area, "Unknown Or Undefined")) %>%
-  filter(st_name == "Michigan") %>%
-  separate(area, into = c("county", "state"), sep = -17) %>% 
-  select(-state) %>%
-  mutate(year = as.numeric(year)) %>% 
-  filter(county == "Manistee") %>% 
-  group_by(year, industry) %>% 
-  ggplot(aes(x = year,
-             y = annual_average_establishment_count,
-             fill = industry,
-             text = paste("Industry: ", industry,
-                          "\nEstablishment: ", annual_average_establishment_count)
-  )) +
-  geom_bar(stat = "identity", color = "black") +
-  labs(title = "Establishments by Industry in Manistee (2015-2019)",
-       x = "Year",
-       y = "Number of Establishments")
-
-est_ind_plotly <- ggplotly(est_ind, tooltip = "text")
-
-quotient = 
-  county_2015_2019 %>% 
-  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
-  subset(!str_detect(area, "Unknown Or Undefined")) %>%
-  filter(st_name == "Michigan") %>%
-  separate(area, into = c("county", "state"), sep = -17) %>% 
-  select(-state) %>%
-  mutate(year = as.numeric(year)) %>% 
-  filter(county == "Manistee") %>% 
-  group_by(year, industry) %>% 
-  ggplot(aes(x = year,
-             y = employment_location_quotient_relative_to_u_s,
-             fill = industry,
-             text = paste("Industry: ", industry,
[…]

### 2020-11-08 EPA:Census Team/viz_scripts/alisha_earn.R

@@ -1,54 +1,0 @@
-library(plm)
-library(plyr)
-library(ggplot2)
-library(tidyverse)
-library(bea.R)
-#install.packages(httr)
-library(httr)
-
-beakey = '386344DA-444A-4E59-A4FC-4DA15935C1A3'
-
-paramvals <- beaParamVals(beakey, 'Regional', 'LineCode' )
-paramvals <- as.data.frame(paramvals)
-
-
-line1 = c( '400', '500', '600', '700', '1000', '1100', '1200', '1700','1800')
-
-
-beaearnindtotal<- data.frame()
-
-
-for (i in line1) {
-  beaspecs2 <- list(
-    'UserID' = beakey,
-    'Method' = 'GetData', 
-    'DatasetName' = 'Regional',
-    'TableName' = 'CAINC5N',
-    'LineCode' = i,
-    'GeoFips' = '26101',
-    'Year' = '2013,2014,2015,2016,2017,2018',
-    'ResultFormat' = 'json'
-  )
-  beaearningsind <- beaGet(beaspecs2, asWide = FALSE)
-  beaearnindtotal <- rbind(beaearnindtotal, beaearningsind, fill=TRUE)
-}
-
-by_year<- beaearnindtotal %>% group_by(TimePeriod)
-
-earn_plot<- ggplot(beaearnindtotal, aes(x = TimePeriod , y= DataValue, fill=factor(Code))) + ggtitle("Earnings Per Industry for Manistee County, Michigan") +
-  geom_bar(stat= "identity", position="dodge") +
-  scale_fill_discrete(name="Industry",
-                      labels=c("CAINC5N-400" = "Construction", 
-                               "CAINC5N-500" = "Manufacturing", "CAINC5N-600" = "Wholesale trade", "CAINC5N-700" = "Retail trade",
-                               "CAINC5N-1000" = "Finance and insurance",
-                               "CAINC5N-1100" = "Real estate and rental and leasing", "CAINC5N-1200" = "Professional, scientific, and technical services",
-                               "CAINC5N-1700" = "Arts, entertainment, and recreation", "CAINC5N-1800" = "Accommodation and food services")) +   
-  labs(
-    x = "Year",
-    y = "Earnings per Industry"
-  )
-
-earn_plot
-
-#ggplotly(earn_plot) <- commented this out because I'm not sure why but when I run this all of my legend labels dissapear
-

### 2020-11-08 EPA:Census Team/viz_scripts/alisha_educ.R

@@ -1,70 +1,0 @@
-library(tidycensus)
-library(tidyverse)
-library(ggplot2)
-library(plotly)
-library(dplyr)
-
-#Data from ACS5 2018 table B15003 - Education level totals for population over 25
-
-api_key <- Sys.getenv("CENSUS_API_KEY")
-census_api_key(api_key, overwrite = TRUE, install = TRUE)
-
-readRenviron("~/.Renviron")
-
-
-acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
-
-
-#pulling data for Manistee Education
-manistee_educ2 <- get_acs(geography = "county",
-                          state = "MI",
-                          county = "Manistee",
-                          table =  'B15003',
-                          year = 2018,
-                          output = "wide",
-                          geometry = TRUE
-)
-
-
-#EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER">
-#adding columns for different education level totals
-manistee_educ2$no_degree <- manistee_educ2$B15003_002E + manistee_educ2$B15003_003E + manistee_educ2$B15003_004E + manistee_educ2$B15003_005E + manistee_educ2$B15003_006E +  
-  manistee_educ2$B15003_007E + manistee_educ2$B15003_008E + manistee_educ2$B15003_009E + manistee_educ2$B15003_010E +
-  manistee_educ2$B15003_011E + manistee_educ2$B15003_012E + manistee_educ2$B15003_013E + manistee_educ2$B15003_014E +
-  manistee_educ2$B15003_015E + manistee_educ2$B15003_016E
-manistee_educ2$high_school <- manistee_educ2$B15003_017E + manistee_educ2$B15003_018E
-manistee_educ2$some_college <- manistee_educ2$B15003_019E + manistee_educ2$B15003_020E
-manistee_educ2$associates <- manistee_educ2$B15003_021E
-manistee_educ2$bachelors <- manistee_educ2$B15003_022E
-manistee_educ2$post_grad <- manistee_educ2$B15003_023E + manistee_educ2$B15003_024E + manistee_educ2$B15003_025E
-manistee_educ2$totalpop <- manistee_educ2$B15003_001E
-
-manistee_educ <- subset(manistee_educ2, select =  c(GEOID, NAME,no_degree, high_school, some_college, associates, bachelors, post_grad,totalpop))
-manistee_educ_pct<- manistee_educ %>% mutate(no_degree_pct = (no_degree/ totalpop),
[…]

### 2020-11-08 EPA:Census Team/viz_scripts/alison_childcare.R

@@ -1,172 +1,0 @@
-
-library(tidyverse)
-library(tidycensus)
-library(ggplot2)
-library(plotly)
-
-Sys.getenv("CENSUS_API_KEY")
-
-state = "26"
-county = "101"
-survey = "acs5"
-year = 2018
-
-
-# Table B11001: Household Type (Including Living Alone): https://censusreporter.org/tables/B11001/
-
-B11001_vars = c(total = "B11001_001E", 
-                t_family = "B11001_002E", 
-                tf_married_couple = "B11001_003E",
-                tf_other = "B11001_004E",
-                tfo_male = "B11001_005E",
-                tfo_female = "B11001_006E",
-                t_nonfamily = "B11001_007E",
-                tn_alone = "B11001_008E",
-                tn_not_alone = "B11001_009E")
-
-B11001_label = B11001_vars %>%
-  enframe() %>%
-  select(name) %>%
-  unlist()
-
-B11001 <- get_acs(geography = "county",
-                  variables = B11001_vars,
-                  state = state,
-                  county = county,
-                  survey = survey,
-                  year = year)
-
-household_type <- B11001 %>%
-  mutate(label = B11001_label) %>%
-  mutate(estimate_scaled = (estimate/9591)) %>%
-  select(GEOID, NAME, variable, label, estimate_scaled, estimate, moe) %>%
-  slice(3, 4, 8, 9) %>%
-  mutate(label = case_when(label == "tf_married_couple" ~ "Lives with spouse",
-                           label == "tf_other" ~ "Lives with other family",
-                           label == "tn_alone" ~ "Lives alone",
-                           label == "tn_not_alone" ~ "Lives with non-family")) %>%
-  mutate(label = factor(label, levels = c("Lives with spouse", "Lives with other family", "Lives with non-family", "Lives alone")))
-
-household_plot <- ggplot(data = household_type, aes(x = label, y = estimate_scaled)) + 
-  geom_bar(stat = "identity") +
-  ggtitle("Household type") +
-  labs(caption = "Source: 2018 ACS 5-year survey")
-
-household_plotly <- ggplotly(household_plot)
-
-
-# Table B09005: Household Type for Children in Households: https://censusreporter.org/tables/B09005/
-
-B09005_vars = c(total = "B09005_001E", 
-                t_family = "B09005_002E", 
-                tf_married_couple = "B09005_003E",
-                tf_male = "B09005_004E",
-                tf_female = "B09005_005E",
[…]

### 2020-11-08 EPA:Census Team/viz_scripts/alison_sociodem.R

@@ -1,126 +1,0 @@
-
-library(knitr)
-library(tidyverse)
-library(tidycensus)
-library(ggplot2)
-library(ggthemes)
-library(jcolors)
-
-Sys.getenv("CENSUS_API_KEY")
-
-acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
-
-write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
-
-
-sex_age_vars <- acs5_2018_vars %>%
-  slice(5:51)
-
-sex_age_vec <- sex_age_vars %>%
-  select(name) %>%
-  unlist()
-
-sex_age_label <- sex_age_vars %>%
-  select(label) %>%
-  unlist()
-
-
-
-MI_place_sex_age_acs5_2018 <- get_acs(geography = "place",
-                                      variables = sex_age_vec,
-                                      place = "Manistee",
-                                      state = "MI",
-                                      year = 2018)
-
-Man_city_sex_age_acs5_2018 <- MI_place_sex_age_acs5_2018 %>%
-  slice(18425:18471) %>%
-  mutate(sex_age_label = sex_age_label) %>%
-  select(sex_age_label, estimate)
-
-
-
-# https://github.com/walkerke/tidycensus/issues/131
-
-years <- lst(2013, 2018)
-
-MI_place_mob_acs5_2013_2018 <- map_dfr(
-  years,
-  ~ get_acs(
-    geography = "place",
-    variables = c("B07001_016", "B07001_033", "B07001_049", "B07001_065", "B07001_081"),
-    state = "MI",
-    place = "Manistee",
-    year = .x
-  ),
-  .id = "year"
-)
-
-
-
-Man_city_mob_acs5_2013_2018 <- MI_place_mob_acs5_2013_2018 %>%
-  filter(NAME == "Manistee city, Michigan") %>%
-  add_column(label = "") %>%
-  mutate(label = case_when(variable == "B07001_016" ~ "Same house",
-                           variable == "B07001_033" ~ "Same county",
-                           variable == "B07001_049" ~ "Different county",
-                           variable == "B07001_065" ~ "Different state",
-                           variable == "B07001_081" ~ "Moved from abroad")) %>%
-  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
-
-
-
-p1_mobility <- ggplot(data = Man_city_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
[…]

### 2020-11-08 EPA:Census Team/viz_scripts/gretchen_bband.R

@@ -1,37 +1,0 @@
-
-library(rgdal)
-library(tidyverse)
-library(stringr)
-library(leaflet)
-library(RColorBrewer)
-
-# Get Data (in this folder)
-maniblocks <- readOGR(dsn = "data_files/manistee bband blocks.shp")
-
-#plot
-#define bins 
-bins <- c(1:9)
-pal <- colorNumeric("YlGnBu", domain = maniblocks@data$ProvideNum)
-
-bbandmap <- leaflet(maniblocks) %>%
-  addPolygons(stroke = TRUE, smoothFactor = 0.5,
-              weight=1, color='#333333',  
-              fillColor = ~pal(ProvideNum), 
-              fillOpacity = 1,
-              #add pinpoint label
-              label = ~stringr::str_c(NAME10, ':',
-                                      formatC(ProvideNum)),
-              labelOptions = labelOptions(direction = 'auto'),
-              #add hover highlight
-              highlightOptions = highlightOptions(
-                color='#000000', weight = 3,
-                bringToFront = TRUE, sendToBack = TRUE))
-
-bbandmap2 <- bbandmap %>% 
-  addLegend("bottomleft", 
-            pal = pal, 
-            values = ~ProvideNum, 
-            title = htmltools::HTML("Number of Providers"))
-
-bbandmap3 <- setView(bbandmap2, lng = -86.1, lat = 44.3350, zoom = 10)
-bbandmap3

### 2020-11-08 EPA:Census Team/viz_scripts/gretchen_demo.R

@@ -1,48 +1,0 @@
-# Data you need to plot is in this folder
-# See "Data Manipulation" folder for OG data & manipulation code" 
-
-library(tidyverse)
-library(ggplot2)
-library(scales)
-
-just_sex <- read.csv("data_files/just_sex")
-age_breakdown_manistee <- read.csv("data_files/age_breakdown_manistee")
-race_data_manistee <- read.csv("data_files/race_data_manistee")
-
-# create "Blank_theme" to remove all the axis markings for pie chart 
-blank_theme <- theme_minimal()+
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x=element_blank(),
-    axis.title.y = element_blank(),
-    panel.border = element_blank(),
-    panel.grid=element_blank(),
-    axis.ticks = element_blank(),
-    plot.title=element_text(size=14, face="bold")
-  )
-
-# Create Chart (in ggplot2 you have to make a bar chart first, then change "coordinates" to make it into a pie)
-genderbar<- ggplot(just_sex, aes(x="", y=`Manistee.County..MI`, fill=`label`))+
-  geom_bar(width = 1, stat = "identity")
-genderpie <- genderbar + coord_polar("y", start=0) +
-  geom_text(aes(y = `Manistee.County..MI`/3 + c(0, cumsum(`Manistee.County..MI`)[-length(`Manistee.County..MI`)]), 
-                label = percent(`Manistee.County..MI`/sum(`Manistee.County..MI`), accuracy=1)), size=5) + 
-  ggtitle(label = "Sex") +
-  scale_fill_discrete(name = "", labels = c("Female", "Male")) +
-  blank_theme
-genderpie 
-
-#plot
-agebar <- ggplot(age_breakdown_manistee, aes(x= `NewCat`, y=`Per`)) + 
-  geom_bar(width = .8, stat = "identity") + 
-  labs(title="Population by Age Range for Manistee County") + 
-  xlab("Age Range") + ylab("Percent of Population")
-agebar
-
-# Note: "Hispanic" is included in 2+ category and "hispanic" category, but not others 
-racebar <- ggplot(race_data_manistee, aes(x= `CompRace`, y=as.numeric(`Per` * 100))) + 
-  geom_bar(width = .8, stat = "identity") + 
-  scale_y_continuous(limits = c(0,90), expand = c(0, 0)) +
-  labs(title="Race & Ethnicity for Manistee County") + 
-  xlab("Race") + ylab("Percent of Population")
-racebar

### 2020-11-08 EPA:Census Team/viz_scripts/gretchen_walk.R

@@ -1,79 +1,0 @@
-library(leaflet)
-library(rgdal)
-library(tidyverse)
-library(dplyr)
-library(RColorBrewer)
-library(stringr)
-library(ggplot2)
-library(plotly)
-
-# All data is in this folder
-mwalk <- readOGR(dsn = "data_files/mwalk.shp")
-statefips <- read.csv("data_files/State FIPS.csv")
-combined <- read.csv("data_files/combined.csv")
-walkmi <- read.csv("data_files/walkmi.csv")
-
-# Walkability map (Michigan Specific) 
-## Target End state: blue water, better names, search by zip code/location, round Walkability Index Score, add points/layers for parks/schools/nature
-walkmap <- leaflet(mwalk) %>%
-  setView(lat = 44.2444, lng = -86.3243, zoom = 6) %>%
-  addProviderTiles(providers$Stamen.TonerLite) %>%
-  #add bloc borders
-  addPolygons(stroke = TRUE, smoothFactor = 0.5,
-              weight=1, color='#333333', opacity=1, 
-              #chloropleth shading
-              fillColor = ~colorQuantile("Greens", NatWalkInd)(NatWalkInd), 
-              fillOpacity = 1, 
-              #add pinpoint label
-              label = ~stringr::str_c(CBSA_Name, ' ',
-                                      formatC(NatWalkInd, big.mark = ',', format='f')),
-              labelOptions = labelOptions(direction = 'auto'),
-              #add hover highlight
-              highlightOptions = highlightOptions(
-                color='#000000', weight = 3,
-                bringToFront = TRUE, sendToBack = TRUE))
-
-walkmap2 <- walkmap %>% 
-  addMiniMap() %>%
-  addLegend(
-    "bottomleft", 
-    pal = colorQuantile("Greens", mwalk$NatWalkInd, n=5),
-    values = ~NatWalkInd,
-    title = htmltools::HTML("Walkability Index"),
-    opacity = 1, 
-    labFormat = function(type, cuts, p) {
-      n = length(cuts)
-      paste0(cuts[-n], " – ", cuts[-1])}
-  )
-
-walkmap3 <- walkmap2 %>%
-  addEasyButton(easyButton(
-    icon="fa-globe", 
-    title="Zoom to Entire US Map",
-    onClick=JS("function(btn, map){ map.setZoom(2.5); }"))) 
-walkmap3
-
-# Next steps: re-define "comparable rural areas"
[…]

### 2020-11-08 EPA:Census Team/viz_scripts/kyung_business.R

@@ -1,151 +1,0 @@
-
-library(tidyverse)
-library(tidycensus)
-library(readxl)
-library(plotly)
-
-theme_set(theme_minimal() + theme(legend.position = "bottom"))
-
-options(
-  ggplot2.continuous.colour="viridis",
-  ggplot2.continuous.fill="viridis")
-
-scale_colour_discrete = scale_color_viridis_d
-scale_fill_discrete = scale_fill_viridis_d
-
-knitr::opts_chunk$set(comment = NA,
-                      message = FALSE,
-                      warning = FALSE,
-                      echo = FALSE)
-
-Sys.getenv("CENSUS_API_KEY")
-readRenviron("~/.Renviron")
-
-county_2015_2019 = 
-  read_csv("data_files/county_2015_2019.csv")
-
-est_ind = 
-  county_2015_2019 %>% 
-  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
-  subset(!str_detect(area, "Unknown Or Undefined")) %>%
-  filter(st_name == "Michigan") %>%
-  separate(area, into = c("county", "state"), sep = -17) %>% 
-  select(-state) %>%
-  mutate(year = as.numeric(year)) %>% 
-  filter(county == "Manistee") %>% 
-  group_by(year, industry) %>% 
-  ggplot(aes(x = year,
-             y = annual_average_establishment_count,
-             fill = industry,
-             text = paste("Industry: ", industry,
-                          "\nEstablishment: ", annual_average_establishment_count)
-  )) +
-  geom_bar(stat = "identity", color = "black") +
-  labs(title = "Establishments by Industry in Manistee (2015-2019)",
-       x = "Year",
-       y = "Number of Establishments")
-
-est_ind_plotly <- ggplotly(est_ind, tooltip = "text")
-
-quotient = 
-  county_2015_2019 %>% 
-  filter(industry != "Unclassified" & industry != "Total, all industries") %>% 
-  subset(!str_detect(area, "Unknown Or Undefined")) %>%
-  filter(st_name == "Michigan") %>%
-  separate(area, into = c("county", "state"), sep = -17) %>% 
-  select(-state) %>%
-  mutate(year = as.numeric(year)) %>% 
-  filter(county == "Manistee") %>% 
-  group_by(year, industry) %>% 
-  ggplot(aes(x = year,
-             y = employment_location_quotient_relative_to_u_s,
-             fill = industry,
-             text = paste("Industry: ", industry,
[…]

### 2020-11-08 KPMG Team/KPMG Datasets & Models/SIR/src/SEIR_Covidactnow.py

@@ -1,181 +1,0 @@
-
-# coding: utf-8
-
-# In[9]:
-
-
-import pandas as pd
-import numpy as np
-from typing import Generator, Tuple
-from collections import namedtuple
-import matplotlib.pyplot as plt
-import copy
-
-
-# ### SEIR CovidActNow Model
-# 
-# mimic similar code structure as the streamlit app
-
-# In[10]:
-
-
-def seir_actnow(
-    S: float,  # Susceptible
-    E: float, # Exposed 
-    I_Mild: float, # Infected (mild, not hosp)
-    I_Hosp: float, # Infected (hosp)
-    I_ICU: float, # Infected (ICU)
-    R: float, # Recovered (from 3 types of Infected)
-    D: float, # Deceased (from I_C)
-    beta_1: float, # from S -> Infected (mild, not hosp)
-    beta_2: float, # from S -> Infected (hosp)
-    beta_3: float, # from S -> Infected (ICU)
-    alpha: float, # from E -> Infected (mild, not hosp)
-    gamma_1: float, # Infected (mild, not hosp)-> R
-    gamma_2: float, # Infected (hosp) -> R
-    gamma_3: float, # Infected (ICU) -> R
-    pop_1: float, # Infected (mild, not hosp) -> Infected (hosp)
-    pop_2: float, # Infected (hosp) -> Infected (ICU)
-    psi: float # Infected (ICU) -> D
-) -> Tuple[float, float, float, float, float, float]:
-    
-    """The SEIR model in BJC, one time step."""
-    
-    N = S + E + I_Mild + I_Hosp + I_ICU + R + D
-    
-    dS = (- beta_1 * S * I_Mild - beta_2 * S * I_Hosp - beta_3 * S * I_ICU)
-    dE = (beta_1 * S * I_Mild + beta_2 * S * I_Hosp + beta_3 * S * I_ICU) - alpha * E
-    dI_Mild = alpha * E  - gamma_1 * I_Mild - pop_1 * I_Mild
-    dI_Hosp = pop_1 * I_Mild  - gamma_2 * I_Hosp - pop_2 * I_Hosp
-    dI_ICU  = pop_2 * I_Hosp  - gamma_3 * I_ICU - psi * I_ICU
-    dR = gamma_1 * I_Mild + gamma_2 * I_Hosp + gamma_3 * I_ICU
-    dD = psi * I_ICU
-    
[…]

### 2020-11-08 KPMG Team/KPMG Datasets & Models/SIR/src/SEIR_bjc.py

@@ -1,207 +1,0 @@
-
-# coding: utf-8
-
-# In[1]:
-
-
-import pandas as pd
-import numpy as np
-from typing import Generator, Tuple
-from collections import namedtuple
-import matplotlib.pyplot as plt
-import copy
-
-
-# ### SEIR BJC Model
-# 
-# mimic similar code structure as the streamlit app
-
-# In[6]:
-
-
-def seir_bjc(
-    S: float,  # Susceptible
-    E: float, # Exposed 
-    IR: float, # Infected (not hosp)
-    IH: float, # Infected (hosp)
-    R: float, # Recovered (from IR)
-    H: float, # Hospitalized (from IH)
-    D: float, # Discahrged (from H)
-    beta: float, # from S -> I
-    sigma: float, # from E -> I
-    hosp_prop: float, # percent of E that go to IH
-    gamma_r: float, # IR -> R
-    gamma_h: float, # IH -> H
-    psi: float # H -> D
-) -> Tuple[float, float, float, float, float, float]:
-    
-    """The SEIR model in BJC, one time step."""
-    
-    N = S + E + IR + IH + R + H + D
-    
-    dS = - beta * S * (IR + IH)
-    dE = beta * S * (IR + IH) - sigma * E
-    dIR = sigma * E * (1 - hosp_prop) - gamma_r * IR
-    dIH = sigma * E * hosp_prop - gamma_h * IH
-    dR = gamma_r * IR
-    dH = gamma_h * IH - psi * H
-    dD = psi * H
-    
-    S_t = S + dS
-    E_t = E + dE
-    IR_t = IR + dIR
-    IH_t = IH + dIH
-    R_t = R + dR
-    H_t = H + dH
-    D_t = D + dD
-    
-    
-    if S_t < 0.0:
-        S_t = 0.0
-    if E_t < 0.0:
-        E_t = 0.0
-    if IR_t < 0.0:
-        IR_t = 0.0
-    if IH_t < 0.0:
-        IH_t = 0.0
-    if R_t < 0.0:
-        R_t = 0.0
-    if H_t < 0.0:
-        H_t = 0.0
-    if D_t < 0.0:
-        D_t = 0.0
-
-    scale = N / (S_t + E_t + IR_t + IH_t + R_t + H_t + D_t)
-    
-    # order is improtant
-    return S_t * scale, E_t * scale, IR_t * scale, IH_t * scale, R_t * scale, H_t * scale, D_t * scale
-
-
-# In[7]:
-
-
[…]

### 2020-11-08 KPMG Team/KPMG Datasets & Models/SIR/src/eval_model.py

@@ -1,38 +1,0 @@
-import pandas as pd
-import numpy as np
-
-def gen_Metrics(data, var_x, var_true):
-    '''
-    provide SSE, SAE, SSPE, SAPE
-    '''
-    
-    data = data.copy()
-    data['SSE'] = data[var_x] - data[var_true]
-    data['SAE'] = data[var_x] - data[var_true]
-    # if true is zero, ignore it 
-    data['SSPE'] = np.where(data[var_true] <= 0,
-                            0,
-                            (data[var_x] - data[var_true])/data[var_true])
-    data['SAPE'] = np.where(data[var_true] <= 0,
-                            0,
-                            (data[var_x] - data[var_true])/data[var_true])
-
-    def Sum_Square_Diff(x):
-        return sum(x**2)
-
-    def Sum_Absolute_Diff(x):
-        return sum(abs(x))
-    
-    result = data.aggregate({
-        'SSE':Sum_Square_Diff,
-        'SAE':Sum_Absolute_Diff,
-        'SSPE':Sum_Square_Diff,
-        'SAPE':Sum_Absolute_Diff
-    })
-    
-    result = pd.DataFrame(result).transpose()
-    
-    return(result)
-
-
-

### 2020-11-08 EPA_Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
Re-upload

--- 2020-10-29 EPA:Census Team/Alison_Sociodemographic_Visualizations/Social_Demo_Chars_tidycensus.Rmd
@@
 
-Man_city_sex_age_acs5_2018 <- MI_sex_age_acs5_2018 %>%
+Man_city_sex_age_acs5_2018 <- MI_place_sex_age_acs5_2018 %>%
   slice(18425:18471) %>%

### 2020-11-08 EPA_Census Team/Alisha/earn_industry_manistee_1.R

@@ -1,0 +1,54 @@
+library(plm)
+library(plyr)
+library(ggplot2)
+library(tidyverse)
+library(bea.R)
+#install.packages(httr)
+library(httr)
+
+beakey = '386344DA-444A-4E59-A4FC-4DA15935C1A3'
+
+paramvals <- beaParamVals(beakey, 'Regional', 'LineCode' )
+paramvals <- as.data.frame(paramvals)
+
+
+line1 = c( '400', '500', '600', '700', '1000', '1100', '1200', '1700','1800')
+
+
+beaearnindtotal<- data.frame()
+
+
+for (i in line1) {
+  beaspecs2 <- list(
+    'UserID' = beakey,
+    'Method' = 'GetData', 
+    'DatasetName' = 'Regional',
+    'TableName' = 'CAINC5N',
+    'LineCode' = i,
+    'GeoFips' = '26101',
+    'Year' = '2013,2014,2015,2016,2017,2018',
+    'ResultFormat' = 'json'
+  )
+  beaearningsind <- beaGet(beaspecs2, asWide = FALSE)
+  beaearnindtotal <- rbind(beaearnindtotal, beaearningsind, fill=TRUE)
+}
+
+by_year<- beaearnindtotal %>% group_by(TimePeriod)
+
+earn_plot<- ggplot(beaearnindtotal, aes(x = TimePeriod , y= DataValue, fill=factor(Code))) + ggtitle("Earnings Per Industry for Manistee County, Michigan") +
+  geom_bar(stat= "identity", position="dodge") +
+  scale_fill_discrete(name="Industry",
+                      labels=c("CAINC5N-400" = "Construction", 
+                               "CAINC5N-500" = "Manufacturing", "CAINC5N-600" = "Wholesale trade", "CAINC5N-700" = "Retail trade",
+                               "CAINC5N-1000" = "Finance and insurance",
+                               "CAINC5N-1100" = "Real estate and rental and leasing", "CAINC5N-1200" = "Professional, scientific, and technical services",
+                              "CAINC5N-1700" = "Arts, entertainment, and recreation", "CAINC5N-1800" = "Accommodation and food services")) +   
+  labs(
+    x = "Year",
+    y = "Earnings per Industry"
+  )
+
+earn_plot
+
+#ggplotly(earn_plot) <- commented this out because I'm not sure why but when I run this all of my legend labels dissapear
+

### 2020-11-08 EPA_Census Team/app.R

@@ -1,0 +1,237 @@
+# 1 load libraries ----
+library(shiny)
+library(shinydashboard)
+library(tidyverse)
+
+# 2 run scripts ----
+# * 2.1 Housing (Asahi) --------------------------------------------------------
+source("Asahi_housing/housing.r")
+source("Asahi_housing/cost_income.R")
+# * 2.2 Sociodemographic (Alison) --------------------------------------------------------
+source("viz_scripts/alison_sociodem.R")
+# * 2.3 Demographics (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_demo.R")
+# * 2.4 Walkability (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_walk.R")
+# * 2.5 Business/Industry (Kyung) --------------------------------------------------------
+source("viz_scripts/kyung_business.R")
+# * 2.6 Broadband (Gretchen) --------------------------------------------------------
+source("viz_scripts/gretchen_bband.R")
+# * 2.7 Households (Alison) --------------------------------------------------------
+source("viz_scripts/alison_childcare.R")
+# * 2.8 Education + Earnings (Alisha) --------------------------------------------------------
+source("viz_scripts/alisha_earn.R")
+
+# Define UI for application that draws a histogram
+ui <- dashboardPage(
+  dashboardHeader(title = "Columbia TOP Tool"),
+  dashboardSidebar(
+    sidebarMenu(
+      menuItem("Demographics",
+               tabName = "demo_tab",
+               icon = icon("dashboard")),
+      menuItem("Childcare",
+               tabName = "child_tab",
+               icon = icon("dashboard")),
+      menuItem("Quality of Life",
+               tabName = "qol_tab",
+               icon = icon("dashboard")),
+      menuItem("Housing",
+               tabName = "housing_tab",
+               icon = icon("dashboard")),
+      menuItem("Industry",
+               tabName = "industry_tab",
+               icon = icon("dashboard")),
+      menuItem("Broadband",
+               tabName = "bband_tab",
+               icon = icon("dashboard"))
+    )
+  ),
+  dashboardBody(
+    fluidRow(
+        box(textInput(inputId = "location", label = "Location", value = NULL, placeholder = "Enter location")
+      )
+    ),
+    fluidRow(
+      tabItems(
+        tabItem(tabName = "demo_tab",
+                box(
+                  solidHeader = TRUE,
+                  collapsible = TRUE,
[…]

### 2020-11-08 EPA_Census Team/viz_scripts/alisha_earn.R

@@ -1,0 +1,54 @@
+library(plm)
+library(plyr)
+library(ggplot2)
+library(tidyverse)
+library(bea.R)
+#install.packages(httr)
+library(httr)
+
+beakey = '386344DA-444A-4E59-A4FC-4DA15935C1A3'
+
+paramvals <- beaParamVals(beakey, 'Regional', 'LineCode' )
+paramvals <- as.data.frame(paramvals)
+
+
+line1 = c( '400', '500', '600', '700', '1000', '1100', '1200', '1700','1800')
+
+
+beaearnindtotal<- data.frame()
+
+
+for (i in line1) {
+  beaspecs2 <- list(
+    'UserID' = beakey,
+    'Method' = 'GetData', 
+    'DatasetName' = 'Regional',
+    'TableName' = 'CAINC5N',
+    'LineCode' = i,
+    'GeoFips' = '26101',
+    'Year' = '2013,2014,2015,2016,2017,2018',
+    'ResultFormat' = 'json'
+  )
+  beaearningsind <- beaGet(beaspecs2, asWide = FALSE)
+  beaearnindtotal <- rbind(beaearnindtotal, beaearningsind, fill=TRUE)
+}
+
+by_year<- beaearnindtotal %>% group_by(TimePeriod)
+
+earn_plot<- ggplot(beaearnindtotal, aes(x = TimePeriod , y= DataValue, fill=factor(Code))) + ggtitle("Earnings Per Industry for Manistee County, Michigan") +
+  geom_bar(stat= "identity", position="dodge") +
+  scale_fill_discrete(name="Industry",
+                      labels=c("CAINC5N-400" = "Construction", 
+                               "CAINC5N-500" = "Manufacturing", "CAINC5N-600" = "Wholesale trade", "CAINC5N-700" = "Retail trade",
+                               "CAINC5N-1000" = "Finance and insurance",
+                               "CAINC5N-1100" = "Real estate and rental and leasing", "CAINC5N-1200" = "Professional, scientific, and technical services",
+                               "CAINC5N-1700" = "Arts, entertainment, and recreation", "CAINC5N-1800" = "Accommodation and food services")) +   
+  labs(
+    x = "Year",
+    y = "Earnings per Industry"
+  )
+
+earn_plot
+
+#ggplotly(earn_plot) <- commented this out because I'm not sure why but when I run this all of my legend labels dissapear
+

### 2020-11-08 EPA_Census Team/viz_scripts/alison_sociodem.R

@@ -1,0 +1,126 @@
+
+library(knitr)
+library(tidyverse)
+library(tidycensus)
+library(ggplot2)
+library(ggthemes)
+library(jcolors)
+
+Sys.getenv("CENSUS_API_KEY")
+
+acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
+
+write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
+
+
+sex_age_vars <- acs5_2018_vars %>%
+  slice(5:51)
+
+sex_age_vec <- sex_age_vars %>%
+  select(name) %>%
+  unlist()
+
+sex_age_label <- sex_age_vars %>%
+  select(label) %>%
+  unlist()
+
+
+
+MI_place_sex_age_acs5_2018 <- get_acs(geography = "place",
+                                      variables = sex_age_vec,
+                                      place = "Manistee",
+                                      state = "MI",
+                                      year = 2018)
+
+Man_city_sex_age_acs5_2018 <- MI_place_sex_age_acs5_2018 %>%
+  slice(18425:18471) %>%
+  mutate(sex_age_label = sex_age_label) %>%
+  select(sex_age_label, estimate)
+
+
+
+# https://github.com/walkerke/tidycensus/issues/131
+
+years <- lst(2013, 2018)
+
+MI_place_mob_acs5_2013_2018 <- map_dfr(
+  years,
+  ~ get_acs(
+    geography = "place",
+    variables = c("B07001_016", "B07001_033", "B07001_049", "B07001_065", "B07001_081"),
+    state = "MI",
+    place = "Manistee",
+    year = .x
+  ),
+  .id = "year"
+)
+
+
+
+Man_city_mob_acs5_2013_2018 <- MI_place_mob_acs5_2013_2018 %>%
+  filter(NAME == "Manistee city, Michigan") %>%
+  add_column(label = "") %>%
+  mutate(label = case_when(variable == "B07001_016" ~ "Same house",
+                           variable == "B07001_033" ~ "Same county",
+                           variable == "B07001_049" ~ "Different county",
+                           variable == "B07001_065" ~ "Different state",
+                           variable == "B07001_081" ~ "Moved from abroad")) %>%
+  mutate(label = factor(label, levels = c("Same house", "Same county", "Different county", "Different state", "Moved from abroad")))
+
+
+
+p1_mobility <- ggplot(data = Man_city_mob_acs5_2013_2018, aes(x = label, y = estimate, fill = year)) + 
[…]

### 2020-11-08 EPA_Census Team/viz_scripts/gretchen_demo.R

@@ -1,0 +1,48 @@
+# Data you need to plot is in this folder
+# See "Data Manipulation" folder for OG data & manipulation code" 
+
+library(tidyverse)
+library(ggplot2)
+library(scales)
+
+just_sex <- read.csv("data_files/just_sex")
+age_breakdown_manistee <- read.csv("data_files/age_breakdown_manistee")
+race_data_manistee <- read.csv("data_files/race_data_manistee")
+
+# create "Blank_theme" to remove all the axis markings for pie chart 
+blank_theme <- theme_minimal()+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x=element_blank(),
+    axis.title.y = element_blank(),
+    panel.border = element_blank(),
+    panel.grid=element_blank(),
+    axis.ticks = element_blank(),
+    plot.title=element_text(size=14, face="bold")
+  )
+
+# Create Chart (in ggplot2 you have to make a bar chart first, then change "coordinates" to make it into a pie)
+genderbar<- ggplot(just_sex, aes(x="", y=`Manistee.County..MI`, fill=`label`))+
+  geom_bar(width = 1, stat = "identity")
+genderpie <- genderbar + coord_polar("y", start=0) +
+  geom_text(aes(y = `Manistee.County..MI`/3 + c(0, cumsum(`Manistee.County..MI`)[-length(`Manistee.County..MI`)]), 
+                label = percent(`Manistee.County..MI`/sum(`Manistee.County..MI`), accuracy=1)), size=5) + 
+  ggtitle(label = "Sex") +
+  scale_fill_discrete(name = "", labels = c("Female", "Male")) +
+  blank_theme
+genderpie 
+
+#plot
+agebar <- ggplot(age_breakdown_manistee, aes(x= `NewCat`, y=`Per`)) + 
+  geom_bar(width = .8, stat = "identity") + 
+  labs(title="Population by Age Range for Manistee County") + 
+  xlab("Age Range") + ylab("Percent of Population")
+agebar
+
+# Note: "Hispanic" is included in 2+ category and "hispanic" category, but not others 
+racebar <- ggplot(race_data_manistee, aes(x= `CompRace`, y=as.numeric(`Per` * 100))) + 
+  geom_bar(width = .8, stat = "identity") + 
+  scale_y_continuous(limits = c(0,90), expand = c(0, 0)) +
+  labs(title="Race & Ethnicity for Manistee County") + 
+  xlab("Race") + ylab("Percent of Population")
+racebar

### 2020-11-08 KPMG Team/KPMG Datasets & Models/SIR/src/eval_model.py

@@ -1,0 +1,38 @@
+import pandas as pd
+import numpy as np
+
+def gen_Metrics(data, var_x, var_true):
+    '''
+    provide SSE, SAE, SSPE, SAPE
+    '''
+    
+    data = data.copy()
+    data['SSE'] = data[var_x] - data[var_true]
+    data['SAE'] = data[var_x] - data[var_true]
+    # if true is zero, ignore it 
+    data['SSPE'] = np.where(data[var_true] <= 0,
+                            0,
+                            (data[var_x] - data[var_true])/data[var_true])
+    data['SAPE'] = np.where(data[var_true] <= 0,
+                            0,
+                            (data[var_x] - data[var_true])/data[var_true])
+
+    def Sum_Square_Diff(x):
+        return sum(x**2)
+
+    def Sum_Absolute_Diff(x):
+        return sum(abs(x))
+    
+    result = data.aggregate({
+        'SSE':Sum_Square_Diff,
+        'SAE':Sum_Absolute_Diff,
+        'SSPE':Sum_Square_Diff,
+        'SAPE':Sum_Absolute_Diff
+    })
+    
+    result = pd.DataFrame(result).transpose()
+    
+    return(result)
+
+
+

### 2020-11-09 EPA_Census Team/Alison_Sociodemographic_Visualizations/_Old/ACS_5-Year_2014-2018_Social_Demographic_Characteristics_Manistee.Rmd

```{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
```

```{r}
acs_social <- read_csv("ACS_5-Year_2014-2018_Social_Characteristics_Manistee.csv")
acs_demo <- read_csv("ACS_5-Year_2014-2018_Demographic_Characteristics_Manistee.csv")
```

```{r}
acs_social_clean <- acs_social %>%
  rename(Michigan = `Michigan!!Estimate`) %>%
  rename(Manistee_City = `Manistee city, Michigan!!Estimate`) %>%
  select(Label, Michigan, Manistee_City)

acs_demo_clean <- acs_demo %>%
  rename(Michigan = `Michigan!!Estimate`) %>%
  rename(Manistee_City = `Manistee city, Michigan!!Estimate`) %>%
  select(Label, Michigan, Manistee_City)
```

```{r}
acs_social_clean_wide <- acs_social_clean %>%
  select(Label, Manistee_City) %>%
  slice(2:19) %>%
  spread(key = Label, value = Manistee_City)
```

```{r}
acs_social <- read_csv("ACS_5-Year_2014-2018_Social_Characteristics_Manistee_WIDE.csv")
acs_demo <- read_csv("ACS_5-Year_2014-2018_Demographic_Characteristics_Manistee_WIDE.csv")
```

```{r}
acs_demo_age <- acs_demo %>%
  slice(7) %>%
  select(`SEX AND AGE!!Total population!!Under 5 years`:`SEX AND AGE!!Total population!!85 years and over`)

acs_demo_age_clean <- acs_demo_age %>%
  rename(`Under 5` = `SEX AND AGE!!Total population!!Under 5 years`) %>%
  rename(`5 to 9` = `SEX AND AGE!!Total population!!5 to 9 years`) %>%
  rename(`10 t0 14` = `SEX AND AGE!!Total population!!10 to 14 years`) %>%
  rename(`15 to 19` = `SEX AND AGE!!Total population!!15 to 19 years`) %>%
  rename(`20 to 24` = `SEX AND AGE!!Total population!!20 to 24 years`) %>%
  rename(`25 to 34` = `SEX AND AGE!!Total population!!25 to 34 years`) %>%
  rename(`35 to 44` = `SEX AND AGE!!Total population!!35 to 44 years`) %>%
  rename(`45 to 54` = `SEX AND AGE!!Total population!!45 to 54 years`) %>%
  rename(`55 to 59` = `SEX AND AGE!!Total population!!55 to 59 years`) %>%
  rename(`60 to 64` = `SEX AND AGE!!Total population!!60 to 64 years`) %>%
  rename(`65 to 74` = `SEX AND AGE!!Total population!!65 to 74 years`) %>%
  rename(`75 to 84` = `SEX AND AGE!!Total population!!75 to 84 years`) %>%
  rename(`85 and older` = `SEX AND AGE!!Total population!!85 years and over`)
```

```{r}
ggplot(acs_demo_age_clean) +
  geom_bar(aes(estimate))
```

```{r}
acs_demo_Manistee <- acs_demo %>%
  slice(7)
```

### 2020-11-09 EPA_Census Team/Getting_started_in_tidycensus.Rmd

### Getting started in tidycensus

Adapted from: https://walker-data.com/tidycensus/articles/basic-usage.html

```{r}
library(tidyverse)
library(tidycensus)
```

tidycensus uses the Census API, so you will need to request a key: https://api.census.gov/data/key_signup.html

To keep your API key private, run the "census_api_key" function:

```{r}
census_api_key("YOUR API KEY GOES HERE", overwrite = FALSE, install = TRUE)
```

From then on you can simply call the key from your system environment:

```{r}
Sys.getenv("CENSUS_API_KEY")
```

To search through all available variables for a given survey and year, use the "load_variables" function:

```{r}
acs5_2018_vars <- load_variables(2018, "acs5", cache = TRUE)
```

You can also export the CSV in order to view the full text of each cell:

```{r}
write_csv(acs5_2018_vars, "acs5_2018_vars.csv")
```

Here is an example call for Manistee County:

```{r}
Man_county_pop_hous <- get_acs(geography = "county",
                          variables = c(population = "B00001_001", housing_units = "B00002_001"),
                          state = "Michigan",
                          county = "Manistee",
                          survey = "acs5",
                          year = 2018)
```

Note that the FIPS code for Michigan is "26" and for Manistee County is "101", so the full GEOID is "26101". You can also call these FIPS codes directly instead of the place names:

```{r}
Man_county_pop_hous <- get_acs(geography = "county",
                          variables = c(population = "B00001_001", housing_units = "B00002_001"),
                          state = "26",
                          county = "101",
                          survey = "acs5",
                          year = 2018)
```

### 2020-11-10 KPMG Team/Dashboard/app.R
dashboard base

@@ -1,0 +1,44 @@
+
+# ------------------------------------------------------------------------------------ # 
+# ************************************************************************************ #
+
+# COVID-19 Intelligent Forecasting Dashboard Functions 
+# KPMG x QMSS Practicum 2020 
+# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong 
+  
+# Dashboard Lead: Louisa Ong
+
+# ------------------------------------------------------------------------------------ # 
+# ************************************************************************************ #
+
+
+# rShiny dashboard app app.R file
+
+library(shiny)
+
+
+# define UI 
+
+ui <- pageWithSidebar(
+  
+  # App title 
+  headerPanel("COVID-19 Intelligent Forecasting"), 
+  
+  # Sidebar panel for inputs 
+  sidebarPanel(), 
+  
+  # Main panel for displaying outputs 
+  mainPanel()
+  
+)
+
+
+server <- function(input, output) {
+  
+}
+
+
+shinyApp(ui, server)
+
+
+runApp()

### 2020-11-10 KPMG Team/Dashboard/msa_areas.Rmd
msa added

--- 2020-10-29 KPMG Team/Visualizations/Louisa - Map/Metro Areas/metro_areas.Rmd
@@
 ```{r}
+### ----- SET DATE RANGE FIRST ----- ## 
+```
+
+```{r}
+
 # ------------------------------------------------------- #
-# -----------  DATA MANIPULATION FOR CBSA --------------- # 
+# -----------  DATA MANIPULATION FOR MSA --------------- # 
 # ------------------------------------------------------- #
 
-# --- *** --- core-based statistical areas  --- *** --- 
+# --- *** --- metropolitan statistical areas  --- *** --- 
 # comprised of counties 
@@
 
-# select only cbsa-related information 
-data_cbsa <- data %>% 
+# select only msa-related information 
+data_msa <- data %>% 
   mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
@@
   mutate(fips_county = `FIPS County Code`) %>%
-  rename(cbsa = `CBSA Code`) %>%
-  rename(cbsa_title = `CBSA Title`) %>% 
-  select(cbsa, cbsa_title, fips, state, fips_state, fips_county)
+  rename(msa = `Metropolitan Division Code`) %>%
+  rename(msa_title = `Metropolitan Division Title`) %>% 
+  select(msa, msa_title, fips, state, fips_state, fips_county)
 
-# join cbsa data and census data by GEOID/FIPS, get population (and other demographic) info.
-census_cbsa_merge <- census %>% 
+# join msa data and census data by GEOID/FIPS, get population (and other demographic) info.
+census_msa_merge <- census %>% 
   rename(fips = GEOID) %>% 
-  left_join(data_cbsa, by = "fips") %>% 
-  group_by(cbsa) %>%
-  mutate(cbsa_population = sum(population))
+  left_join(data_msa, by = "fips") %>% 
+  group_by(msa) %>%
+  mutate(msa_population = sum(population))
 
 # -------- FILTER OR NOT? ----------- # 
-# only top 12 csa and census data (run this)
+# only top 12 msa and census data (run this)
 
-# list of top cbsas by population, if we want
-# top_cbsa_popn <- census_cbsa_merge %>%
-#   select(cbsa, cbsa_title, cbsa_population) %>%
-#   arrange(desc(cbsa_population)) %>%
+# list of top msas by population, if we want
+# top_msa_popn <- census_msa_merge %>%
+#   select(msa, msa_title, msa_population) %>%
+#   arrange(desc(msa_population)) %>%
 #   distinct() %>%
@@
 
-# census_cbsa <- census_cbsa_merge %>%
-#   semi_join(top_cbsa_popn, by="cbsa")
+# census_msa <- census_msa_merge %>%
+#   semi_join(top_msa_popn, by="msa")
 
@@
 # without filtering (run this)
-census_cbsa <- census_cbsa_merge
+census_msa <- census_msa_merge
 # ------------------------------------ #
 
-# only need county to merge both databsets, will keep just metro areas (csa)
-cbsa_covid_merge <- census_cbsa %>% 
+# only need county to merge both datasets, will keep just metro areas (csa)
+msa_covid_merge <- census_msa %>% 
   left_join(nyt, by = "fips") %>% 
-  select(cbsa, cbsa_title, cbsa_population, date, cases, deaths) %>% 
+  select(msa, msa_title, msa_population, date, cases, deaths) %>% 
   filter(date >= lubridate::as_date("2020-01-24") & 
-           date <= lubridate::as_date("2020-10-18")) # so that the 7 day mean wouldn't get messed up
+           date <= lubridate::as_date(Sys.Date() - 8)) # 8 days prior, so that the 7 day mean wouldn't get messed up
 
 # get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
-cbsa_covid_cases <- data.frame(cbsa_covid_merge) %>% 
-  group_by(cbsa, date) %>% 
+msa_covid_cases <- data.frame(msa_covid_merge) %>% 
+  group_by(msa, date) %>% 
   dplyr::summarise(cases_sum = sum(cases, na.rm = TRUE)) %>% 
@@
 # same as above, but for deaths
-cbsa_covid_deaths <- data.frame(cbsa_covid_merge) %>% 
-  group_by(cbsa, date) %>% 
+msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
+  group_by(msa, date) %>% 
   dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
@@
 # contains
-cbsa_covid <- cbsa_covid_merge %>% 
-  left_join(cbsa_covid_cases, by = c("cbsa", "date")) %>% 
-  left_join(cbsa_covid_deaths, by = c("cbsa", "date")) %>% 
-  arrange(desc(cbsa_population))
+msa_covid <- msa_covid_merge %>% 
+  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
+  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
+  arrange(desc(msa_population))
 
-# final count of cbsa cases, in order to get ranking
-cbsa_covid_sum <- cbsa_covid %>% 
-  group_by(cbsa, cbsa_title, cbsa_population) %>%
+# final count of msa cases, in order to get ranking
+msa_covid_sum <- msa_covid %>% 
+  group_by(msa, msa_title, msa_population) %>%
   mutate(cases = sum(cases)) %>% 
@@
   arrange(desc(cases)) %>% 
-  select(cbsa, cbsa_title, cbsa_population, cases, deaths) %>% 
-  distinct() %>% 
+  select(msa, msa_title, msa_population, cases, deaths) %>% 
+  distinct() %>% # remove repeated values 
   na.omit()
+  # drop_na(msa) # remove rows with NA in msa column
 
@@
 # time series data of all new cases per day (7-day rolling average) during time period specified
-# names: names(cbsa_covid)
-head(cbsa_covid, 10)
+# names: names(msa_covid)
+head(msa_covid, 50)
 
-# Rank of CBSAs according to total cases
-head(cbsa_covid_sum, 10)
+# Rank of MSAs according to total cases
+head(msa_covid_sum, 10)
 
@@
   filter(date >= lubridate::as_date("2020-01-24") & 
-         date <= lubridate::as_date("2020-10-18")
+         date <= lubridate::as_date(Sys.Date() - 8))
 
@@
 head(state_covid, 10)
+
 ```
@@
 
-metroData <- function(cbsa_id) {
-  cbsa_covid %>% 
-    filter(cbsa == cbsa_id)
+metroData <- function(msa_id) {
+  msa_covid %>% 
+    filter(msa == msa_id)
 }
@@
 
+# ---- To Find MSA codes ---- #
+# head(msa_covid_sum, 10)
+
 # --------------------- TO MAKE A PLOT FOR ONE STATE -------------------------- #
@@
 fl_covid <- stateData("Florida")
-miami_covid <- metroData(33100) # from cbsa table
+miami_covid <- metroData(33124) # from msa table
 
@@
 ny_covid <- stateData("New York")
-nyc_covid <- metroData(35620)
+nyc_covid <- metroData(35614)
 
@@
 tx_covid <- stateData("Texas")
-dal_covid <- metroData(19100)
+dal_covid <- metroData(19124)
 

### 2020-11-10 KPMG Team/Dashboard/app.R

@@ -15,6 +15,7 @@
 # rShiny dashboard app app.R file
 
 library(shiny)
+library(shinythemes)
 
 
 # define UI 
@@ -28,7 +29,11 @@ ui <- pageWithSidebar(
   sidebarPanel(), 
   
   # Main panel for displaying outputs 
-  mainPanel()
+  mainPanel(
+    
+    # using a module 
+    # metroArea("", "")
+  )
   
 )
 
@@ -38,7 +43,13 @@ server <- function(input, output) {
 }
 
 
-shinyApp(ui, server)
+# run the app
+shinyApp(
+  
+  ui,# = fluidPage(theme = shinytheme("sandstone")), 
+  server
+  
+)
 
 
-runApp()
+# runApp()

### 2020-11-10 KPMG Team/Dashboard/functions.R

@@ -10,3 +10,19 @@
 
 # ------------------------------------------------------------------------------------ # 
 # ************************************************************************************ #
+
+# module server function
+metroAreas <- function(id, label = "Metro Areas") {
+  
+  moduleServer(
+    id, 
+    
+    # actual module function
+    function(input, output, session) {
+      
+    return()
+      
+    }
+  )
+  
+}

### 2020-11-11 KPMG Team/Dashboard/msa_areas.Rmd
state county map viz
MSA Latest
with link to socialexplorer variables

@@
 # census data API, variables can be changed here 
+# find variable names: https://www.socialexplorer.com/data/ACS2010/metadata/?ds=ACS10
 census <- get_acs(geography = "county",
@@
 ```
+
+```{r}
+
+# Plot State Map with Metro Areas
+
+# mapping packages needed 
+library(leaflet)
+library(maptools)
+library(rgdal)
+
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
+
+```

### 2020-11-11 KPMG Team/Dashboard/msa_areas_map.Rmd

--- 2020-11-10 KPMG Team/Dashboard/msa_areas.Rmd
@@
 # ------------------------------------ #
+names(nyt)
+
+county_sum <- nyt %>% 
+  group_by(county, state, fips) %>%
+  summarise(cases_sum = sum(cases, na.rm = TRUE))
 
@@
 # Rank of MSAs according to total cases
-head(msa_covid_sum, 10)
+head(msa_covid_sum, 50)
 
@@
 # ---- To Find MSA codes ---- #
+
 # head(msa_covid_sum, 10)
@@
 ```
+
+```{r}
+
+# Plot State Map with Metro Areas
+
+# mapping packages needed 
+library(leaflet)
+library(maptools)
+library(rgdal)
+
+# interactive visualization
+library(plotly)
+
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
+
+library(tigris)
+library(usmap) # to get US Map Data
+
+library(jsonlite) # from FIPS -> Geo (lat, long)
+
+# states in Texas 
+plot_usmap(include = .midwest_region, labels = TRUE)
+
+# get FIPS codes (state + county)
+# fips(state = "CA", county = "Orange")
+
+# FIPS county codes to Lat Long
+geoFips <- function() {
+
+}
+
+# get lat long data from US Census 
+# https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html
+county_lines <- st_read("county_shape/cb_2018_us_county_500k.shp")
+
+# add FIPS as a variable from State Fips and County Fips
+county_lines <- county_lines %>% 
+  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))
+
+# add lat long attributes to county dataset          
+county_geo <- merge(county_lines, county_sum, by = "fips")
+county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')
+
+outside_mainland <- c("02", "15", "72", "2", "3")
+islands <- c("Virgin Islands", "Northern Mariana Islands")
+
+county_map_US <- county_map %>% 
+  filter(!STATEFP %in% outside_mainland) %>% 
+  filter(!state %in% islands) # remove alaska and hawaii
+
+ggplot(data = county_map_US) +
+  geom_sf(aes(fill = cases_sum)) + 
+  scale_fill_viridis_c(option = "magma", direction = 1)
+
+county_map_NY <- county_map %>% 
+  filter(STATEFP == 36) # remove alaska and hawaii
+
+ggplot(data = county_map_NY) +
+  geom_sf(aes(fill = cases_sum)) + 
+  scale_fill_viridis_c(option = "magma", direction = -1)
+
+names(county_map_US)
+
+fips(state = "New York")
+
+table(c(county_map_US$statefp, county_map_US$state))
+
+# PLOT FUNCTION 
+# county state map 
+county_state_map <- function(state_name) {
+
+  fips_code <- fips(state = state_name)
+
+  county_map_state <- county_map %>% 
+  filter(STATEFP == fips_code)
+
+  ggplot(data = county_map_state) +
+  geom_sf(aes(fill = cases_sum)) + 
+  scale_fill_viridis_c(option = "magma", direction = -1)
+
+}
+
+W <- county_state_map("Wisconsin")
+
+```

### 2020-11-11 KPMG Team/Dashboard/visualization_scripts/cities.Rmd
cities area

---
title: "Cities (Deaths & Case)"
output: html_notebook
author: Louisa Ong (QMSS Team)
---

```{r}
```{r}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus) # US census bureau data
library(zoo) # rolling mean

# plotting 
library(ggplot2)
# library(hrbrthemes) 
# library(geofacet)

# ----------- LOADING DATASETS ----------- #

#nytimes county-level cases & deaths data from online link 
nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")

# don't need it, unless we want more variables
# census data API, variables can be changed here 
# census <- get_acs(geography = "county",
#                    variables = c(population = "B01003_001E"), # total population), # black veterans 
#                    output = "wide",
#                    county = "",
#                    year = 2018)

# US City & FIPS code boundaries 
# https://simplemaps.com/data/us-cities license 
city <- read_csv("uscities.csv")

View(city)
names(city)
```

```{r}
names(city)

# ------------------------------------------------------- # 
# county covid sum cases & deaths 
county_sum <- nyt %>%
  filter(date == max(nyt$date))

# ---------------------------------------------- #
# -----------  DATA MANIPULATION --------------- # 
# ---------------------------------------------- #

# CITY DATASET 
data_city <- city %>% 
  rename(fips = county_fips) %>% 
  group_by(city) %>% 
  mutate(city_population = sum(population))

View(nyt)

# CITY & COVID MERGE 
# only need county to merge both datasets, will keep just city areas 
city_covid_merge <- data_city %>% 
  left_join(nyt, by = "fips") %>% 
  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date(Sys.Date() - 8))

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
city_covid_cases <- data.frame(city_covid_merge) %>% 
  group_by(city, date) %>% 
  dplyr::mutate(cases_sum = max(cases)) %>% # choose max cases for each date because some dates report more than once
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

View(city_covid_cases)
# same as above, but for deaths
msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
  group_by(msa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
msa_covid <- msa_covid_merge %>% 
  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
  arrange(desc(msa_population))

```

### 2020-11-11 KPMG Team/Dashboard/visualization_scripts/msa_areas_map.Rmd
cities area

--- 2020-11-11 KPMG Team/Dashboard/msa_areas_map.Rmd
@@
 
-```
+city <- read_csv("cities.csv")
 
-```{r}
-### ----- SET DATE RANGE FIRST ----- ## 
+View(city)
 ```
@@
 census_msa <- census_msa_merge
+
 # ------------------------------------ #
@@
 county_sum <- nyt %>%
-  group_by(county, state, fips) %>%
-  summarise(cases_sum = sum(cases, na.rm = TRUE))
+  filter(date == max(nyt$date)) # latest
 
@@
   group_by(msa, msa_title, msa_population) %>%
-  mutate(cases = sum(cases)) %>% 
-  mutate(deaths = sum(deaths)) %>%
-  arrange(desc(cases)) %>% 
-  select(msa, msa_title, msa_population, cases, deaths) %>% 
+  mutate(total_cases = sum(new_cases)) %>% 
+  mutate(total_deaths = sum(new_deaths)) #%>%
+  arrange(desc(total_cases)) %>% 
+  select(msa, msa_title, msa_population, total_cases, total_deaths) %>% 
   distinct() %>% # remove repeated values 
@@
   # drop_na(msa) # remove rows with NA in msa column
+
+msa_covid_sum <- msa_covid %>% 
+  filter(date == max(msa_covid$date)) %>% 
+  select(msa, msa_title, msa_population, date, cases_sum, deaths_sum) #%>% 
+  # distinct() #%>%
+  # drop_na(msa)
+
+View(msa_covid_sum)
+
+table(msa_covid$msa_title)
 
@@
 
-# states in Texas 
-plot_usmap(include = .midwest_region, labels = TRUE)
-
-# get FIPS codes (state + county)
-# fips(state = "CA", county = "Orange")
-
-# FIPS county codes to Lat Long
-geoFips <- function() {
-
-}
-
 # get lat long data from US Census 
@@
 ggplot(data = county_map_US) +
-  geom_sf(aes(fill = cases_sum)) + 
+  geom_sf(aes(fill = cases)) + 
   scale_fill_viridis_c(option = "magma", direction = 1)
@@
 
-ggplot(data = county_map_NY) +
-  geom_sf(aes(fill = cases_sum)) + 
-  scale_fill_viridis_c(option = "magma", direction = -1)
-
-names(county_map_US)
-
-fips(state = "New York")
-
-table(c(county_map_US$statefp, county_map_US$state))
+# # plot for states 
+# ggplot(data = county_map_NY) +
+#   geom_sf(aes(fill = cases)) + 
+#   scale_fill_viridis_c(option = "magma", direction = -1) +
+# 
+# 
+# names(county_map_US)
+# 
+# fips(state = "New York")
 
@@
   ggplot(data = county_map_state) +
-  geom_sf(aes(fill = cases_sum)) + 
+  geom_sf(aes(fill = cases)) + 
   scale_fill_viridis_c(option = "magma", direction = -1)
@@
 
+range(county_map_US$cases)
+
 ```

### 2020-11-11 EPA_Census Team/Basic Demographic Plots - Race, Age, Gender/Basic Demo Plots.Rmd

---
title: "Basic Demographic Plots"
author: "Gretchen Streett"
date: "11/3/2020"
output: html_document
---

```{r setup}
# Data you need to plot is in this folder
# See "Data Manipulation" folder for OG data & manipulation code"

library(tidyverse)
library(ggplot2)
library(scales)

just_sex <- read.csv("just_sex")
age_breakdown_manistee <- read.csv("age_breakdown_manistee")
race_data_manistee <- read.csv("race_data_manistee")
```

```{r Gender Pie chart}
# create "Blank_theme" to remove all the axis markings for pie chart 
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.text.x=element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )

# Create Chart (in ggplot2 you have to make a bar chart first, then change "coordinates" to make it into a pie)
genderbar<- ggplot(just_sex, aes(x="", y=`Manistee.County..MI`, fill=`label`))+
geom_bar(width = 1, stat = "identity")
genderpie <- genderbar + coord_polar("y", start=0) +
  geom_text(aes(y = `Manistee.County..MI`/3 + c(0, cumsum(`Manistee.County..MI`)[-length(`Manistee.County..MI`)]), 
            label = percent(`Manistee.County..MI`/sum(`Manistee.County..MI`), accuracy=1)), size=5) + 
  ggtitle(label = "Sex") +
  scale_fill_discrete(name = "", labels = c("Female", "Male")) +
  blank_theme
genderpie 
```

```{R Age Bar Chart}
#plot
agebar <- ggplot(age_breakdown_manistee, aes(x= `NewCat`, y=`Per`)) + 
  geom_bar(width = .8, stat = "identity") + 
  labs(title="Population by Age Range for Manistee County") + 
  xlab("Age Range") + ylab("Percent of Population")
agebar
```

```{R Race Bar Chart}
# Note: "Hispanic" is included in 2+ category and "hispanic" category, but not others 
racebar <- ggplot(race_data_manistee, aes(x= `CompRace`, y=as.numeric(`Per` * 100))) + 
  geom_bar(width = .8, stat = "identity") + 
  scale_y_continuous(limits = c(0,90), expand = c(0, 0)) +
  labs(title="Race & Ethnicity for Manistee County") + 
  xlab("Race") + ylab("Percent of Population")
racebar
```

### 2020-11-11 EPA_Census Team/Basic Demographic Plots - Race, Age, Gender/Data Manipulation/Basic Demos Data Manipulation.Rmd

---
title: "Graphing Demos"
author: "Gretchen Streett"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)

# get data - See "Data Manipulation" Folder & re-write paths
sex_by_age_all_counties <- read.csv("~/Documents/Columbia/Practicum - Aracelis Torres/Data/sex_by_age_all_counties.csv")
race_data <- read.csv("~/Documents/Columbia/Practicum - Aracelis Torres/Data/race_data.csv")
```

```{r Gender Pie}
# Filter Data
sex_by_age_manistee <- sex_by_age_all_counties[, c("United.States", "Michigan", "Manistee.County..MI", "label", "name")]
just_sex <- sex_by_age_manistee %>% filter(name == "B01001002" | name == "B01001026")
#write_csv(just_sex, "just_sex")

# Gender Pie Chart 
# create "Blank_theme" to remove all the axis markings for pie chart 
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.text.x=element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )

# Create Chart (in ggplot2 you have to make a bar chart first, then change "coordinates" to make it into a pie)
genderbar<- ggplot(just_sex, aes(x="", y=`Manistee.County..MI`, fill=`label`))+
geom_bar(width = 1, stat = "identity")
genderpie <- genderbar + coord_polar("y", start=0) +
  geom_text(aes(y = `Manistee.County..MI`/3 + c(0, cumsum(`Manistee.County..MI`)[-length(`Manistee.County..MI`)]), 
            label = percent(`Manistee.County..MI`/sum(`Manistee.County..MI`), accuracy=1)), size=5) + 
  ggtitle(label = "Sex") +
  scale_fill_discrete(name = "", labels = c("Female", "Male")) +
  blank_theme
genderpie

```

```{R Age Distribution, warning = FALSE, message = FALSE}
# Data wrangling:  (Census has men + women listed separately. Get Totals.)
age_breakdown_manistee <- sex_by_age_manistee %>% separate(label, into = c(NA, NA, "Gender", "Age"), sep = "\\!!", remove = TRUE)
age_breakdown_manistee <- age_breakdown_manistee[, c("Manistee.County..MI", "Gender", "Age")]
age_breakdown_manistee <- age_breakdown_manistee %>% group_by(`Age`) %>% summarize(Total = sum(`Manistee.County..MI`))
age_breakdown_manistee <- age_breakdown_manistee %>% filter(Age != "NA")

# Too many categories to plot well. Combine Further
age_breakdown_manistee <- age_breakdown_manistee %>% mutate(NewCat = case_when(
              Age == "Under 5 years" | Age == "5 to 9 years" ~ "0 to 9",
              Age == "10 to 14 years" | Age == "15 to 17 years" | Age == "18 and 19 years" ~ "10 to 19",
              Age == "20 years" | Age == "21 years" | Age == "22 to 24 years" | Age == "25 to 29 years" ~ "20 to 29",
              Age == "30 to 34 years" | Age == "35 to 39 years" ~ "30 - 39",
              Age == "40 to 44 years" | Age == "45 to 49 years" ~ "40 - 49",
              Age == "50 to 54 years" | Age == "55 to 59 years" ~ "50 - 59",
              Age == "60 and 61 years" | Age == "62 to 64 years" | Age == "65 and 66 years" | Age == "67 to 69 years" ~ "60 to 69",
              Age == "70 to 74 years" | Age == "75 to 79 years" ~ "70 - 79",
              Age == "80 to 84 years" | Age == "85 years and over" ~ "80+",
              TRUE ~ "else" ))
age_breakdown_manistee <- age_breakdown_manistee %>% group_by(`NewCat`) %>% summarize(Total = sum(Total))
age_breakdown_manistee <- age_breakdown_manistee %>% mutate(Per = percent(Total/sum(Total), accuracy = 1))
#write_csv(age_breakdown_manistee, "age_breakdown_manistee")

#plot
agebar <- ggplot(age_breakdown_manistee, aes(x= `NewCat`, y=`Per`)) + 
  geom_bar(width = .8, stat = "identity") + 
  labs(title="Population by Age Range for Manistee County") + 
  xlab("Age Range") + ylab("Percent of Population")
agebar
```

```{R Race Distribution, warning = FALSE, message = FALSE}
# filter data 
race_data_manistee <- race_data[, c("Manistee.County..MI", "label", "name")]
race_data_manistee <- race_data_manistee %>% separate(col = label, into = c(NA, NA, "Hispanic", "Race"), sep="!!", remove = TRUE)
race_data_manistee <- race_data_manistee %>% filter(`Hispanic` != "NA")
hispanic_data <- race_data_manistee %>% filter(name == "B03002012")
race_data_manistee <- race_data_manistee %>% filter(Hispanic != "Hispanic or Latino")
race_data_manistee <- rbind(race_data_manistee, hispanic_data)

# Add better Race names for ease of graphing 
race_data_manistee <- race_data_manistee %>% mutate(CompRace = case_when(
              name == "B03002002" ~ "Not Hispanic Total",
              name == "B03002003" ~ "White",
              name == "B03002004" ~ "Black",
              name == "B03002005" ~ "American Indian",
              name == "B03002006" ~ "Asian",
              name == "B03002007" ~ "Pacific Islander",
              name == "B03002008" ~ "Other",
              name == "B03002009" | name == "B03002010" | name == "B03002011" ~ "Two or More",
              name == "B03002012" ~ "Hispanic or Latino", 
              TRUE ~ "Else"))
race_data_manistee <- race_data_manistee %>% group_by(CompRace) %>% summarize(Total = sum(`Manistee.County..MI`))
race_data_manistee <- filter(race_data_manistee, CompRace != "Not Hispanic Total")
race_data_manistee <- race_data_manistee %>% mutate(Per = Total/sum(Total))

#plot 
# Turn race into an ordered factor 
race_data_manistee$CompRace <- factor(race_data_manistee$CompRace, levels = race_data_manistee$CompRace[order(race_data_manistee$Per)])
#write_csv(race_data_manistee, "race_data_manistee")

racebar <- ggplot(race_data_manistee, aes(x= `CompRace`, y=as.numeric(`Per` * 100))) + 
  geom_bar(width = .8, stat = "identity") + 
  scale_y_continuous(limits = c(0,90), expand = c(0, 0)) +
  labs(title="Race & Ethnicity for Manistee County") + 
  xlab("Race") + ylab("Percent of Population")
racebar
```

### 2020-11-12 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd

--- 2020-11-11 KPMG Team/Dashboard/visualization_scripts/msa_areas_map.Rmd
@@
 ---
+
+# Data Manipulation
 
@@
 library(ggplot2)
-library(hrbrthemes) 
-library(geofacet)
+# library(hrbrthemes) 
+# library(geofacet)
 
@@
 # metropolitan state areas delineation from file taken off Census Bureau
-data <- read_csv("msa.csv")
+data <- read_csv("data/msa.csv")
 
 #nytimes county-level cases & deaths data from online link 
-nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
+covid <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
 
-# census data API, variables can be changed here 
+# census data API, variables can be changed and added here 
 census <- get_acs(geography = "county",
-                   variables = c(population = "B01003_001E"), # total population), # black veterans 
+                   variables = c(population = "B01003_001E"), # total population
                    output = "wide",
                    county = "",
-                   year = 2018)
+                   year = 2018) # ACS 2018 Survey
 
-city <- read_csv("cities.csv")
+# US City & FIPS code boundaries 
+# https://simplemaps.com/data/us-cities license 
+city <- read_csv("data/uscities.csv")
 
-View(city)
+```
+
+```{r}
+
+# ---------------------------------------------------------- #
+# -----------  DATA MANIPULATION FOR COUNTY  --------------- # 
+# ---------------------------------------------------------- #
+
+# not much to do
+county_sum <- covid %>%
+  filter(date == max(covid$date)) # latest
+
+# Preview 
+head(county_sum, 50)
+
 ```
@@
 # --- *** --- metropolitan statistical areas  --- *** --- 
+# only 31 MSA in this list
 # comprised of counties 
@@
 # list of top msas by population, if we want
-# top_msa_popn <- census_msa_merge %>%
-#   select(msa, msa_title, msa_population) %>%
-#   arrange(desc(msa_population)) %>%
-#   distinct() %>%
-#   na.omit() %>%
-#   data.frame() %>%
-#   slice(1:12)
+top_msa_popn <- census_msa_merge %>%
+  select(msa, msa_title, msa_population) %>%
+  arrange(desc(msa_population)) %>%
+  distinct() %>%
+  na.omit() %>%
+  data.frame() %>%
+  slice(1:12)
 
-# census_msa <- census_msa_merge %>%
-#   semi_join(top_msa_popn, by="msa")
+top_census_msa <- census_msa_merge %>%
+  semi_join(top_msa_popn, by="msa")
 
@@
 # ------------------------------------ #
-names(nyt)
-
-county_sum <- nyt %>%
-  filter(date == max(nyt$date)) # latest
 
@@
 msa_covid_merge <- census_msa %>% 
-  left_join(nyt, by = "fips") %>% 
+  left_join(covid, by = "fips") %>%  # join nytimes county-covid df with population and msa ares
   select(msa, msa_title, msa_population, date, cases, deaths) %>% 
+  # set date range
   filter(date >= lubridate::as_date("2020-01-24") &
@@
   mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>% 
-  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))
+  # rollmean function from zoo package to help calculate rolling mean, k = no. of days
+  mutate(new_cases_mean = zoo::rollmean(new_cases, k = 7, fill = NA))
 
@@
   mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
-  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))
+  # rollmean function from zoo package to help calculate rolling mean, k = no. of days
+  mutate(new_deaths_mean = zoo::rollmean(new_deaths, k = 7, fill = NA))
 
@@
   left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
-  arrange(desc(msa_population))
+  arrange(desc(msa_population)) # arrange by msa with highest population
 
@@
 msa_covid_sum <- msa_covid %>% 
-  group_by(msa, msa_title, msa_population) %>%
-  mutate(total_cases = sum(new_cases)) %>% 
-  mutate(total_deaths = sum(new_deaths)) #%>%
-  arrange(desc(total_cases)) %>% 
-  select(msa, msa_title, msa_population, total_cases, total_deaths) %>% 
-  distinct() %>% # remove repeated values 
-  na.omit()
-  # drop_na(msa) # remove rows with NA in msa column
-
-msa_covid_sum <- msa_covid %>% 
   filter(date == max(msa_covid$date)) %>% 
-  select(msa, msa_title, msa_population, date, cases_sum, deaths_sum) #%>% 
-  # distinct() #%>%
-  # drop_na(msa)
-
-View(msa_covid_sum)
-
-table(msa_covid$msa_title)
+  select(msa, msa_title, msa_population, date, cases_sum, deaths_sum) %>% 
+  distinct() %>%
+  drop_na(msa)
 
@@
 # similarly, calculate rolling daily 7-day average separately for cases
-nyt_state_cases <- nyt %>% 
+covid_state_cases <- covid %>% # nyt dataset 
   group_by(state, date) %>%
@@
 # rolling daily 7-day average for deaths
-nyt_state_deaths <- nyt %>% 
+covid_state_deaths <- covid %>% 
   group_by(state, date) %>%
@@
 # combined total figures
-state_covid <- nyt_state_cases %>% 
-  left_join(nyt_state_deaths, by = c("state", "date")) %>%
+state_covid <- covid_state_cases %>% 
+  left_join(covid_state_deaths, by = c("state", "date")) %>%
   filter(date >= lubridate::as_date("2020-01-24") & 
@@
 
-# -------------------- FUNCTION TO MAKE A PLOT FOR ONE STATE  ------------------- #
+# ------------------------------------------- #
+# ----------- GETTING CITIES ---------------- # 
+# ------------------------------------------- #
 
+# cities are actually smaller than counties, e.g. New York City is part of New York County
+# some are incorporated, and some would have mayors 
+# this is useful data also to get labels and plot points
+
+head(city)
+```
+
+# Visualizations
+
+## Covid Line Graph (7-day average) per State
+
+```{r}
+
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\\/\/\/ #
+# *************** FUNCTION: PLOT DAILY COVID ONE STATE GRAPH ********************** # 
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\\/\/\/ #
+
+# plot line graph of daily cases for state and metropolitan area
+
+# choose state name
 stateData <- function(state_name) {
@@
 
+# choose metropolitan state area ID 
+# find them here: head(msa_covid_sum, 10)
 metroData <- function(msa_id) {
@@
 
+# line graph
 covidCasesPlot <- function(state_data, state_name, metro_data) {
@@
 
-`
+### Examples
+
 ```{r}
@@
 
-# --------------------- TO MAKE A PLOT FOR ONE STATE -------------------------- #
+# ----------------- EG TO MAKE A DAILY CASES/ DEATHS PLOT FOR ONE STATE ------------------- #
 
@@
 
+## Geo Mapping Visualizations
+
+### County Cases
+
 ```{r}
@@
 
-# mapping packages needed 
-library(leaflet)
-library(maptools)
-library(rgdal)
+library(sf)
 
-# interactive visualization
-library(plotly)
-
-library(sf)
-library(rnaturalearth)
-library(rnaturalearthdata)
-
-library(tigris)
 library(usmap) # to get US Map Data
@@
 library(jsonlite) # from FIPS -> Geo (lat, long)
+
+library(cowplot)
+library(gridExtra) # for grid plot
 
@@
 
+# remove non-mainland US parts 
 outside_mainland <- c("02", "15", "72", "2", "3")
@@
 
-county_map_NY <- county_map %>% 
-  filter(STATEFP == 36) # remove alaska and hawaii
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\\/\/\/ #
+# *************** FUNCTION: PLOT DAILY COVID ONE STATE GRAPH ********************** # 
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\\/\/\/ #
 
-# # plot for states 
-# ggplot(data = county_map_NY) +
-#   geom_sf(aes(fill = cases)) + 
-#   scale_fill_viridis_c(option = "magma", direction = -1) +
-# 
-# 
-# names(county_map_US)
-# 
-# fips(state = "New York")
-
-# PLOT FUNCTION 
 # county state map 
@@
     geom_sf(aes(fill = cases)) + 
-  scale_fill_viridis_c(option = "magma", direction = -1)
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = paste(state_name))
 
@@
 
-W <- county_state_map("Wisconsin")
+# e.g. 
+TX <- county_state_map("Texas")
+CA <- county_state_map("California")
+FL <- county_state_map("Florida")
+NY <- county_state_map("New York")
+IL <- county_state_map("Illinois")
+GA <- county_state_map("Georgia")
 
-range(county_map_US$cases)
+# TO DO: 
+# Facet it
+
+plot_grid(TX, CA, FL, NY, IL, GA, ncol = 2)
+
+ggsave("county-map-cases.pdf")
 
 ```
+### Metropolitan Cases
+
+```{r}
+
+# in progress
+
+# limit to state 
+# view by state, won't be able to plot all 
+# check out why there aren't all 31 MSAs
+
+msa_map <- county_map_US %>% 
+  inner_join(top_census_msa, by = "fips")
+
+View(msa_map)
+
+names(top_census_msa)
+
+dim(top_census_msa)
+
+  ggplot(data = msa_map) +
+    geom_sf(aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1)
+
+```

### 2020-11-12 KPMG Team/Visualizations/Louisa/Metropolitan Areas/msa_county_state_cities.Rmd

--- 2020-11-12 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
@@
 # list of top msas by population, if we want
-top_msa_popn <- census_msa_merge %>%
-  select(msa, msa_title, msa_population) %>%
-  arrange(desc(msa_population)) %>%
-  distinct() %>%
-  na.omit() %>%
-  data.frame() %>%
-  slice(1:12)
+# top_msa_popn <- census_msa_merge %>%
+#   select(msa, msa_title, msa_population) %>%
+#   arrange(desc(msa_population)) %>%
+#   distinct() %>%
+#   na.omit() %>%
+#   data.frame() %>%
+#   slice(1:12)
 
-top_census_msa <- census_msa_merge %>%
-  semi_join(top_msa_popn, by="msa")
+# census_msa <- census_msa_merge %>%
+#   semi_join(top_msa_popn, by="msa")
 
@@
 
-### County Cases
-
 ```{r}
@@
 
+# # mapping packages needed 
+# library(leaflet)
+# library(maptools)
+# library(rgdal)
+# 
+# # interactive visualization
+# library(plotly)
+
 library(sf)
+# library(rnaturalearth)
+# library(rnaturalearthdata)
 
+# library(tigris)
 library(usmap) # to get US Map Data
@@
 library(jsonlite) # from FIPS -> Geo (lat, long)
-
-library(cowplot)
-library(gridExtra) # for grid plot
 
@@
   geom_sf(aes(fill = cases)) + 
-    scale_fill_viridis_c(option = "magma", direction = -1) + 
-    labs(title = paste(state_name))
+  scale_fill_viridis_c(option = "magma", direction = -1)
 
@@
 # e.g. 
-TX <- county_state_map("Texas")
-CA <- county_state_map("California")
-FL <- county_state_map("Florida")
-NY <- county_state_map("New York")
-IL <- county_state_map("Illinois")
-GA <- county_state_map("Georgia")
-
-# TO DO: 
-# Facet it
-
-plot_grid(TX, CA, FL, NY, IL, GA, ncol = 2)
-
-ggsave("county-map-cases.pdf")
+W <- county_state_map("Wisconsin")
 
 ```
-### Metropolitan Cases
-
-```{r}
-
-# in progress
-
-# limit to state 
-# view by state, won't be able to plot all 
-# check out why there aren't all 31 MSAs
-
-msa_map <- county_map_US %>% 
-  inner_join(top_census_msa, by = "fips")
-
-View(msa_map)
-
-names(top_census_msa)
-
-dim(top_census_msa)
-
-  ggplot(data = msa_map) +
-    geom_sf(aes(fill = cases)) + 
-    scale_fill_viridis_c(option = "magma", direction = -1)
-
-```

### 2020-11-12 EPA_Census Team/Bband Tables Simp/Broadband Tables.Rmd
Add Simplified maps and Bband Tables

---
title: "Broadband Tables"
author: "Gretchen Streett"
date: "11/12/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(DT)

#Import Residential Broadband Data 
MI_bband_res <- read.csv("MI_bband_res.csv")

#Import Commercial Broadband Data
MI_bband_com <- read.csv("MI_bband_com.csv")
```

```{r Residential Broadband Table}
#Residential Table Function
databycounty <- function(input){
df_display <- MI_bband_res %>% filter(County == input)
df_display <- df_display %>% group_by(ProviderName, `Technology.of.Transmission`) %>% summarize("Blks" = n(), "MaxDown" = max(MaxAdDown), "MaxUp" = max(MaxAdUp))
df_display <- df_display %>% arrange(desc(Blks))
colnames(df_display) <- c("Provider Name", "Technology", "Blocks", "MaxDown", "MaxUp")
datatable(df_display)
}

#Call function (takes County name as character string, i.e. "Manistee")
databycounty("Manistee")
```

```{r Commercial Broadband Table}
# Commercial Table Function 
bizdatabycounty <- function(input){
df_display <- MI_bband_com %>% filter(County == input)
df_display <- df_display %>% group_by(ProviderName, `Technology.of.Transmission`) %>% summarize("Blks" = n(), "MaxDown" = max(MaxCIRDown), "MaxUp" = max(MaxCIRUp))
df_display <- df_display %>% arrange(desc(Blks))
colnames(df_display) <- c("Provider Name", "Technology", "Blocks", "MaxDown", "MaxUp")
datatable(df_display)
}

#Call function
bizdatabycounty("Benzie")
```

### 2020-11-12 EPA_Census Team/Simplified Maps/Simplified Maps.Rmd
Add Simplified maps and Bband Tables

---
title: "Simplified Maps"
author: "Gretchen Streett"
date: "11/12/2020"
output: html_document
---

``` {r}
library(tidyverse)
library(tmap)
library(ggplot2)
library(sf)
library(ggthemes)
library(viridis)

MIdata <- read.csv("MIbbandwalkdata")
```

```{R Walkability Map}
#Function
walkbycounty <- function(input){
subsetdata <- MIdata %>% filter(V2 == input)
ggplot() + geom_polygon(data = subsetdata, aes(x=long, y=lat, group=group, fill=NatWalkInd), color = "black") +
  scale_fill_gradientn(colours=c("royalblue", "green"), 
           name="Walkability") + theme_map()
}

#call function (Takes County Name as Character String, i.e. "Manistee")
walkbycounty("Emmet")
```

```{R Broadband Map}
# Mapping Function:
bbandbycounty <- function(input){
subsetdata <- MIdata %>% filter(V2 == input)
ggplot() + geom_polygon(data = subsetdata, aes(x=long, y=lat, group=group, fill=GroupProvideNum), color = "black") +
  scale_fill_viridis(option = "A", name="Number of \n Broadband Providers") + theme_map()
}

#Call Function (Takes County Name as a character string, I.e. "Manistee")
bbandbycounty("Manistee")
```

### 2020-11-13 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
More Visualizations
Metropolitan Areas within States

@@
 library(ggplot2)
+library(ggthemes)
+library(extrafonts)
 # library(hrbrthemes) 
@@
 
+# set working directory to source file location
+
 # metropolitan state areas delineation from file taken off Census Bureau
-data <- read_csv("data/msa.csv")
+msa <- read_csv("data/msa.csv")
 
@@
 # select only msa-related information 
-data_msa <- data %>% 
+data_msa <- msa %>% 
   mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
@@
 # only need county to merge both datasets, will keep just metro areas (csa)
-msa_covid_merge <- census_msa %>% 
-  left_join(covid, by = "fips") %>%  # join nytimes county-covid df with population and msa ares
+msa_county_covid_merge <- census_msa %>% 
+  left_join(covid, by = c("fips", "state")) %>%
+  select(!B01003_001M) %>%
+  drop_na(msa)
+
+# ###########
+# sum of county, showing msa as well 
+msa_county_sum <- msa_county_covid_merge %>%
+  filter(date == max(date))
+# ###########
+
+# join nytimes county-covid df with population and msa areas (remove county)
+msa_covid_merge <- msa_county_covid_merge %>% 
   select(msa, msa_title, msa_population, date, cases, deaths) %>% 
@@
 
+head(msa_county_sum)
+
 # ----------------------------------------------------------------- #
@@
 head(state_covid, 10)
+
+# final count of msa cases, in order to get ranking
+state_sum <- state_covid %>% 
+  filter(date == max(state_covid$date)) %>% 
+  select(state, date, cases_sum, deaths_sum) %>% 
+  distinct()
+
+```
+
+```{r}
+
+# --------------------------------------------------------- #
+# ----------- GETTING STATE & MSA TOGETHER ---------------- # 
+# --------------------------------------------------------- #
+
+# making both datasets bindable 
+msa_covid_sum_merge <- msa_covid_sum %>% 
+  rename(msa_state = "msa_title") %>% 
+  ungroup() %>%
+  select(msa_state, date, cases_sum, deaths_sum)
+
+state_covid_sum_merge <- state_covid_sum %>% 
+  rename(msa_state = "state") %>% 
+  select(msa_state, date, cases_sum, deaths_sum)
+
+# binding row  
+msa_state_covid_sum <- msa_covid_sum_merge %>%
+  bind_rows(state_covid_sum_merge)
 
@@
 # https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html
-county_lines <- st_read("county_shape/cb_2018_us_county_500k.shp")
+county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")
+state_lines <- st_read("data/state_shape/cb_2018_us_state_500k.shp")
 
+names(state_sum)
+head(state_lines)
+
+# merge county 
 # add FIPS as a variable from State Fips and County Fips
@@
   mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))
-
 # add lat long attributes to county dataset          
@@
 
+# merge MSA
+msa_geo <- merge(county_lines, msa_county_sum, by = "fips")
+msa_map <- msa_geo %>% st_transform('+proj=longlat +datum=WGS84')
+
+# merge states and state lines
+state_lines <- state_lines %>% 
+  rename(state = NAME)
+state_geo <- merge(state_lines, state_sum, by = "state")
+state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')
+
 # remove non-mainland US parts 
-outside_mainland <- c("02", "15", "72", "2", "3")
-islands <- c("Virgin Islands", "Northern Mariana Islands")
+# outside_mainland <- c("02", "15", "72", "2", "3")
+outside_mainland <- c("Alaska", 
+                      "Hawaii", 
+                      "Puerto Rico", 
+                      "Virgin Islands", 
+                      "Northern Mariana Islands",
+                      "Guam")
 
 county_map_US <- county_map %>% 
-  filter(!STATEFP %in% outside_mainland) %>% 
-  filter(!state %in% islands) # remove alaska and hawaii
+  filter(!state %in% outside_mainland)
+
+msa_map_US <- msa_map %>% 
+  filter(!state %in% outside_mainland)
+
+state_map_US <- state_map %>% 
+  filter(!state %in% outside_mainland) %>% 
+  rename(cases = cases_sum) %>% 
+  rename(deaths = deaths_sum)
+
+names(state_map_US)
 
@@
   scale_fill_viridis_c(option = "magma", direction = 1)
+
+ggplot(data = msa_map_US) +
+  geom_sf(aes(fill = cases)) + 
+  scale_fill_viridis_c(option = "magma", direction = 1)
+
+ggplot(data = state_map_US) +
+  geom_sf(aes(fill = cases)) + 
+  scale_fill_viridis_c(option = "magma", direction = -1) + 
+  theme_void()
 
@@
 # TO DO:
+
+ fips_code <- fips(state = "New York") 
 # Facet it
@@
 
-ggsave("county-map-cases.pdf")
+# msa state map 
+msa_state_map <- function(msa) {
 
+  msa_map_state <- msa_map %>% 
+  filter(msa == msa)
+
+  ggplot(data = msa_map_state) +
+    geom_sf(aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = msa_map_state$NAME)
+
+}
+
+msa_state_map(40484)
+
+head(msa_county_sum)
+
+msa_map_NWNJ <- msa_map %>% 
+  filter(msa == 35084)
+
+31084
+
+msa_map_LA <- msa_map %>% 
+  filter(msa == 41884)
+
+msa_map_SF <- msa_map %>% 
+  filter(msa == 31084)
+
+msa_map_OC <- msa_map %>% 
+  filter(msa == 11244)
+
+msa_map_OAK <- msa_map %>% 
+  filter(msa == 36084)
+
+msa_map_RA <- msa_map %>% 
+  filter(msa == 42034)
+
+state_map_CA <- state_map_US %>% 
+  filter(state == "California")
+
+state_map_CA_region <- state_map_CA
+
+# state_map_CA <- state_map_CA %>% 
+#   mutate(region_cases = cases - msa_map_LA$cases
+#          
+state_map_CA_region$cases <- state_map_CA$cases -
+  sum(msa_map_LA$cases) - 
+  msa_map_SF$cases - 
+  msa_map_OC$cases - 
+  sum(msa_map_OAK$cases) - 
+  msa_map_RA$cases
+
+ggplot() +
+    geom_sf(data = state_map_CA_region, aes(fill = cases)) + 
+    geom_sf(data = msa_map_LA, aes(fill = cases)) + 
+    geom_sf(data = msa_map_SF, aes(fill = cases)) + 
+    geom_sf(data = msa_map_OC, aes(fill = cases)) + 
+    geom_sf(data = msa_map_OAK, aes(fill = cases)) + 
+    geom_sf(data = msa_map_RA, aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = state_map_CA$state,
+         subtitle = "Metropolitan Statistical Areas") + 
+    theme_void()
+
+ggplot() +
+    geom_sf(data = state_map_CA, aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(subtitle = state_map_CA$state) + 
+    theme_void()
+
+names(msa_map_NWNJ)
+
+### TEXAS ###
+
+state_map_TX <- state_map_US %>% 
+  filter(state == "Texas")
+
+texas_msa <- msa_county_sum %>% 
+  select("state", "msa_title", "msa") %>%
+  filter(state == "Texas") %>%
+  distinct()
+
+texas_msa_code <- pull(texas_msa, msa) # get msa codes into a vector
+
+msa_map_TX <- msa_map %>% 
+  filter(msa %in% texas_msa_code)
+
+state_map_TX$region_cases <- state_map_TX$cases - sum(msa_map_TX$cases)
+
+ggplot() +
+    geom_sf(data = state_map_TX, aes(fill = region_cases)) +
+    geom_sf(data = msa_map_TX, aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = state_map_TX$state,
+         subtitle = "Metropolitan Statistical Areas") + 
+    theme_void()
+
+head(msa_map_TX)
+
+### FUNCTION ###
+
+msa_state_map_US <- function(state_name) {
+
+  # e.g. filter to get state's map
+  state_map <- state_map_US %>% 
+    filter(state == state_name)
+
+  state_msa <- msa_county_sum %>% 
+    select("state", "msa_title", "msa") %>%
+    filter(state == state_name) %>%
+    distinct()
+
+  # get msas from the state 
+  state_msa_code <- pull(state_msa, msa) # get msa codes into a vector
+
+  msa_map_state <- msa_map %>% 
+    filter(msa %in% state_msa_code)
+
+  # get regional numbers
+  state_map$region_cases <- state_map$cases - sum(msa_map_state$cases)
+
+  #   # get two-letter e.g. NY
+  state_code <- state_map$STUSPS
+
+  # df <- list()
+  # 
+  # df$state_map <- state_map 
+  # df$msa_map_state <- msa_map_state
+
+#   return(df)
+# }  
+  # plot
+  ggplot() +
+    geom_sf(data = state_map, aes(fill = region_cases)) +
+    geom_sf(data = msa_map_state, aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = state_map$state,
+         subtitle = "Metropolitan Statistical Areas") + 
+    theme_void()
+
+}
+
+35614 # nyc is missing
+
+df_ny <- msa_state_map_US("New York")
+
+View(df_ny$msa_map_state)
+
+msa_map %>% 
+  filter(state == "New York")
+
+plot_grid(TX, CA, FL, NY, IL, GA, ncol = 2)
+
+IL_msa <- msa_state_map_US("Illinois")
+CA_msa <- msa_state_map_US("California")
+TX_msa <- msa_state_map_US("Texas")
+NJ_msa <- msa_state_map_US("New Jersey")
+MA_msa <- msa_state_map_US("Massachusetts")
+FL_msa <- msa_state_map_US("Florida")
+
+plot_grid(TX_msa, CA_msa, FL_msa, IL_msa, MA_msa, NJ_msa, ncol = 2)
+
+# ggsave("metro-state.png")
 ```

### 2020-11-13 KPMG Team/Dashboard/visualization_scripts/cities.Rmd

@@
 
-```{r}
 ```{r}
@@
 
-#nytimes county-level cases & deaths data from online link 
+# nytimes county-level cases & deaths data from online link 
+# a few less steps of data manipulation required, but essentially the same from JHU.
 nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
@@
 # https://simplemaps.com/data/us-cities license 
-city <- read_csv("uscities.csv")
+city_data <- read_csv("data/uscities.csv")
 
-View(city)
-names(city)
 ```
@@
 ```{r}
-names(city)
 
@@
 # CITY DATASET 
-data_city <- city %>% 
-  rename(fips = county_fips) %>% 
-  group_by(city) %>% 
-  mutate(city_population = sum(population))
+# cities are actually smaller than counties, e.g. New York City is part of New York County
 
-View(nyt)
+# 
+# city_data %>% 
+#   data.frame() %>% 
+#   arrange(desc(population)) %>% 
+#   slice(1:300)
+# 
+# View(city_data_2)
+# 
+# View(city_data)
+#   
+# 
+# city_data_2 <- city_data %>% 
+#   data.frame() %>%
+#   mutate(fips = county_fips) %>% 
+#   # mutate(state = state_name) %>% 
+#   # mutate(state_code = state_id) %>% 
+#   # mutate(county = county_name) %>% 
+#   # mutate(county_popn = population) %>%
+#   group_by(state_name, city) %>% 
+#   mutate(city_population = sum(population)) %>% 
+#   select(city, state_name, state_id, city_population, fips, county_name, population)
+
+# 
+# city_state <- city_data_renamed %>% 
+#   select(city, state_id, state_name) %>% 
+#   distinct()
+# 
+# top300 <- city_data_renamed %>%
+#   data.frame() %>%
+#   select(city, city_population) %>%
+#   arrange(desc(city_population)) %>% # %>% 
+#   distinct() %>%
+#   slice(1:300)
+# 
+# city_state_300 <- top300 %>%
+#   left_join(data.frame(city_data_renamed), by = c("city", "city_population"))
+# 
+# View(city_state_300)
+# 
+# city_data_renamed
+# 
+# city_df <- top300
+# View(top300)
+# 
+# 
+# city_data_2 %>% 
+#   arrange(desc(city_population))
+
+View(city_data_2)
+```
+
+```{r}
 
@@
 # only need county to merge both datasets, will keep just city areas 
-city_covid_merge <- data_city %>% 
+city_covid_merge <- city_df %>% 
   left_join(nyt, by = "fips") %>% 
   # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
-  select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
+  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
+    select(city, state_id, state_name, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
   filter(date >= lubridate::as_date("2020-01-24") & 
@@
 city_covid_cases <- data.frame(city_covid_merge) %>% 
-  group_by(city, date) %>% 
-  dplyr::mutate(cases_sum = max(cases)) %>% # choose max cases for each date because some dates report more than once
+  group_by(city, city_population, date) %>% 
+  summarise(cases_sum = sum(cases, na.rm = TRUE)) %>%
   mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
@@
 
-View(city_covid_cases)
 # same as above, but for deaths
-msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
-  group_by(msa, date) %>% 
-  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
+city_covid_deaths <- data.frame(city_covid_merge) %>% 
+  group_by(city, city_population, date) %>% 
+  summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
   mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
@@
 # join deaths and cases numbers together (not sure if I can do it in one?)
-# contains
-msa_covid <- msa_covid_merge %>% 
-  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
-  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
-  arrange(desc(msa_population))
+city_covid <- city_covid_merge %>% 
+  left_join(city_covid_cases, by = c("city", "date")) %>% 
+  left_join(city_covid_deaths, by = c("city", "date")) %>% 
+  select(city, state_id, state_name, city_population, 
+       date, cases, new_cases, new_cases_mean, cases_sum,
+       deaths, new_deaths, new_deaths_mean, deaths_sum) %>%
+  arrange(desc(city_population))
+
+city_covid_sum <- city_covid %>% 
+  filter(date == max(city_covid$date)) %>% 
+  select(city, state_id, state_name, city_population, 
+       date, cases, new_cases, new_cases_mean, cases_sum, 
+       deaths, new_deaths, new_deaths_mean, deaths_sum)
+
+city_only <- data_city %>% 
+  select(city, state_id, state_name, city_population) %>% 
+  distinct()
+
+city_covid_2 <- city_covid_cases %>%
+  left_join(city_only, by = c("city", "city_population"))
+
+View(city_covid_2)
+
+View(data_city)
+View(city_covid_merge)
+View(city_covid_deaths)
+
+names(data_city)
+names(city_covid)
+
+View(city_only)
+
+View(city)
+```
+
+```{r}
 

### 2020-11-13 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
Exported Visualizations for Presentation

@@
 
-state_covid_sum_merge <- state_covid_sum %>% 
+state_covid_sum_merge <- state_sum %>% 
   rename(msa_state = "state") %>% 
@@
 ```{r}
-
 # Plot State Map with Metro Areas
@@
 
+ggplot() +
+  geom_sf(data = state_map_US, aes(fill = cases)) + 
+  geom_sf(data = msa_map_US, aes(fill = cases)) + 
+  scale_fill_viridis_c(option = "magma", direction = -1) + 
+  theme_void()
+
 ggplot(data = state_map_US) +
@@
 
+ggsave("state-msa-us.png", width = 8, height = 5, dpi = 300)
+```
+
+```{r}
 # \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\\/\/\/ #
@@
     scale_fill_viridis_c(option = "magma", direction = -1) + 
-    labs(title = paste(state_name))
+    labs(title = paste(state_name)) +
+    theme_void()
 
@@
 FL <- county_state_map("Florida")
-NY <- county_state_map("New York")
+MA <- county_state_map("Massachusetts")
+# NY <- county_state_map("New York")
+NJ <- county_state_map("New Jersey")
 IL <- county_state_map("Illinois")
-GA <- county_state_map("Georgia")
+# GA <- county_state_map("Georgia")
 
@@
 
-plot_grid(TX, CA, FL, NY, IL, GA, ncol = 2)
+plot_grid(TX, CA, FL, IL, MA, NJ, ncol = 3)
+```
 
+```{r}
 # msa state map 
@@
     scale_fill_viridis_c(option = "magma", direction = -1) + 
-    labs(title = state_map$state,
-         subtitle = "Metropolitan Statistical Areas") + 
+    labs(title = state_map$state) +
+         # subtitle = "Metropolitan Statistical Areas") + 
     theme_void()
@@
 
-35614 # nyc is missing
+# 35614 # nyc is missing
 
-df_ny <- msa_state_map_US("New York")
-
-View(df_ny$msa_map_state)
-
-msa_map %>% 
-  filter(state == "New York")
-
-plot_grid(TX, CA, FL, NY, IL, GA, ncol = 2)
+# df_ny <- msa_state_map_US("New York")
+# 
+# View(df_ny$msa_map_state)
+# 
+# msa_map %>% 
+#   filter(state == "New York")
 
@@
 
-plot_grid(TX_msa, CA_msa, FL_msa, IL_msa, MA_msa, NJ_msa, ncol = 2)
+plot_grid(TX_msa, CA_msa, FL_msa, IL_msa, MA_msa, NJ_msa, ncol = 3)
 
-# ggsave("metro-state.png")
+# ggsave("county-state.png", width = 8, height = 5, dpi = 300)
 ```

### 2020-11-17 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
figures per pax

@@
 library(ggthemes)
-library(extrafonts)
-# library(hrbrthemes) 
-# library(geofacet)
+library(showtext)
+
+showtext_auto()
+
+font_add(family = "raleway", regular = "C:/Users/louis/Dropbox/Stuff/Fonts/Raleway-Regular.ttf")
+
+```
+
+```{r cache = TRUE}
 
@@
 
-# not much to do
-county_sum <- covid %>%
+county_sum_total <- covid %>%
   filter(date == max(covid$date)) # latest
 
-# Preview 
-head(county_sum, 50)
+# wave 2 
+county_sum_wave_2_start <- covid %>%
+  filter(date == "2020-06-15") # %>% # latest
 
+county_sum_w2 <- county_sum_total %>% 
+  inner_join(county_sum_wave_2_start, by = c("fips", "county")) %>% 
+  mutate(cases = cases.x - cases.y) %>%
+  mutate(deaths = deaths.x - deaths.y) %>% 
+  rename(start_date = date.y) %>% 
+  rename(end_date = date.x) %>%
+  select(fips, county, population, start_date, end_date, cases, deaths)
+
+# DATA WITH POPULATION
+county_census <-census %>% 
+  rename(fips = GEOID) %>% 
+  left_join(covid, by = "fips")
+
+county_census_total <- county_census %>% 
+  filter(date == "2020-11-14")
+# idk why max isn't working
+
+# WAVE 2 
+county_census_sum_w2_start <- county_census %>%
+  filter(date == "2020-06-15")
+
+county_census_sum_w2 <- county_census_total %>% 
+  inner_join(county_census_sum_w2_start, by = c("fips", "county", "population", "state")) %>% 
+  mutate(cases = cases.x - cases.y) %>%
+  mutate(deaths = deaths.x - deaths.y) %>% 
+  rename(start_date = date.y) %>% 
+  rename(end_date = date.x) %>%
+  select(fips, county, population, start_date, end_date, cases, deaths)
+
+county_census_sum_w2
+
+View(county_census_sum_w2)
+dim(county_census_sum_w2)
 ```
@@
 
+county_geo_w2 <- merge(county_lines, county_sum_w2, by = "fips")
+county_map_w2 <- county_geo_w2 %>% st_transform('+proj=longlat +datum=WGS84')
+
+# census county, wave 2
+county_census_geo_w2 <- merge(county_lines, county_census_sum_w2, by = "fips") 
+county_census_map_w2 <- county_census_geo_w2 %>% st_transform('+proj=longlat +datum=WGS84')
+
 # merge MSA
@@
 
+county_map_US_w2 <- county_map_w2 %>% 
+  filter(!state %in% outside_mainland)
+
+county_census_map_US_w2 <- county_census_map_w2 %>% 
+  filter(!state %in% outside_mainland)
+
+# ---
+
 msa_map_US <- msa_map %>% 
@@
 names(state_map_US)
+
+# --- 
+# PLOTS
 
@@
 
-ggsave("state-msa-us.png", width = 8, height = 5, dpi = 300)
+# save to image function - last plot 
+# ggsave("state-msa-us.png", width = 8, height = 5, dpi = 300)
 ```
@@
     scale_fill_viridis_c(option = "magma", direction = -1) + 
-    labs(title = paste(state_name)) +
+    labs(title = paste(state_name),
+         subtitle = "Total Cases") +
+    theme_void()
+
+}
+
+# census
+
+county_pax_cases_w2 <- function(state_name) {
+
+  fips_code <- fips(state = state_name)
+
+  county_census_map_state <- county_census_map_US_w2 %>% 
+    filter(STATEFP == fips_code) %>%
+    filter(cases >= 0)
+
+  ggplot(data = county_census_map_state) +
+    geom_sf(aes(fill = cases/population)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = paste(state_name),
+         subtitle = "Cases per Capita after Second Wave") +
+    theme_void() + 
+    theme(text = element_text(family = "raleway", size = 25))
+
+}
+
+county_pax_deaths_w2 <- function(state_name) {
+
+  fips_code <- fips(state = state_name)
+
+  county_census_map_state <- county_census_map_US_w2 %>% 
+    filter(STATEFP == fips_code) %>% 
+    filter(deaths >= 0)
+
+  ggplot(data = county_census_map_state) +
+      geom_sf(aes(fill = deaths/population)) + 
+      scale_fill_viridis_c(option = "viridis", direction = -1) + 
+      labs(title = paste(state_name),
+           subtitle = "Deaths per Capita after Second Wave") +
+      theme_void() + 
+      theme(text = element_text(family = "raleway", size = 25))
+
+}
+
+county_state_map_w2 <- function(state_name) {
+
+  fips_code <- fips(state = state_name)
+
+  county_map_state <- county_map_w2 %>% 
+  filter(STATEFP == fips_code)
+
+  ggplot(data = county_map_state) +
+    geom_sf(aes(fill = cases)) + 
+    scale_fill_viridis_c(option = "magma", direction = -1) + 
+    labs(title = paste(state_name),
+         subtitle = "After Second Wave") +
     theme_void()
@@
 MA <- county_state_map("Massachusetts")
-# NY <- county_state_map("New York")
+NY <- county_state_map("New York")
 NJ <- county_state_map("New Jersey")
 IL <- county_state_map("Illinois")
-# GA <- county_state_map("Georgia")
+GA <- county_state_map("Georgia")
+
+# cases after second wave 
+# e.g. 
+TX_w2 <- county_state_map_w2("Texas")
+CA_w2 <- county_state_map_w2("California")
+FL_w2 <- county_state_map_w2("Florida")
+MA_w2 <- county_state_map_w2("Massachusetts")
+NY_w2 <- county_state_map_w2("New York")
+
+NJ_w2 <- county_state_map_w2("New Jersey")
+IL_w2 <- county_state_map_w2("Illinois")
+GA_w2 <- county_state_map_w2("Georgia")
+
+county_pax_w2
+
+CA_pax_w2_cases <- county_pax_cases_w2("California")
+CA_pax_w2_deaths <- county_pax_deaths_w2("California")
+
+CA_3 <- county_pax_cases_w2("California")
+CA_4 <- county_pax_deaths_w2("California")
+
+NY_3 <- county_pax_cases_w2("New York")
+NY_4 <- county_pax_deaths_w2("New York")
+
+IL_3 <- county_pax_cases_w2("Illinois")
+IL_4 <- county_pax_deaths_w2("Illinois")
+
+FL_3 <- county_pax_cases_w2("Florida")
+FL_4 <- county_pax_deaths_w2("Florida")
+
+WI_3 <- county_pax_cases_w2("Wisconsin")
+WI_4 <- county_pax_deaths_w2("Wisconsin")
+
+ND_3 <- county_pax_cases_w2("North Dakota")
+ND_4 <- county_pax_deaths_w2("North Dakota")
+
+SD_3 <- county_pax_cases_w2("South Dakota")
+SD_4 <- county_pax_deaths_w2("South Dakota")
+
+GA_3 <- county_pax_cases_w2("Georgia")
+GA_4 <- county_pax_deaths_w2("Georgia")
 
@@
 
-plot_grid(TX, CA, FL, IL, MA, NJ, ncol = 3)
+plot_grid(TX, TX_w2, 
+          CA, CA_w2, 
+          FL, FL_w2, 
+          IL, IL_w2, 
+          MA, MA_w2, 
+          NJ, NJ_w2, 
+          NY, NY_w2, 
+          ncol = 2)
+
+plot_grid(FL_3, FL_4, 
+          IL_3, IL_4, 
+          ncol = 2)
+
+plot_grid(NY_3, NY_4, 
+          CA_3, CA_4,
+          WI_3, WI_4,
+          GA_3, GA_4,
+          ncol = 2)
+
+# ggsave("CA-comparison.png", plot_grid(CA, CA_msa), width = 8, height = 5, dpi = 300)
+# ggsave("IL-comparison.png", plot_grid(IL, IL_msa), width = 8, height = 5, dpi = 300)
+# ggsave("MA-comparison.png", plot_grid(MA, MA_msa), width = 8, height = 5, dpi = 300)
+# 
+# ggsave("CA-county-state.png", CA, width = 5, height = 6, dpi = 300)
+# ggsave("IL-county-state.png", IL, width = 5, height = 6, dpi = 300)
+# ggsave("MA-county-state.png", MA, width = 5, height = 6, dpi = 300)
+
+# ggsave("county-state-waves.png", width = 8, height = 25, dpi = 300)
+
+# ggsave("county-pax-w2_NY_CA_WI_GA.png", width = 14, height = 25, dpi = 300)
+# 
+# ggsave("county-pax-w2_FL_IL.png", width = 10, height = 8, dpi = 300)
+
 ```
@@
 
-# ggsave("county-state.png", width = 8, height = 5, dpi = 300)
+# ggsave("CA-metro-state.png", CA_msa, width = 5, height = 6, dpi = 300)
+# ggsave("IL-metro-state.png", IL_msa, width = 5, height = 6, dpi = 300)
+# ggsave("MA-metro-state.png", MA_msa, width = 5, height = 6, dpi = 300)
+# ggsave("metro-state.png", width = 8, height = 5, dpi = 300)
 ```

### 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa_county_state_cities.Rmd
county viz wave 2

@@
   rename(end_date = date.x) %>%
-  select(fips, county, population, start_date, end_date, cases, deaths)
+  select(fips, county, all_of(population), start_date, end_date, cases, deaths)
 
@@
 
+data_msa$fips
+
 # join msa data and census data by GEOID/FIPS, get population (and other demographic) info.
@@
   mutate(msa_population = sum(population))
+
+census_msa_merge$fips
 
@@
 
-library(cowplot)
+# library(cowplot)
 library(gridExtra) # for grid plot

### 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd

---
title: "Visualizations: Metropolitan Areas per Capita, Wave I, II, III (Deaths)"
output: html_notebook 
---

# CONTENTS OF THIS SCRIPT

### Loading Packages 
### Datasets 
### Merging County data to Top 50 populated MSA & Region Dataset (for spatial attributes)
### Adding Spatial Attributes 
### Filtering by waves
### Visualization in Maps

```{r message = FALSE}
# ----------- LOAD PACKAGES ------------- #

library(tidyverse)
library(tidycensus)

# plotting 
library(ggplot2)
library(ggthemes)
library(showtext)

showtext_auto()

font_add(family = "raleway", regular = "C:/Users/louis/Dropbox/Stuff/Fonts/Raleway-Regular.ttf")

```

```{r cache = TRUE}

# ----------- LOADING DATASETS ----------- #

# MSA Things
# from Sydney
msa_index <- read_csv("data/msa/msa_mix_index.csv") # csa and msa with population 
covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split

# metropolitan state areas delineation from file taken off Census Bureau
msa <- read_csv("data/msa.csv") # using this bc FIPS code needs to be 5 digits

# -------------- MSA_INDEX dataset check
# top 50 most populated csa and msas
# msa_index %>% 
#   select("mix_index") %>% 
#   distinct() %>% 
#   count()

# -------------- MSA_SPLIT dataset check
# Regional minus State count 
# CSA/MSA count

# View(msa)
```

# Adding County Data (to match with spatial lines)

```{r}

# need to merge county back in to get the lines

# # rename to get 5-digit fips, select only needed variables, remove NAs
msa_w_fips <- msa %>%
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>% #
  select(fips, `CSA Title`, `County/County Equivalent`, `State Name`)

# # left join index of 50 with fips codes
# msa_index_fips <- msa_index %>% 
#   left_join(msa_w_fips, by = c("CSA Title", "County/County Equivalent", "State Name")) %>% 
#   rename(county = `County/County Equivalent`) %>% 
#   rename(csa = `CSA Title`) %>%
#   rename(state = `State Name`) %>%
#   rename(agg_index = mix_index)

msa_index_2 <- msa_index %>% 
  rename(county = `County/County Equivalent`) %>% 
  rename(csa = `CSA Title`) %>%
  rename(state = `State Name`) %>%
  rename(agg_index = mix_index) %>% 
  rename(fips = FIPS_combined) %>% 
  names()

# anti join to get the counties not in msa_index 
region_fips <- msa_w_fips %>% 
  anti_join(msa_index, by = "CSA Title") %>% 
  rename(county = `County/County Equivalent`) %>%
  mutate(agg_index = `State Name`) %>%
  rename(state = `State Name`) %>%
  rename(csa = `CSA Title`)

# joining fips info with covid dataset (has states + selected msas)
state_msa_fips_join <- covid %>% 
  full_join(msa_index_fips, by = "agg_index")# %>%

# keeping only msas and counties, since they have fips
msa_covid_only <- state_msa_fips_join %>%
  drop_na(fips)

# keeping only states, don't have counties
state_covid_only <- state_msa_fips_join %>% 
  anti_join(msa_covid_only) %>% 
  select(-fips, -mix_index_population, -FIPS_population, -FIPS_combined, -`Central/Outlying County`, -state, -county, -csa, -`Metropolitan Division Title`)

names(msa_region)

region_fips_join <- msa_region %>% 
  full_join(region_fips, by = "agg_index") %>%
  drop_na(fips)

View(msa_covid_only)

View(region_fips_join)
View(covid)

final_daily <- msa_fips_join %>% # or final total 
  bind_rows(region_fips_join) %>% 
  select(index, Date, state, agg_index, fips, csa, county, population, 
         Confirmed, Deaths, new_confirmed, new_deaths, 
         new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
         per_100k_new_confirmed_7, per_100k_new_deaths_7)

names(final_daily)
# write.csv(final_daily, "data/final-50-county-daily.csv")

```

# Adding spatial attributes

```{r cache = TRUE}

# Adding spatial attributes

library(sf)
library(jsonlite) # from FIPS -> Geo (lat, long)
library(gridExtra) # for grid plot

county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")

# merge county 
# add FIPS as a variable from State Fips and County Fips
county_lines <- county_lines %>% 
  mutate(fips = paste(STATEFP, COUNTYFP, sep = ""))

# add lat long attributes to county dataset          
county_geo <- merge(county_lines, final_daily, by = "fips")
county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')

map_daily <- county_map # daily covid map

```

# Wave/Dates Filter  
### For total cases

```{r}
# library(spdplyr) # spatial version of dplyr

# Dates of Waves

wave_2 <- as.Date("2020-06-15") # start of wave 2
wave_3 <- as.Date("2020-10-15") # start of wave 3

# all cases map 
map_total <- county_map %>% 
  filter(Date == max(as.Date(county_map$Date), na.rm = TRUE)) %>% # need to remove NA
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 start date 
map_w2_start <- map_daily %>%
  filter(Date == "2020-06-15") %>% 
  rename(state_fips = STATEFP) %>%
  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)

# wave 2 map 
map_w2 <- map_total %>% 
  st_join(map_w2_start, by = "fips") %>% 
  mutate(Confirmed = Confirmed.x - Confirmed.y) %>% 
  mutate(Deaths = Deaths.x - Deaths.y) %>% 
  rename(start_date = Date.y) %>% 
  rename(end_date = Date.x) %>% 
  rename(fips = fips.x) %>% 
  rename(county = NAME.x) %>% 
  rename(population = population.x) %>% 
  rename(csa = csa.x) %>% 
  rename(agg_index = agg_index.x) %>% 
  rename(state = state.x) %>%
  rename(index = index.x) %>%
  rename(county_fips = COUNTYFP.x) %>%
  rename(state_fips = state_fips.x) %>%
  select(fips, index, agg_index, state, county, csa, population, Confirmed, Deaths, start_date, end_date, state_fips, county_fips, geometry, LSAD.x, ALAND.x, AWATER.x)

```

# Visualizations (finally)

```{r}
library(usmap) # to get US Map Data

names(map_w2)
fips("New York")
#state_name,

# county state map 
mapping <- function(map_type, pax) {
  # 
  # fips_code <- fips(state = state_name) 
  # 
  # map <- map_type %>% 
  #   filter(state_fips == fips_code) %>% 
  #   filter(Deaths >= 0)
  #

  map <- map_type

  ggplot(data = map) +
    geom_sf(aes(fill = Deaths, color = Deaths)) + 
    scale_fill_viridis_c(option = "viridis", direction = -1) + 
    labs(title = "something", #paste(state),
         subtitle = "Total Deaths") +
    theme_void() + 
    theme(text = element_text(family = "raleway"))

  # if (pax = FALSE) {
  #   
  #   map <- 
  #   
  # } else if (pax = TRUE) {
  #   
  #   
  #

}

dim(map_w2)

length(map_total$Deaths)
length(map_w2$Deaths)

mapping(map_w2)

```

### 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
US States per Capita - 3 waves
Counties next

@@
 
-names(msa_region)
+# problematic with states without fips 
+# region_fips_join <- msa_region %>% 
+#   full_join(region_fips, by = "agg_index") %>%
+#   drop_na(fips)
 
-region_fips_join <- msa_region %>% 
-  full_join(region_fips, by = "agg_index") %>%
-  drop_na(fips)
-
-View(msa_covid_only)
-
-View(region_fips_join)
-View(covid)
-
-final_daily <- msa_fips_join %>% # or final total 
-  bind_rows(region_fips_join) %>% 
-  select(index, Date, state, agg_index, fips, csa, county, population, 
-         Confirmed, Deaths, new_confirmed, new_deaths, 
-         new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
-         per_100k_new_confirmed_7, per_100k_new_deaths_7)
-
-names(final_daily)
+# final_daily <- msa_fips_join %>% # or final total 
+#   bind_rows(region_fips_join) %>% 
+#   select(index, Date, state, agg_index, fips, csa, county, population, 
+#          Confirmed, Deaths, new_confirmed, new_deaths, 
+#          new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
+#          per_100k_new_confirmed_7, per_100k_new_deaths_7)
+#          
+# names(final_daily)
 # write.csv(final_daily, "data/final-50-county-daily.csv")
 
+```
+
+# Getting Wave Count
+
+```{r}
+
+wave_2 <- as.Date("2020-06-15") # start of wave 2
+wave_3 <- as.Date("2020-10-15") # start of wave 3
+
+state_covid_only <- state_covid_only %>% 
+  rename("NAME" = "agg_index") %>%
+  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))
+
+state_total <- state_covid_only %>% 
+  filter(Date == max(as.Date(state_covid_only$Date), na.rm = TRUE)) %>% # need to remove NA
+  mutate(Date_Total = Date)
+
+state_w2_start <- state_covid_only %>% 
+  filter(Date == max(wave_2)) %>%
+  mutate(Date_W2 = Date)
+
+state_w3_start <- state_covid_only %>% 
+  filter(Date == max(wave_3)) %>% 
+  mutate(Date_W3 = Date)
+
+state_all <- state_total %>% 
+  # joining wave 2 and beyond numbers
+  left_join(state_w2_start, by = "NAME") %>% 
+  rename(Confirmed_total = Confirmed.x) %>%
+  rename(Deaths_total = Deaths.x) %>%
+  # wave 1 only %>% 
+  mutate(Confirmed_W1 = Confirmed.y) %>% 
+  mutate(Deaths_W1 = Deaths.y) %>%
+  # wave 2 and beyond column 
+  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed.y) %>% 
+  mutate(Deaths_W2_plus = Deaths_total - Deaths.y) %>% 
+  # joining wave 3 numbers 
+  left_join(state_w3_start, by = "NAME") %>%
+  # only wave 2 
+  mutate(Deaths_W2_only = Deaths - Deaths_W2_plus) %>% 
+  mutate(Confirmed_W2_only = Confirmed - Confirmed_W2_plus) %>% 
+  # wave 3 and beyond
+  mutate(Deaths_W3 = Deaths_total - Deaths) %>% 
+  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>%
+  # tidy up, selecting only relevant variables
+  select(index, NAME, population, 
+         Date_Total, Confirmed_total, Deaths_total, 
+         Confirmed_W1, Deaths_W1,
+         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
+         Date_W3, Confirmed_W3, Deaths_W3)
+
+View(state_all)
 ```
@@
 county_lines <- st_read("data/county_shape/cb_2018_us_county_500k.shp")
+state_lines <- st_read("data/state_shape/cb_2018_us_state_500k.shp")
 
@@
 # add lat long attributes to county dataset          
-county_geo <- merge(county_lines, final_daily, by = "fips")
+county_geo <- merge(county_lines, msa_covid_only, by = "fips")
 county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')
 
-map_daily <- county_map # daily covid map
+# merge state 
+# add lat long attributes to state dataset
 
-```
-
-# Wave/Dates Filter  
-### For total cases
-
-```{r}
-# library(spdplyr) # spatial version of dplyr
-
-# Dates of Waves
-
-wave_2 <- as.Date("2020-06-15") # start of wave 2
-wave_3 <- as.Date("2020-10-15") # start of wave 3
-
-# all cases map 
-map_total <- county_map %>% 
-  filter(Date == max(as.Date(county_map$Date), na.rm = TRUE)) %>% # need to remove NA
-  rename(state_fips = STATEFP) %>%
-  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)
-
-# wave 2 start date 
-map_w2_start <- map_daily %>%
-  filter(Date == "2020-06-15") %>% 
-  rename(state_fips = STATEFP) %>%
-  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)
-
-# wave 2 map 
-map_w2 <- map_total %>% 
-  st_join(map_w2_start, by = "fips") %>% 
-  mutate(Confirmed = Confirmed.x - Confirmed.y) %>% 
-  mutate(Deaths = Deaths.x - Deaths.y) %>% 
-  rename(start_date = Date.y) %>% 
-  rename(end_date = Date.x) %>% 
-  rename(fips = fips.x) %>% 
-  rename(county = NAME.x) %>% 
-  rename(population = population.x) %>% 
-  rename(csa = csa.x) %>% 
-  rename(agg_index = agg_index.x) %>% 
-  rename(state = state.x) %>%
-  rename(index = index.x) %>%
-  rename(county_fips = COUNTYFP.x) %>%
-  rename(state_fips = state_fips.x) %>%
-  select(fips, index, agg_index, state, county, csa, population, Confirmed, Deaths, start_date, end_date, state_fips, county_fips, geometry, LSAD.x, ALAND.x, AWATER.x)
+state_geo <- merge(state_lines, state_all, by = "NAME")
+state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')
 
@@
 ```{r}
+
+options(scipen = 999)
+
 library(usmap) # to get US Map Data
+library(cowplot) # for grid plot
 
@@
 
-# county state map 
-mapping <- function(map_type, pax) {
-  # 
-  # fips_code <- fips(state = state_name) 
-  # 
-  # map <- map_type %>% 
-  #   filter(state_fips == fips_code) %>% 
-  #   filter(Deaths >= 0)
-  #
+g_total <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_total/population), size = 0.3) +  # color = NA 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Total Deaths\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
 
-  map <- map_type
+g_w1 <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_W1/population), size = 0.3) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Wave 1 Deaths Only\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
 
-  ggplot(data = map) +
-    geom_sf(aes(fill = Deaths, color = Deaths)) + 
-    scale_fill_viridis_c(option = "viridis", direction = -1) + 
-    labs(title = "something", #paste(state),
-         subtitle = "Total Deaths") +
+g_w2 <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_W2_only/population), size = 0.3) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Wave 2 Deaths Only\n") + 
   theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
+
+g_w3 <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_W3/population), size = 0.3) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Wave 3 Deaths Onwards\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
+
+p <- plot_grid(g_total, g_w2, g_w3, ncol = 3)
+
+title <- ggdraw() + 
+  draw_label("Total Deaths per Capita in U.S. by State per Wave", fontface='bold') + 
   theme(text = element_text(family = "raleway"))
 
-  # if (pax = FALSE) {
-  #   
-  #   map <- 
-  #   
-  # } else if (pax = TRUE) {
-  #   
-  #   
-  #
+plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
 
-}
-
-dim(map_w2)
-
-length(map_total$Deaths)
-length(map_w2$Deaths)
-
-mapping(map_w2)
+# ggsave("viz/state-waves.png", width = 8, height = 3, dpi = 300)
+ggsave("viz/state-waves-total.png", g_total, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-waves-W1.png", g_w1, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-waves-W2.png", g_w2, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-waves-W3.png", g_w3, width = 8, height = 5, dpi = 300)
 

### 2020-11-25 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
msa and region waves I, II, III

--- 2020-11-24 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
@@
 covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split 
+covid_state <- read_csv("data/us_state_level_clean_2020-11-20.csv") # states only
 
@@
 # View(msa)
+
+covid_state
 ```
@@
 
-# # left join index of 50 with fips codes
-# msa_index_fips <- msa_index %>% 
-#   left_join(msa_w_fips, by = c("CSA Title", "County/County Equivalent", "State Name")) %>% 
-#   rename(county = `County/County Equivalent`) %>% 
-#   rename(csa = `CSA Title`) %>%
-#   rename(state = `State Name`) %>%
-#   rename(agg_index = mix_index)
-
 msa_index_2 <- msa_index %>% 
@@
   rename(agg_index = mix_index) %>% 
-  rename(fips = FIPS_combined) %>% 
-  names()
-
-# anti join to get the counties not in msa_index 
-region_fips <- msa_w_fips %>% 
-  anti_join(msa_index, by = "CSA Title") %>% 
-  rename(county = `County/County Equivalent`) %>%
-  mutate(agg_index = `State Name`) %>%
-  rename(state = `State Name`) %>%
-  rename(csa = `CSA Title`)
+  rename(fips = FIPS_combined)
 
@@
 state_msa_fips_join <- covid %>% 
-  full_join(msa_index_fips, by = "agg_index")# %>%
+  full_join(msa_index_2, by = "agg_index")# %>%
 
@@
   anti_join(msa_covid_only) %>%
-  select(-fips, -mix_index_population, -FIPS_population, -FIPS_combined, -`Central/Outlying County`, -state, -county, -csa, -`Metropolitan Division Title`)
+  select(-fips, -mix_index_population, -FIPS_population, -`Central/Outlying County`, -state, -county, -csa, -`Metropolitan Division Title`)
 
-names(msa_region)
+```
 
-region_fips_join <- msa_region %>% 
-  full_join(region_fips, by = "agg_index") %>%
-  drop_na(fips)
+# Getting Wave Count
 
-View(msa_covid_only)
+## for States
 
-View(region_fips_join)
-View(covid)
+```{r}
+wave_2 <- as.Date("2020-06-15") # start of wave 2
+wave_3 <- as.Date("2020-10-15") # start of wave 3
 
-final_daily <- msa_fips_join %>% # or final total 
-  bind_rows(region_fips_join) %>% 
-  select(index, Date, state, agg_index, fips, csa, county, population, 
-         Confirmed, Deaths, new_confirmed, new_deaths, 
-         new_confirmed_7, new_deaths_7, new_confirmed_14, new_deaths_14,
-         per_100k_new_confirmed_7, per_100k_new_deaths_7)
+state_covid_reg <- covid_state %>% 
+  rename("NAME" = "Province_State") %>%
+  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))
 
-names(final_daily)
-# write.csv(final_daily, "data/final-50-county-daily.csv")
+state_reg_total <- state_covid_reg %>% 
+  filter(Date == max(as.Date(state_covid_reg$Date), na.rm = TRUE)) %>% # need to remove NA
+  mutate(Date_total = Date)
+
+state_reg_w2_start <- state_covid_reg %>% 
+  filter(Date == wave_2) %>%
+  mutate(Date_W2 = Date)
+
+state_reg_w3_start <- state_covid_reg %>% 
+  filter(Date == wave_3) %>% 
+  mutate(Date_W3 = Date)
+
+View(state_reg_w2_start)
+View(state_reg_total)
+View(state_all)
+
+state_reg_all <- state_reg_total %>% 
+  rename(Confirmed_total = Confirmed) %>% # Start - Latest
+  rename(Deaths_total = Deaths) %>%
+  # joining wave 2 and beyond numbers
+  left_join(state_reg_w2_start, by = "NAME") %>%  # Start - June 15
+  # wave 1 only %>% 
+  rename(Confirmed_W1 = Confirmed) %>%  # Start - June 15
+  rename(Deaths_W1 = Deaths) %>% # Start - June 15
+  # wave 2 and beyond column 
+  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 - Latest
+  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 - Latest
+  # joining wave 3 numbers 
+  left_join(state_reg_w3_start, by = "NAME") %>% # Start - Oct 15 ie Wave 1 & Wave 2
+  # only wave 2 
+  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 - Oct 15
+  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 - Oct 15
+  # wave 3 and beyond
+  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 - Latest
+  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 - Latest
+  # tidy up, selecting only relevant variables
+  select(NAME, population, 
+         Date_total, Confirmed_total, Deaths_total, 
+         Confirmed_W1, Deaths_W1,
+         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
+         Date_W3, Confirmed_W3, Deaths_W3, stringency, containment, RegionCode)
+
+```
+
+## for States without MSA
+
+```{r}
+
+state_covid_only_2 <- state_covid_only %>% 
+  rename("NAME" = "agg_index") %>%
+  filter(! NAME %in% c("Alaska", "Hawaii", "Rhode Island", "District of Columbia"))
+
+state_total <- state_covid_only_2 %>% 
+  filter(Date == max(as.Date(state_covid_only_2$Date), na.rm = TRUE)) %>% # need to remove NA
+  mutate(Date_total = Date)
+
+state_w2_start <- state_covid_only_2 %>% 
+  filter(Date == max(wave_2)) %>%
+  mutate(Date_W2 = Date)
+
+state_w3_start <- state_covid_only_2 %>% 
+  filter(Date == max(wave_3)) %>% 
+  mutate(Date_W3 = Date)
+
+state_all <- state_total %>% 
+  rename(Confirmed_total = Confirmed) %>% # Start to Latest
+  rename(Deaths_total = Deaths) %>%
+  # joining wave 2 and beyond numbers
+  left_join(state_w2_start, by = "NAME") %>%  # Start to June 15
+  # wave 1 only %>% 
+  rename(Confirmed_W1 = Confirmed) %>%  # Start to June 15
+  rename(Deaths_W1 = Deaths) %>% # Start to June 15
+  # wave 2 and beyond column 
+  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 to Latest
+  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 to Latest
+  # joining wave 3 numbers 
+  left_join(state_w3_start, by = "NAME") %>% # Start to Oct 15 ie Wave 1 & Wave 2
+  # only wave 2 
+  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 to Oct 15
+  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 to Oct 15
+  # wave 3 and beyond
+  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 to Latest
+  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 to Latest
+  # tidy up, selecting only relevant variables
+  select(index, NAME, population, 
+         Date_total, Confirmed_total, Deaths_total, 
+         Confirmed_W1, Deaths_W1,
+         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only, 
+         Date_W3, Confirmed_W3, Deaths_W3)
+
+View(msa_all)
+
+```
+
+## for MSAs/Top 50
+
+```{r}
+
+msa_covid_only_2 <- msa_covid_only %>% 
+  rename("NAME" = "agg_index")
+
+msa_total <- msa_covid_only_2 %>% 
+  filter(Date == max(as.Date(msa_covid_only_2$Date), na.rm = TRUE)) %>% # need to remove NA
+  mutate(Date_total = Date)
+
+msa_w2_start <- msa_covid_only_2 %>% 
+  filter(Date == max(wave_2)) %>%
+  mutate(Date_W2 = Date)
+
+msa_w3_start <- msa_covid_only_2 %>% 
+  filter(Date == max(wave_3)) %>% 
+  mutate(Date_W3 = Date)
+
+msa_all <- msa_total %>% 
+  rename(Confirmed_total = Confirmed) %>% # Start - Latest
+  rename(Deaths_total = Deaths) %>%
+  # joining wave 2 and beyond numbers
+  left_join(msa_w2_start, by = "NAME") %>%  # Start - June 15
+  # wave 1 only %>% 
+  rename(Confirmed_W1 = Confirmed) %>%  # Start - June 15
+  rename(Deaths_W1 = Deaths) %>% # Start - June 15
+  # wave 2 and beyond column 
+  mutate(Confirmed_W2_plus = Confirmed_total - Confirmed_W1) %>% # June 15 - Latest
+  mutate(Deaths_W2_plus = Deaths_total - Deaths_W1) %>% # June 15 - Latest
+ left_join(msa_w3_start, by = "NAME") %>% # Start - Oct 15 ie Wave 1 & Wave 2
+  # only wave 2 
+  mutate(Deaths_W2_only = Deaths - Deaths_W1) %>% # June 15 - Oct 15
+  mutate(Confirmed_W2_only = Confirmed - Confirmed_W1) %>% # June 15 - Oct 15
+  # wave 3 and beyond
+  mutate(Deaths_W3 = Deaths_total - Deaths) %>%  # Oct 15 - Latest
+  mutate(Confirmed_W3 = Confirmed_total - Confirmed) %>% # Oct 15 - Latest
+  # tidy up, rename
+  rename(metro_title = "Metropolitan Division Title") %>%
+  rename(central_outlying = "Central/Outlying County") %>%
+  # selecting only relevant variables
+  select(index, NAME, fips, csa, county, metro_title, population, FIPS_population, mix_index_population,
+         Date_total, Confirmed_total, Deaths_total,
+         Confirmed_W1, Deaths_W1,
+         Date_W2, Confirmed_W2_plus, Confirmed_W2_only, Deaths_W2_plus, Deaths_W2_only,
+         Date_W3, Confirmed_W3, Deaths_W3) %>%
+  distinct()
 
@@
 # add lat long attributes to county dataset          
-county_geo <- merge(county_lines, final_daily, by = "fips")
+county_geo <- merge(county_lines, msa_all, by = "fips")
 county_map <- county_geo %>% st_transform('+proj=longlat +datum=WGS84')
 
-map_daily <- county_map # daily covid map
+# merge state 
+# add lat long attributes to state dataset
 
-```
+state_geo <- merge(state_lines, state_all, by = "NAME")
+state_map <- state_geo %>% st_transform('+proj=longlat +datum=WGS84')
 
-# Wave/Dates Filter  
-### For total cases
-
-```{r}
-# library(spdplyr) # spatial version of dplyr
-
-# Dates of Waves
-
-wave_2 <- as.Date("2020-06-15") # start of wave 2
-wave_3 <- as.Date("2020-10-15") # start of wave 3
-
-# all cases map 
-map_total <- county_map %>% 
-  filter(Date == max(as.Date(county_map$Date), na.rm = TRUE)) %>% # need to remove NA
-  rename(state_fips = STATEFP) %>%
-  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)
-
-# wave 2 start date 
-map_w2_start <- map_daily %>%
-  filter(Date == "2020-06-15") %>% 
-  rename(state_fips = STATEFP) %>%
-  select(-new_confirmed, -new_deaths, -new_confirmed_7, -new_deaths_7, -new_confirmed_14, -new_deaths_14, -per_100k_new_confirmed_7, -per_100k_new_deaths_7, -COUNTYNS, -AFFGEOID, -GEOID, -county)
-
-# wave 2 map 
-map_w2 <- map_total %>% 
-  st_join(map_w2_start, by = "fips") %>% 
-  mutate(Confirmed = Confirmed.x - Confirmed.y) %>% 
-  mutate(Deaths = Deaths.x - Deaths.y) %>% 
-  rename(start_date = Date.y) %>% 
-  rename(end_date = Date.x) %>% 
-  rename(fips = fips.x) %>% 
-  rename(county = NAME.x) %>% 
-  rename(population = population.x) %>% 
-  rename(csa = csa.x) %>% 
-  rename(agg_index = agg_index.x) %>% 
-  rename(state = state.x) %>%
-  rename(index = index.x) %>%
-  rename(county_fips = COUNTYFP.x) %>%
-  rename(state_fips = state_fips.x) %>%
-  select(fips, index, agg_index, state, county, csa, population, Confirmed, Deaths, start_date, end_date, state_fips, county_fips, geometry, LSAD.x, ALAND.x, AWATER.x)
+state_reg_geo <- merge(state_lines, state_reg_all, by = "NAME")
+state_reg_map <- state_reg_geo %>% st_transform('+proj=longlat +datum=WGS84')
 
@@
 
+## state
+
 ```{r}
+
+options(scipen = 999)
+
 library(usmap) # to get US Map Data
+library(cowplot) # for grid plot
 
-names(map_w2)
-fips("New York")
+# fips("New York")
 #state_name,
 
-# county state map 
-mapping <- function(map_type, pax) {
-  # 
-  # fips_code <- fips(state = state_name) 
-  # 
-  # map <- map_type %>% 
-  #   filter(state_fips == fips_code) %>% 
-  #   filter(Deaths >= 0)
-  #
+g_total <- ggplot() +
+  geom_sf(data = state_map, 
+          aes(fill = Deaths_total/population), 
+          color = NA) +  # color = NA 
+  geom_sf(data = county_map, aes(fill = Deaths_total/population), color = NA) + # MSA Things
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Total Deaths\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
 
-  map <- map_type
+max_W1 <- max(c(state_map$Deaths_W1 / state_map$population, county_map$Deaths_W1 / county_map$population))
+max_W2 <- max(c(state_map$Deaths_W2_only / state_map$population, county_map$Deaths_W2_only / county_map$population))
+max_W3 <- max(c(state_map$Deaths_W3 / state_map$population, county_map$Deaths_W3 / county_map$population))
 
-  ggplot(data = map) +
-    geom_sf(aes(fill = Deaths, color = Deaths)) + 
-    scale_fill_viridis_c(option = "viridis", direction = -1) + 
-    labs(title = "something", #paste(state),
-         subtitle = "Total Deaths") +
+min_W1 <- min(c(state_map$Deaths_W1 / state_map$population, county_map$Deaths_W1 / county_map$population))
+min_W2 <- min(c(state_map$Deaths_W2_only / state_map$population, county_map$Deaths_W2_only / county_map$population))
+min_W3 <- min(c(state_map$Deaths_W3 / state_map$population, county_map$Deaths_W3 / county_map$population))
+
+max_W1 
+max_W2
+max_W3
+min_W1
+min_W2
+min_W3
+
+max(state_map$Deaths_W1 / state_map$population)
+max(county_map$Deaths_W1 / county_map$population)
+min(state_map$Deaths_W1 / state_map$population)
+min(county_map$Deaths_W1 / county_map$population)
+
+max(state_map$Deaths_W2_only / state_map$population)
+max(county_map$Deaths_W2_only / county_map$population)
+min(state_map$Deaths_W2_only / state_map$population)
+min(county_map$Deaths_W2_only / county_map$population)
+
+max(state_map$Deaths_W3 / state_map$population)
+max(county_map$Deaths_W3 / county_map$population)
+min(state_map$Deaths_W3 / state_map$population)
+min(county_map$Deaths_W3 / county_map$population)
+
+g_w1 <- ggplot(data = state_reg_map) +
+  geom_sf(aes(fill = Deaths_W1/population), color = NA) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W1, max_W1)) + 
+  labs(title = "U.S. States - Wave 1 Deaths per Capita",
+      subtitle = "Until 15 June 2020\n") + 
   theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+g_w2 <- ggplot(data = state_reg_map) +
+  geom_sf(aes(fill = Deaths_W2_only/population), color = NA) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W2, max_W2)) + 
+  labs(title = "U.S. States - Wave 2 Deaths per Capita",
+       subtitle = "From 15 June 2020 - 15 October 2020\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+g_w3 <- ggplot(data = state_reg_map) +
+  geom_sf(aes(fill = Deaths_W3/population), color = NA) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W3, round(max_W3))) + 
+  labs(title = "U.S. States - Wave 3 Onwards Deaths per Capita",
+       subtitle = "After 15 October 2020\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+# state and county msa 
+g_c_w1 <- ggplot() +
+  geom_sf(data = state_map, aes(fill = Deaths_W1/population), color = NA) +
+  geom_sf(data = county_map, aes(fill = Deaths_W1/population), color = NA) + # MSA Things
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W1, max_W1)) + 
+  labs(title = "Metropolitan State Areas, Regional Areas - Wave 1 Deaths per Capita", 
+       subtitle = "Until 15 June 2020\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+g_c_w2 <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_W2_only/population), color = NA) + # size = 0.3
+  geom_sf(data = county_map, aes(fill = Deaths_W2_only/population), color = NA) + # MSA Things
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W2, max_W2)) +
+  labs(title = "Metropolitan State Areas, Regional Areas - Wave 2 Deaths per Capita",
+       subtitle = "From 15 June 2020 - 15 October 2020\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+g_c_w3 <- ggplot(data = state_map) +
+  geom_sf(aes(fill = Deaths_W3/population), color = NA) + 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1, limits = c(min_W3, round(max_W3))) + 
+    geom_sf(data = county_map, aes(fill = Deaths_W3/population), color = NA) + # MSA Things
+  labs(title = "Metropolitan State Areas, Regional Areas - Wave 3 Deaths per Capita",
+       subtitle = "After 15 October 2020\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 30))
+
+# Grid Plots 
+p <- plot_grid(g_total, g_w2, g_w3, ncol = 3)
+
+p_2 <- plot_grid(g_w2, g_c_w2, ncol = 2)
+p_3 <- plot_grid(g_w3, g_c_w3, ncol = 2)
+
+title <- ggdraw() + 
+  draw_label("Total Deaths per Capita in U.S. by State per Wave", fontface='bold') + 
   theme(text = element_text(family = "raleway"))
 
-  # if (pax = FALSE) {
-  #   
-  #   map <- 
-  #   
-  # } else if (pax = TRUE) {
-  #   
-  #   
-  #
+plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
 
-}
+# ggsave("viz/state-waves.png", width = 8, height = 3, dpi = 300)
+# ggsave("viz/state-waves-total.png", g_total, width = 8, height = 5, dpi = 300)
+# ggsave("viz/state-waves-W1.png", g_w1, width = 8, height = 5, dpi = 300)
+# ggsave("viz/state-waves-W2.png", g_w2, width = 8, height = 5, dpi = 300)
+# ggsave("viz/state-waves-W3.png", g_w3, width = 8, height = 5, dpi = 300)
 
-dim(map_w2)
+ggsave("viz/state-wave-w1.png", g_w1, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-msa-wave-w1.png", g_c_w1, width = 8, height = 5, dpi = 300)
 
-length(map_total$Deaths)
-length(map_w2$Deaths)
+ggsave("viz/state-wave-w2.png", g_w2, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-msa-wave-w2.png", g_c_w2, width = 8, height = 5, dpi = 300)
 
-mapping(map_w2)
+ggsave("viz/state-wave-w3.png", g_w3, width = 8, height = 5, dpi = 300)
+ggsave("viz/state-msa-wave-w3.png", g_c_w3, width = 8, height = 5, dpi = 300)
+
+ggsave("viz/state-msa-waves-W1.png", plot_grid(g_w1, g_c_w1, ncol = 2), width = 14, height = 5, dpi = 300)
+ggsave("viz/state-msa-waves-W2.png", plot_grid(g_w2, g_c_w2, ncol = 2), width = 14, height = 5, dpi = 300)
+ggsave("viz/state-msa-waves-W3.png", plot_grid(g_w3, g_c_w3, ncol = 2), width = 14, height = 5, dpi = 300)
+
+# ggsave("viz/state-msa-waves-temp.png", width = 14, height = 5, dpi = 300)
 
 ```
+
+## top 50
+
+```{r}
+
+c_total <- ggplot(data = county_map) +
+  geom_sf(aes(fill = Deaths_total/population), size = 0.3) +  # color = NA 
+  scale_fill_viridis_c(name = "Deaths per Capita", option = "viridis", direction = -1) + 
+  labs(title = "Total Deaths\n") + 
+  theme_void() + 
+  theme(text = element_text(family = "raleway", size = 50))
+
+```

### 2020-12-01 KPMG Team/Dashboard/app.R
dashboard done

@@ -1,14 +1,14 @@
 
-# ------------------------------------------------------------------------------------ # 
+# ------------------------------------------------------------------------------------ #
 # ************************************************************************************ #
 
-# COVID-19 Intelligent Forecasting Dashboard Functions 
-# KPMG x QMSS Practicum 2020 
-# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong 
-  
+# COVID-19 Intelligent Forecasting Dashboard Functions
+# KPMG x QMSS Practicum 2020
+# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong
+
 # Dashboard Lead: Louisa Ong
 
-# ------------------------------------------------------------------------------------ # 
+# ------------------------------------------------------------------------------------ #
 # ************************************************************************************ #
 
 
@@ -16,40 +16,242 @@
 
 library(shiny)
 library(shinythemes)
+library(plotly)
+library(tidyverse)
+library(ggplot2)
+library(RColorBrewer)
+library(showtext)
+library(janitor)
+
+# Fonts
+showtext_auto()
+font_add(family = "raleway", regular = "fonts/Raleway-Regular.ttf")
+font_add(family = "instruction", regular = "fonts/Instruction.otf")
+font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
+font_add(family = "proxima-nova", regular = "fonts/ProximaNova-Reg.ttf")
+font_add(family = "proxima-nova-light", regular = "fonts/ProximaNova-Light.ttf")
+# font_add(family = "montserratlight", regular = "fonts/Montserrat-Light.ttf")
+# font_add(family = "montserratthin", regular = "fonts/Montserrat-Thin.ttf")
+
+# Turn off scientific notation
+options(scipen = 999)
+
+# import latest csv with updated predictions
+covid <- read_csv("data/us_state_level_clean_2020-11-29.csv")
 
 
-# define UI 
+details <- covid %>%
+  filter(Date == as.Date(max(Date))) %>%
+  select(Province_State, population, Deaths, Confirmed, stringency) %>%
+  rename("State" = Province_State) %>%
+  rename("Stringency Index" = stringency) %>%
[…]

### 2020-12-01 KPMG Team/Dashboard/for later/functions.R

@@ -1,0 +1,28 @@
+
+# ------------------------------------------------------------------------------------ # 
+# ************************************************************************************ #
+
+# COVID-19 Intelligent Forecasting Dashboard Functions 
+# KPMG x QMSS Practicum 2020 
+# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong 
+
+# Dashboard Lead: Louisa Ong
+
+# ------------------------------------------------------------------------------------ # 
+# ************************************************************************************ #
+
+# module server function
+metroAreas <- function(id, label = "Metro Areas") {
+  
+  moduleServer(
+    id, 
+    
+    # actual module function
+    function(input, output, session) {
+      
+    return()
+      
+    }
+  )
+  
+}

### 2020-12-01 KPMG Team/Dashboard/for later/server.R

@@ -1,0 +1,29 @@
+library(shiny)
+library(tidyverse)
+library(dygraphs)
+
+source('setup.R')
+
+
+shinyServer(function(input, output, session) {
+
+  # Build reactive expression that takes values from the input to fetch the data
+  # state <- reactive({
+  #   merge_to_wdi(input$state)
+  # })
+  checkboxGroupInput(
+    "state", label = "States:", choices = unique(covid$Province_State), selected = "New York",
+     width = "100%")
+
+  output$plot_d <- renderPlotly({
+
+    covid_sel <- covid_data %>%
+      filter(Province_State %in% state)
+
+    ggplotly(g_deaths)
+
+  })
+
+})
+
+unique(covid$Province_State)

### 2020-12-01 KPMG Team/Dashboard/for later/setup.R

@@ -1,0 +1,69 @@
+library(shiny)
+library(tidyverse)
+library(readr)
+library(ggplot2)
+library(RColorBrewer)
+library(showtext)
+
+# Fonts
+showtext_auto()
+font_add(family = "raleway", regular = "fonts/Raleway-Regular.ttf")
+font_add(family = "instruction", regular = "fonts/Instruction.otf")
+
+# Turn off scientific notation
+options(scipen=999)
+
+# import latest csv with updated predictions
+covid <- read_csv("visualization_scripts/data/us_state_level_clean_2020-11-29.csv")
+
+lookup <- covid
+
+names(lookup) <- c("a", "b", "c", "d", "e", "f")
+# Lookup variable names / descriptions
+lookup_list <- setNames(as.list(lookup$b), lookup$a)
+lookup_list2 <- setNames(as.list(lookup$a), lookup$b)
+
+lookup_state <- setNames(as.list(covid$Province_State), covid$State_FIPS)
+
+
+# selecting states of interest
+states_sel <- c("New York", "Florida", "Texas", "California")
+
+# filtering dataframe for states
+covid_sel <- covid %>%
+  filter(Province_State %in% states_sel)
+
+
+# LINE GRAPH PLOTS
+
+# Deaths
+g_deaths <- ggplot(data = covid_sel,
+                   aes(x = Date, y = new_deaths_7,
+                       group = Province_State, color = Province_State)) +
+  geom_line(size = 1.1) +
+  scale_color_viridis_d(option = "viridis", name = "State") +
+  geom_vline(xintercept = as.Date("2020-10-01"), color = "slategray", size = 2, alpha = 0.3) +
+  # geom_rect(aes(xmin = as.Date("2020-10-01"), xmax = max(Date), ymin = 0, ymax = max(new_deaths_7)), alpha = 0.1, inherit_aes = FALSE) +
+  labs(title = "\nCOVID-19 Daily Deaths Count\n",
+       x = "\nDate\n",
+       y = "\n7-Day Average of Daily Deaths\n") +
+  theme_minimal() +
+  theme(text = element_text(family = "raleway"),
+        plot.title = element_text(family = "instruction", color = "grey20"),
+        axis.line = element_line(color = "grey60"))
+
+
+
+# Confirmed
+g_confirmed <- ggplot(data = covid_sel, aes(x = Date, y = new_confirmed_7, group = Province_State, color = Province_State)) +
+  geom_line(size = 1.1) +
[…]

### 2020-12-01 KPMG Team/Dashboard/for later/ui.R

@@ -1,0 +1,56 @@
+library(shiny)
+library(dplyr)
+
+source('setup.R')
+
+shinyUI(
+  fluidPage(
+
+    titlePanel("COVID-19 Intelligent Forecasting"),
+
+    sidebarLayout(
+      sidebarPanel(title="Make selections",
+                   selectInput("Province_State", label = "State",
+                               lookup_state,
+                               selected = "New York")
+      ),
+
+      mainPanel(h2("Visualization"),
+
+                g_deaths
+
+                )
+
+
+    )
+
+
+    # tabsetPanel(
+    #
+    #   tabPanel(
+    #     "Controls",
+    #     selectInput(
+    #       'indicator',
+    #       'COVID-19',
+    #       lookup_list2,
+    #       selected = 'Fertility rate, total (births per woman)'
+    #     ),
+    #
+    #   selectInput(
+    #     "ctry",
+    #     label = "Country",
+    #     lookup_list_countries,
+    #     selected = "US"
+    #     )
+    #   ),
+    #
+    # tabPanel(
+    #   "Chart",
+    #   p("Select a country for an interactive chart"),
+    #   dygraphOutput("dchart", width = "100%")
+    #   )
+
+    )
+  )
+)
+

### 2020-12-01 KPMG Team/Dashboard/functions.R

@@ -1,28 +1,0 @@
-
-# ------------------------------------------------------------------------------------ # 
-# ************************************************************************************ #
-
-# COVID-19 Intelligent Forecasting Dashboard Functions 
-# KPMG x QMSS Practicum 2020 
-# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong 
-
-# Dashboard Lead: Louisa Ong
-
-# ------------------------------------------------------------------------------------ # 
-# ************************************************************************************ #
-
-# module server function
-metroAreas <- function(id, label = "Metro Areas") {
-  
-  moduleServer(
-    id, 
-    
-    # actual module function
-    function(input, output, session) {
-      
-    return()
-      
-    }
-  )
-  
-}

### 2020-12-01 KPMG Team/Dashboard/app.R
dashboard update
numbesr per capita

@@ -29,7 +29,7 @@ font_add(family = "raleway", regular = "fonts/Raleway-Regular.ttf")
 font_add(family = "instruction", regular = "fonts/Instruction.otf")
 font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
 font_add(family = "proxima-nova", regular = "fonts/ProximaNova-Reg.ttf")
-font_add(family = "proxima-nova-light", regular = "fonts/ProximaNova-Light.ttf")
+font_add(family = "proximanovalight", regular = "fonts/ProximaNova-Light.ttf")
 # font_add(family = "montserratlight", regular = "fonts/Montserrat-Light.ttf")
 # font_add(family = "montserratthin", regular = "fonts/Montserrat-Thin.ttf")
 
@@ -42,7 +42,10 @@ covid <- read_csv("data/us_state_level_clean_2020-11-29.csv")
 
 details <- covid %>%
   filter(Date == as.Date(max(Date))) %>%
-  select(Province_State, population, Deaths, Confirmed, stringency) %>%
+  mutate("Total Deaths per Capita" = signif(Deaths/population, 3)) %>%
+  mutate("Total Cases per Capita" = signif(Confirmed/population, 3)) %>%
+  select(Province_State, population, Deaths, Confirmed,
+         `Total Deaths per Capita`, `Total Cases per Capita`, stringency) %>%
   rename("State" = Province_State) %>%
   rename("Stringency Index" = stringency) %>%
   # rename("Containment Index" = containment) %>%
@@ -97,6 +100,8 @@ gCon_fun <- function(input_state) {
           axis.line = element_line(color = "grey60"))
 }
 
+
+gDeaths_fun(c("New York", "Florida", "California"))
 
 detailsFill_fun <- function(input_state) {
 
@@ -228,7 +233,6 @@ ui <- fluidPage(
     )
 
 )
-
 
 server <- function(input, output) {
 

### 2020-12-02 KPMG Team/Dashboard/app.R
multiple tabs

@@ -6,16 +6,25 @@
 # KPMG x QMSS Practicum 2020
 # Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong
 
-# Dashboard Lead: Louisa Ong
+# Visualizations Lead: Louisa Ong
 
 # ------------------------------------------------------------------------------------ #
 # ************************************************************************************ #
 
 
+
 # rShiny dashboard app app.R file
+
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/ #
+# ------------------------------------------------------------------ #
+# ---------------------------- SETUP ------------------------------- #
+# ------------------------------------------------------------------ #
+# \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/ #
+
 
 library(shiny)
 library(shinythemes)
+library(shinydashboard)
 library(plotly)
 library(tidyverse)
 library(ggplot2)
@@ -23,13 +32,14 @@ library(RColorBrewer)
 library(showtext)
 library(janitor)
 
+
 # Fonts
 showtext_auto()
 font_add(family = "raleway", regular = "fonts/Raleway-Regular.ttf")
 font_add(family = "instruction", regular = "fonts/Instruction.otf")
-font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
-font_add(family = "proxima-nova", regular = "fonts/ProximaNova-Reg.ttf")
-font_add(family = "proximanovalight", regular = "fonts/ProximaNova-Light.ttf")
+# font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
+# font_add(family = "proxima-nova", regular = "fonts/ProximaNova-Reg.ttf")
+# font_add(family = "proximanovalight", regular = "fonts/ProximaNova-Light.ttf")
 # font_add(family = "montserratlight", regular = "fonts/Montserrat-Light.ttf")
 # font_add(family = "montserratthin", regular = "fonts/Montserrat-Thin.ttf")
 
@@ -37,134 +47,342 @@ font_add(family = "proximanovalight", regular = "fonts/ProximaNova-Light.ttf")
 options(scipen = 999)
 
 # import latest csv with updated predictions
[…]

### 2020-12-05 KPMG Team/Dashboard/visualization_scripts/msa-waves-pax.Rmd
dashboard done
Dashboard Update
All tabs, viz of US Waves and Clusters

@@
 
+install.packages("profvis")
+library(profvis)
+
 ```
@@
 # from Sydney
+# msa_index <- read_csv("data/msa/msa_mix_index.csv") # csa and msa with population 
+# covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split 
+# covid_state <- read_csv("data/us_state_level_clean_2020-11-20.csv") # states only
+
 msa_index <- read_csv("data/msa/msa_mix_index.csv") # csa and msa with population
-covid <- read_csv("data/msa/us_daily_msa_split_2020-11-16.csv") # split 
-covid_state <- read_csv("data/us_state_level_clean_2020-11-20.csv") # states only
+covid <- read_csv("data/msa/us_daily_msa_final_2020-12-02.csv") # split
+covid_state <- read_csv("data/us_state_level_clean_2020-12-02.csv") # states only
 
@@
 
-covid_state
 ```
@@
   mutate(Date_W3 = Date)
-
-View(state_reg_w2_start)
-View(state_reg_total)
-View(state_all)
 

### 2020-12-05 KPMG Team/Dashboard/app.R

@@ -4,7 +4,7 @@
 
 # COVID-19 Intelligent Forecasting Dashboard Functions
 # KPMG x QMSS Practicum 2020
-# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Sherry Huang, Jen Woo, Louisa Ong
+# Team: Ariel Luo, Sydney Bolim Son, Andrew Thvedt, Jen Woo, Louisa Ong
 
 # Visualizations Lead: Louisa Ong
 
@@ -21,16 +21,26 @@
 # ------------------------------------------------------------------ #
 # \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/ #
 
-
+# shiny essentials
 library(shiny)
 library(shinythemes)
 library(shinydashboard)
-library(plotly)
-library(tidyverse)
-library(ggplot2)
-library(RColorBrewer)
-library(showtext)
+
+library(plotly) # for interactive plots 
+library(tidyverse) # datacleaning
+library(ggplot2) # graphing
+library(RColorBrewer) # graph color palette 
+library(showtext) # for fonts
 library(janitor)
+
+# spatial essentials 
+library(sf) # for merging spatial datasets 
+library(jsonlite) # from FIPS -> Geo (lat, long)
+library(gridExtra) # for grid plot
+# library(gganimate) # for interactive plot 
+# library(tmap)
+library(profvis)
+
 
 
 # Fonts
@@ -46,43 +56,347 @@ font_add(family = "instruction", regular = "fonts/Instruction.otf")
 # Turn off scientific notation
 options(scipen = 999)
 
-# import latest csv with updated predictions
-# TO DO: use Sydney's branch on Andrew's Github
-covid <- read_csv("data/us_state_level_clean_2020-11-29.csv")
+# ------------------------  Importing Datasets -------------------------- # 
 
+# US State Level Clean
+covid_state <- read_csv("data/us_state_level_clean_2020-12-02.csv") # States only
+
+# MSAs & Regional Areas 
+covid <- read_csv("data/us_daily_msa_final_2020-12-02.csv") # includes MSAs and Regional Areas
+
+# MSA fips codes 
+# msa <- read_csv("data/msa.csv")
+msa_index <- read_csv("data/msa_mix_index.csv") # csa and msa with population
+
+
+# html viz from Andrew & Sydney ML Model Analysis 
+cat_d_plot <- "viz/deaths_per_100k.html"
[…]

### 2020-12-08 KPMG Team/Dashboard/app.R
last week

@@ -92,7 +92,6 @@ covid_current <- covid_state %>%
     filter(Date == as.Date(min(c(Sys.Date(), max(Date)))))
   
 
-
 details <- function(covid_type) {
   
   covid_type %>%
@@ -376,8 +375,14 @@ g_c_w3 <- ggplot() +
         axis.line = element_line(color = NA))
 
 
-
+# attempts to make it faster
 gp_c_w1 <- ggplotly(g_c_w1) %>% 
+  partial_bundle()
+
+gp_c_w2 <- ggplotly(g_c_w2) %>% 
+  partial_bundle()
+
+gp_c_w3 <- ggplotly(g_c_w3) %>% 
   partial_bundle()
 
 
@@ -507,6 +512,69 @@ detailsCurrent_fun <- function(input_state) {
 # }
 
 
+# 
+#  
+# covid_renamed %>% select(y_val)
+# 
+# 
+# y_id <- c("death_7", "death_100", "death_cap", 
+#           "case_7", "case_100", "case_cap")
+# 
+# y_var <- c("7-Day Average Deaths", "Daily Deaths per 100k", "Daily Deaths per Capita", 
+#            "7-Day Average Cases", "Daily Cases per 100k", "Daily Cases per Capita")
+# 
+# g_title <- c("\nCOVID-19 Daily Death Count\n \n ", 
+#              "\nCOVID-19 Daily Deaths per 100k\n \n ", 
+#              "\nCOVID-19 Daily Deaths per Capita\n \n ",
+#              "\nCOVID-19 Daily Case Count\n \n ", 
+#              "\nCOVID-19 Daily Cases per 100k\n \n ", 
+#              "\nCOVID-19 Daily Cases per Capita\n \n ")y
+# 
+# y_lab <- c(" \n 7-Day Average Daily Deaths\n \n ", 
+#            "\n Daily Deaths per 100k\n \n ",
+#            " \n Daily Deaths per Capita\n \n ",
+#            " \n 7-Day Average Daily Cases\n \n ", 
+#            "\n Daily Cases per 100k\n \n ",
+#            " \n Daily Cases per Capita\n \n ")
+# 
+# g_df <- data.frame(y_id, y_var, g_title, y_lab)
+# 
+# g_row <- g_df %>% 
+#   filter(y_id == "death_7")
+# 
+# covid_renamed$y_val <- covid_renamed %>% select(g_row$y_var)
[…]

### 2020-12-09 KPMG Team/Dashboard/app.R
with dygraphs

@@ -32,6 +32,12 @@ library(ggplot2) # graphing
 library(RColorBrewer) # graph color palette 
 library(showtext) # for fonts
 library(janitor)
+library(viridis)
+
+
+library(dygraphs)
+library(xts)
+
 
 # spatial essentials 
 library(sf) # for merging spatial datasets 
@@ -47,7 +53,7 @@ library(profvis)
 showtext_auto()
 font_add(family = "raleway", regular = "fonts/Raleway-Regular.ttf")
 font_add(family = "instruction", regular = "fonts/Instruction.otf")
-# font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
+font_add(family = "montserrat", regular = "fonts/Montserrat-Regular.ttf")
 # font_add(family = "proxima-nova", regular = "fonts/ProximaNova-Reg.ttf")
 # font_add(family = "proximanovalight", regular = "fonts/ProximaNova-Light.ttf")
 # font_add(family = "montserratlight", regular = "fonts/Montserrat-Light.ttf")
@@ -59,10 +65,10 @@ options(scipen = 999)
 # ------------------------  Importing Datasets -------------------------- # 
 
 # US State Level Clean
-covid_state <- read_csv("data/us_state_level_clean_2020-12-02.csv") # States only
+covid_state <- read_csv("data/us_state_level_clean_2020-12-06.csv") # States only
 
 # MSAs & Regional Areas 
-covid <- read_csv("data/us_daily_msa_final_2020-12-02.csv") # includes MSAs and Regional Areas
+covid <- read_csv("data/us_daily_msa_final_2020-12-05.csv") # includes MSAs and Regional Areas
 
 # MSA fips codes 
 # msa <- read_csv("data/msa.csv")
@@ -125,13 +131,13 @@ details_current <- details(covid_current)
 
 # Renaming so that the variable names in plotly shows up well
 covid_renamed <- covid_state %>% 
-  mutate("7-Day Average Deaths" = new_deaths_7) %>%
-  mutate("Daily Deaths per 100k" = new_deaths_7/100000) %>%
-  mutate("Daily Deaths per Capita" = covid_state$new_deaths_7/covid_state$population) %>%
-  rename("Province/State" = Province_State) %>%
-  mutate("7-Day Average Cases" = new_confirmed_7) %>%
-  mutate("Daily Cases per 100k" = new_confirmed_7/100000) %>%
[…]

### 2020-12-09 KPMG Team/Dashboard/app.R
Latest

@@ -393,6 +393,8 @@ gcw3 <- gWave(state_map_2$`Wave 3 Cases per Capita`,
 
 # split up to make it *hopefully* faster
 
+# Include cluster data 
+
 detailsForecast_fun <- function(input_state) {
   
   details_forecast %>%
@@ -555,24 +557,24 @@ ui <- navbarPage(
 
                               # daily deaths & cases
                               tabPanel("Daily", br(),
-                                       dygraphOutput("gdeaths_daily"), 
-                                       br(),
-                                       br(),
-                                       dygraphOutput("gconfirmed_daily")),
+                                       dygraphOutput("gdeaths_daily")), 
+                                       # br(),
+                                       # br(),
+                                       # dygraphOutput("gconfirmed_daily")),
                               
                               # daily per 100k deaths & cases
                               tabPanel("Daily per 100k", br(),
-                                       dygraphOutput("gdeaths_100"), 
-                                       br(),
-                                       br(),
-                                       dygraphOutput("gconfirmed_100")),
+                                       dygraphOutput("gdeaths_100")), 
+                                       # br(),
+                                       # br(),
+                                       # dygraphOutput("gconfirmed_100")),
                               
                               # daily per capitadeaths & cases
                               tabPanel("Daily per Capita", br(),
-                                       dygraphOutput("gdeaths_pc"), 
-                                       br(),
-                                       br(),
-                                       dygraphOutput("gconfirmed_pc")), 
+                                       dygraphOutput("gdeaths_pc")), 
+                                       # br(),
+                                       # br(),
+                                       # dygraphOutput("gconfirmed_pc")), 
                   
                                br()
                   ),
@@ -725,7 +727,7 @@ server <- function(input, output) {
 
 
 # \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/ #
-# ------------------------------------------------------------------ #
+# -----------------------------------------            ------------------------- #
 # ----------------------------- RUN APP ---------------------------- #
 # ------------------------------------------------------------------ #
 # \/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/ #

### 2020-12-09 KPMG Team/Dashboard/app.R
renames

@@ -97,27 +97,23 @@ covid_forecast <- covid_state %>%
 covid_current <- covid_state %>%
     filter(Date == as.Date(min(c(Sys.Date(), max(Date)))))
   
-
 details <- function(covid_type) {
   
   covid_type %>%
-    mutate("Total Deaths per Capita" = signif(Deaths/population, 3)) %>%
-    mutate("Total Cases per Capita" = signif(Confirmed/population, 3)) %>%
-    
+
     rename("State" = Province_State) %>%
     rename("Population" = population) %>%
-    
     rename("Total Deaths" = Deaths) %>%
-    mutate("Total Deaths per 100k" = signif(`Total Deaths`/100000, 3)) %>% 
-    rename("Total Cases" = Confirmed) %>%
-    mutate("Total Cases per 100k" = signif(`Total Cases`/100000, 3)) %>% 
+    mutate("Total Deaths per 100k" = signif(per_100k_new_deaths_7, 3)) %>% 
+    # rename("Total Cases" = Confirmed) %>%
+    # mutate("Total Cases per 100k" = signif(`Total Cases`/100000, 3)) %>% 
     mutate("Spread Category" = "Semi-Controlled") %>% # arbitrary
     mutate("Mobility Index" = round(-22)) %>% 
     mutate("Containment Index" = round(containment)) %>%
     # days since peak 
     select("State", "Population", 
-    "Total Deaths", "Total Deaths per 100k", "Total Deaths per Capita", 
-    "Total Cases", "Total Cases per 100k", "Total Cases per Capita", 
+    "Total Deaths", "Total Deaths per 100k",
+    # "Total Cases", "Total Cases per 100k", "Total Cases per Capita", 
     "Containment Index", "Mobility Index")
   
 }
@@ -127,17 +123,16 @@ details_forecast <- details(covid_forecast)
 details_current <- details(covid_current)
 
 
+
 # ------------------------ Renaming Variables for Line Plot -------------------------- # 
 
 # Renaming so that the variable names in plotly shows up well
 covid_renamed <- covid_state %>% 
   mutate("7-Day Average Deaths" = round(new_deaths_7)) %>%
-  mutate("Daily Deaths per 100k" = signif(new_deaths_7/100000), 3) %>%
-  mutate("Daily Deaths per Capita" = signif(covid_state$new_deaths_7/covid_state$population), 3) %>%
-  rename("State" = Province_State) %>%
-  mutate("7-Day Average Cases" = round(new_confirmed_7)) %>%
[…]
